### 2021-11-30 analyses.RmD
- Implemented several changes after ICA - Imputed missing data

---
title: "Analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    # code_folding: hide
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
  # pdf_document:
  #   toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
options(digits = 3, width = 200, scipen=999)
```

# Set-up

Load packages, and make repo replicable.

```{r r-setup, results='hide'}
# install packages
# devtools::install_github("https://github.com/tdienlin/td")

# define packages
packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
              "gridExtra", "haven", "kableExtra", "knitr", 
              "lavaan", "lme4", "magrittr", "mice", "mvnormalTest", "PerFit", 
              "psych", "quanteda.textstats",
              "pwr", "purrr", "semPlot", "semTools", "tidyverse")

# renv::install(packages)
# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install

# load packages
lapply(c(packages, "td"), library, character.only = TRUE)
```

## Custom functions

Find variables in dataset.

```{r}
find_var <- function(name, data = doc)(
  # finds the variables names for an item for each wave
    data %>% 
    filter(Label == name) %>% 
    select(variable) %>% 
    unlist() %>% 
    set_names(sub("\\_.*", "", .))
)
```

Extract characteristics from fitted lmer models.

```{r}
get_specs <- function(model){
    # Get mean, max, and min values

    dat <- coefficients(model)$wave
    specs <- data.frame(
    sd = attr(VarCorr(model), "sc"),
    min = min(dat),
    max = max(dat),
    mean = mean(dat$`(Intercept)`)
  )
}
```

Get data from fitted lmer objects for descriptives.

```{r}
get_dat <- function(model){
  coefficients(model)$wave %>% 
  tibble::rownames_to_column("wave") %>% 
  rename(value = "(Intercept)") %>% 
  mutate(wave = as.integer(.$wave),
         value = as.numeric(.$value)) 
} 
```

Determine average reliability for measures across waves

```{r}
get_rel <- function(data, waves=24){

  # extract average reliability from lavaan fitted cfa with several groups
  reliability(data) %>% 
  unlist() %>% 
  matrix(5, waves) %>% 
  as.data.frame() %>% 
  set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
  summarise(omega = rowMeans(.["omega",])) %>% 
  return()
}
```

Make graph of variables' development.

```{r}
make_graph <- function(model, title, ll, ul, line = FALSE, labels = FALSE, lmer=TRUE, legend=TRUE){

    if(isTRUE(lmer)){
      dat <- get_dat(model)
    } else{
    dat <- model
  }

  graph <- 
    ggplot(dat, aes(wave, value, color = dimension)) +
    {if(legend) theme(
      legend.position = c(0.85, 0.8),
      legend.title = element_blank()
      )} + 
    {if(!legend) theme(
      legend.position = "none"
    )} +
    theme(axis.title.x=element_blank()) +
    coord_cartesian(ylim = c(ll, ul)) +
    ggtitle(title) +
    {if(line) geom_smooth(se = FALSE, method = 'loess')} +
    # geom_point() +
    scale_color_manual(values=c("dodgerblue3", "deepskyblue2", "magenta2", "green2", "red"))
  graph
}
```

Get data of lmer objects for results.

```{r}
# get data
get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, type){

  dat_fig_results <- 
  broom.mixed::tidy(data_aff_neg, conf.int = T) %>% 
  mutate(dv = "aff_neg") %>% 
  rbind(
    broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
    mutate(dv = "aff_pos")
  ) %>% 
  rbind(
    broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
    mutate(dv = "life_sat")
  ) %>% 
  mutate(
    type = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
    iv = gsub("_(w|b)\\>", "", .$term)
         ) %>%
  mutate(
    type = factor(.$type, 
                  levels = c("within", "between")),
    dv = factor(.$dv, 
                levels = c("life_sat", "aff_pos", "aff_neg"),
                labels = c("Life satisfaction", "Positive affect", "Negative affect"))
  ) %>% 
  select(dv, iv, type, estimate, conf.low, conf.high, p.value)

  # select Social Media type of activity
  if(type == "activity") {
    dat_fig_results %<>%
    filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
    mutate(
      iv = factor(.$iv, 
                  levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
                  labels = c("Posting", "Liking & Sharing", "Reading"))
    )
  } else if(type == "channel"){
    dat_fig_results %<>%
    filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
    mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
  } else if(type == "control"){
    dat_fig_results %<>%
      filter(iv %in% c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem")) %>% 
      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem"), 
                         labels = c("Male", "Health", "Internal locus of control", "Employed public service", "Sport", "Satisfaction democracy")))
  }
  return(dat_fig_results)
}

```

Make graph of effects

```{r}
make_graph_res <- function(data_1, data_2 = NULL, sesoi = NULL){
  ggplot(data_1, aes(x = estimate, y = iv)) + 
  scale_color_manual(values = c("grey40", "deepskyblue")) +
  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                lwd = .75, height = .2) +
  geom_point(size = 2) +
  {if(!is.null(data_2)) geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), data_2,
                lwd = .75, height = .2, color = "grey",
                position = position_nudge(y = -.3)
                )} +
  {if(!is.null(data_2)) geom_point(data = data_2, size = 2, color = "grey",
             position = position_nudge(y = -.3)
             )} +
  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    panel.spacing = unit(.9, "lines"),
    text = element_text(size = 12),
    axis.text.y = element_text(hjust = 0)
    ) +
  facet_grid(rows = vars(type),
             cols = vars(dv),
             scales = "free",
             space = "free_y") 
}
```

Extract factor scores.

```{r}
get_fs <- function(model) {
  # the aim is to get factor scores on original scaling
  # hence export factor values for all items
  # then compute average
  dat_ov <- lavaan::lavPredict(model, type = "ov", assemble = TRUE) %>% 
    mutate(fs = rowMeans(.[1:3]))
  return(dat_ov$fs)
}
```

# Data Wrangling

```{r}
doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
d_raw <- read_csv("data/10094_da_de_v4_0.csv")
```

Identify the names of each item for each wave.

```{r}
# identify variable names of items, using custom function
health <- find_var("Gesundheitszustand") # wsn't collected at wave 21
aff_neg_1 <- find_var("Depressivitaet: einsam")
aff_neg_2 <- find_var("Depressivitaet: aergerlich")
aff_neg_3 <- find_var("Depressivitaet: so niedergeschlagen")
aff_neg_4 <- find_var("Depressivitaet: sehr nervoes")
aff_neg_5 <- find_var("Depressivitaet: aengstlich")
aff_neg_6 <- find_var("Depressivitaet: bedrueckt und traurig")
aff_pos_1 <- find_var("Depressivitaet: ruhig und gelassen")
aff_pos_2 <- find_var("Depressivitaet: gluecklich")
aff_pos_3 <- find_var("Depressivitaet: voller Energie")
act_wrk <- find_var("Zuhause verlassen: Arbeit")
act_spo <- find_var("Zuhause verlassen: Sport")
act_frn <- find_var("Zuhause verlassen: Freunde oder Verwandte treffen")
act_sho <- find_var("Zuhause verlassen: Essen einkaufen")
act_pet <- find_var("Zuhause verlassen: Haustier ausfuehren")
sat_dem <- find_var("Demokratiezufriedenheit: Oesterreich")
work_h <- find_var("Arbeitsstunden: Jetzt pro Woche")
work_homeoff <- find_var("Aenderung berufliche Situation: Home-Office")
hh_income <- find_var("Aktuelles Haushaltseinkommen")
med_txt_kro <- find_var("Mediennutzung: Kronen Zeitung oder www.krone.at")
med_txt_sta <- find_var("Mediennutzung: Der Standard oder derstandard.at")
med_txt_pre <- find_var("Mediennutzung: Die Presse oder diepresse.com")
med_txt_oes <- find_var("Mediennutzung: Oesterreich oder oe24.at")
med_txt_kur <- find_var("Mediennutzung: Kurier oder kurier.at")
med_txt_slz <- find_var("Mediennutzung: Salzburger Nachrichten oder salzburg.at")
med_txt_son <- find_var("Mediennutzung: Sonstige oesterreichische Tageszeitungen")
med_vid_orf <- find_var("Mediennutzung: ORF (Nachrichten)")
med_vid_pri <- find_var("Mediennutzung: Privatfernsehen (Nachrichten)")
soc_med_fb <- find_var("Soziale Medien: Facebook")
soc_med_tw <- find_var("Soziale Medien: Twitter")
soc_med_ig <- find_var("Soziale Medien: Instagram")
soc_med_yt <- find_var("Soziale Medien: Youtube")
soc_med_wa <- find_var("Soziale Medien: WhatsApp")
soc_med_read <- find_var("Soziale Medien Aktivitaet: Postings zu Corona anderer lesen")
soc_med_like_share <- find_var("Soziale Medien Aktivitaet: Postings liken, teilen oder retweeten")
soc_med_post <- find_var("Soziale Medien Aktivitaet: selber Postings zu Corona verfassen")
life_sat <- find_var("Lebenszufriedenheit")
risk_prop <- find_var("Risikobereitschaft")
loc_cntrl_int_1 <- find_var("Psychologie: habe Leben selbst in der Hand")
loc_cntrl_int_2 <- find_var("Psychologie: Belohnung durch Anstrengung")
loc_cntrl_int_3 <- find_var("Psychologie: Fremdbestimmung")
loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
```

Select variables.

```{r}
d_wide <- d_raw %>% 
  select(
    id = RESPID,
    gender = SD_GENDER,
    acc_bal = SD_ACCESS_BALCONY,
    acc_gar = SD_ACCESS_GARDEN,
    year_birth = SD_BIRTHYEAR,
    born_aus = SD_BORN_AUSTRIA,
    born_aus_prnts = SD_MIGR_BACKGR,
    county = SD_BULA,
    edu = SD_EDU,
    employment = SD_EMPLSTATUS_FEB2020,
    hh_adults = SD_HH_ADULTS,
    hh_child18 = SD_HH_CHILD18,
    hh_child17 = SD_HH_TEENS,
    hh_child14 = SD_HH_CHILD14,
    hh_child5 = SD_HH_CHILD5,
    hh_child2 = SD_HH_CHILD2,
    hh_oldfam = SD_HH_OLDERFAM,
    hh_outfam = SD_HH_OUTERFAM,
    hh_partner = SD_HH_PARTNER,
    # hh_income = SD_HHINCOME_FEB2020,
    home_sqm = SD_HOME_SQM,
    home_owner = SD_HOMEOWNER,
    # work_h = SD_WORKHOURS_FEB2020,
    health = all_of(health),
    life_sat = all_of(life_sat),
    aff_neg_1 = all_of(aff_neg_1),
    aff_neg_2 = all_of(aff_neg_2),
    aff_neg_3 = all_of(aff_neg_3),
    aff_neg_4 = all_of(aff_neg_4),
    aff_neg_5 = all_of(aff_neg_5),
    aff_neg_6 = all_of(aff_neg_6),
    aff_pos_1 = all_of(aff_pos_1),
    aff_pos_2 = all_of(aff_pos_2),
    aff_pos_3 = all_of(aff_pos_3),
    act_wrk = all_of(act_wrk),
    act_spo = all_of(act_spo),
    act_frn = all_of(act_frn),
    act_sho = all_of(act_sho),
    act_pet = all_of(act_pet),
    sat_dem = all_of(sat_dem),
    sat_dem = all_of(sat_dem),
    work_h = all_of(work_h),
    work_homeoff = all_of(work_homeoff),
    hh_income = all_of(hh_income),
    med_txt_kro = all_of(med_txt_kro),
    med_txt_sta = all_of(med_txt_sta),
    med_txt_pre = all_of(med_txt_pre),
    med_txt_oes = all_of(med_txt_oes),
    med_txt_kur = all_of(med_txt_kur),
    med_txt_slz = all_of(med_txt_slz),
    med_txt_son = all_of(med_txt_son),
    med_vid_orf = all_of(med_vid_orf),
    med_vid_pri = all_of(med_vid_pri),
    soc_med_fb = all_of(soc_med_fb),
    soc_med_tw = all_of(soc_med_tw),
    soc_med_ig = all_of(soc_med_ig),
    soc_med_yt = all_of(soc_med_yt),
    soc_med_wa = all_of(soc_med_wa),
    soc_med_like_share = all_of(soc_med_like_share),
    soc_med_read = all_of(soc_med_read),
    soc_med_post = all_of(soc_med_post),
    risk_prop = all_of(risk_prop),
    loc_cntrl_int_1 = all_of(loc_cntrl_int_1),
    loc_cntrl_int_2 = all_of(loc_cntrl_int_2),
    loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
    loc_cntrl_int_4 = all_of(loc_cntrl_int_4)
  )

# rename variables
# names(d_wide) <- sub("\\.\\.\\.", "§", names(d_wide)) # replace "..." with "_"
```

Make new documentation with selected variables.

```{r}
doc_selected <- filter(doc, variable %in% c(
  "RESPID",
  "SD_GENDER",
  "SD_ACCESS_BALCONY",
  "SD_ACCESS_GARDEN",
  "SD_BIRTHYEAR",
  "SD_BORN_AUSTRIA",
  "SD_MIGR_BACKGR",
  "SD_BULA",
  "SD_EDU",
  "SD_EMPLSTATUS_FEB2020",
  "SD_HH_ADULTS",
  "SD_HH_CHILD18",
  "SD_HH_TEENS",
  "SD_HH_CHILD14",
  "SD_HH_CHILD5",
  "SD_HH_CHILD2",
  "SD_HH_OLDERFAM",
  "SD_HH_OUTERFAM",
  "SD_HH_PARTNER",
  "SD_HHINCOME_FEB2020",
  "SD_HOME_SQM",
  "SD_HOMEOWNER",
  "SD_WORKHOURS_FEB2020",
  health[1],
  life_sat[1],
  aff_neg_1[1],
  aff_neg_2[1],
  aff_neg_3[1],
  aff_neg_4[1],
  aff_neg_5[1],
  aff_neg_6[1],
  aff_pos_1[1],
  aff_pos_2[1],
  aff_pos_3[1],
  act_wrk[1],
  act_spo[1],
  act_frn[1],
  act_sho[1],
  act_pet[1],
  sat_dem[1],
  work_h[1],
  work_homeoff[1],
  hh_income[1],
  med_txt_kro[1],
  med_txt_sta[1],
  med_txt_pre[1],
  med_txt_oes[1],
  med_txt_kur[1],
  med_txt_slz[1],
  med_txt_son[1],
  med_vid_orf[1],
  med_vid_pri[1],
  soc_med_fb[1],
  soc_med_tw[1],
  soc_med_ig[1],
  soc_med_yt[1],
  soc_med_wa[1],
  soc_med_like_share[1],
  soc_med_post[1],
  soc_med_read[1],
  risk_prop[1],
  loc_cntrl_int_1[1],
  loc_cntrl_int_2[1],
  loc_cntrl_int_3[1],
  loc_cntrl_int_4[1]
  ))
```

Recode Variables

```{r}
d_wide %<>%
  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
  mutate(
    male = 2 - .$gender,
    age = 2021 - .$year_birth,
    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
    born_aus = 2 - .$born_aus,
    home_owner = 2 - .$home_owner,
    employment_fac = factor(.$employment, 
                           # labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft",
                           #            "Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
                           #            "Selbststaendig erwerbstaetig",
                           #            "Arbeitslos",
                           #            "Pensionist*in",
                           #            "Ausschliesslich haushaltsfuehrend",
                           #            "Schueler/in, Student/in, Bildungskarenz",
                           #            "Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
                           #            "In Elternkarenz")
                           labels = c("Unemployed",
                                      "Industrie",
                                      "Public service",
                                      "Self-employed",
                                      "Retired",
                                      "Housekeeping",
                                      "Student",
                                      "Incapacitated",
                                      "Parental Leave"),
                            levels = c(4, 1:3, 5:8, 10) # make unemployment reference cat
                           ),
    edu_fac = factor(.$edu,
                     # labels = c("Volksschule oder weniger", 
                     #            "Hauptschule oder AHS Unterstufe",
                     #            "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
                     #            "Lehre, Berufsschule",
                     #            "AHS mit Matura",
                     #            "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
                     #            "Hochschulverwandte Lehranstalt oder Kolleg",
                     #            "Bachelor",
                     #            "Magister / Master / Diplomingenieur / Fachhochschule",
                     #            "Doktor / PhD"),
                     labels = c("No degree", 
                                "Middle school",
                                "Vocational school",
                                "Technical school",
                                "High school",
                                "Applied high school",
                                "State college",
                                "Bachelor",
                                "Master",
                                "PhD")
                     ))
```

Make data from wide to long.

```{r}
d_long <-
  d_wide %>% 
  pivot_longer(
    cols = health...W1:loc_cntrl_int_4...W9,
    names_to = "item",
    values_to = "value"
  ) %>% 
  separate(item, c("item", "wave"), sep = "\\.\\.\\.", extra = "merge") %>% 
  pivot_wider(names_from = "item", values_from = "value")
```

<!-- Filter waves without data -->

<!-- ```{r eval=F} -->
<!-- d_long_test <- d_long %>% -->
<!--   filter(!is.na(.[vars_varying])) -->
<!-- ``` -->

Recode values

```{r}
# recode such that higher values imply more strength / align with wording
d_long %<>% 
  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem, soc_med_fb:soc_med_post), 
            funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
  mutate_at(vars(loc_cntrl_int_1:loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L))) %>% 
  mutate_at(vars(born_aus_prnts), 
            funs(recode(., `3` = 0L, `2` = 2L, `1` = 1L)))

# recode inverted items
d_long %<>%
  mutate_at(vars(loc_cntrl_int_3, loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L)))

# recode other
d_long %<>%
  mutate(
    wave = gsub("W", "", .$wave) %>% as.integer(),
    id = as.integer(id)
  )
```

Order dataset. First id, then wave.

```{r}
d_long %<>%
  arrange(id, wave)
```

<!-- Create vectors with names -->

<!-- ```{r} -->
<!-- # create vector of names -->
<!-- vars_varying <- names(select(d_long, "health_b":"loc_cntrl_int_4_w")) -->
<!-- predictors_varying <- select(d_long, "health_b":"loc_cntrl_int_4_w") %>%  -->
<!--   select(-starts_with("aff_"), -starts_with("life_")) %>%  -->
<!--   names() -->
<!-- ``` -->

# Impute data

Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is not executed but loaded from memory. To execute, set eval == TRUE.

```{r eval=TRUE}
vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem")

# filter respondents with more than 50% missing data
d_long_filt <- 
  d_long %>% 
  mutate(na_perc = rowSums(is.na(select(., all_of(vars_used))) / ncol(select(., all_of(vars_used))))) %>% 
  filter(na_perc < .5)

# impute missing data
# exclude social media use data, because they were measured only on selected waves and would lead to the imputation of too much missing data
vars_excl <- c("na_perc", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
incl_ma <- d_long_filt %>% 
  mutate(across(.cols = everything(), .fns = is.na))
incl_ma[vars_excl] <- FALSE

# init <- mice(select(d_long_test_filt, maxit = 0))
# pred_ma <- init$predictorMatrix
# pred_ma[, vars_excl] <- 0
d_long_imp <- mice(d_long_filt, method = "pmm", m = 1, maxit = 30, 
              where = incl_ma,
              # predictorMatrix = pred_ma, 
              seed = 180719, print = FALSE)
d_long_imp <- mice::complete(d_long_imp)

write_csv(d_long_imp, "data/data_imputed.csv")
```

Load imputed data from disc.

```{r}
d_long_imp <- read_csv("data/data_imputed.csv")
```

# Create within-between data

Create means for scales

```{r}
d_long_imp %<>%
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
```

Create mean values for each person (between), and unique changes per wave (within).

```{r}
d_long_imp %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

And because later we want to rerun the results without imputed data, now the same for data without imputation.

```{r}
d_long %<>%
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

# Measures
## Well-being
### Life satisfaction

```{r}
model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_life_sat)
```

### Positive affect

Positive affect was measured as a scale. We hence first inspect factorial validity using CFA.

We first test assumption of multivariate normality.

```{r eval=FALSE}
mardia(d_long_imp %>% select(aff_pos_1, aff_pos_2, aff_pos_3))
```

Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.

```{r}
model <- "
aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
"
cfa_aff_pos <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
```

The data fit the model very well, `#td::fit_txt(cfa_aff_pos)`.
Let's next inspect reliability

```{r}
rel_aff_pos <- get_rel(cfa_aff_pos)
rel_aff_pos
```

The average reliability across all waves was omega = `r rel_aff_pos`, hence good.

Let's next expect the development across waves.

```{r}
model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_aff_pos)
```

Let's now export factor scores.

```{r}
d_long_imp$aff_pos_fs <- get_fs(cfa_aff_pos)
```

### Negative affect

Negative affect was measured as a scale. We hence inspect factorial validty using CFA.

We first test assumption of multivariate normality. (Warning, as these analyses take long and might require additional working memory, they're deactived currently.)

```{r eval=FALSE}
mardia(d_long_imp %>% select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6))
```

Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.

```{r}
model <- "
aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
"
cfa_aff_neg <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
```

The data fit the model very well, `#td::fit_txt(cfa_aff_neg)`.

Let's next inspect reliability.

```{r}
rel_aff_neg <- get_rel(cfa_aff_neg)
rel_aff_neg
```

The average reliability across all waves was omega = `r rel_aff_neg`, hence good.

```{r}
model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_aff_neg)
```

Let's now export factor scores.

```{r}
d_long_imp$aff_neg_fs <- get_fs(cfa_aff_neg)
```

## Social media use
### Reading

```{r}
model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_read)
```

### Liking & sharing

```{r}
model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_like_share)
```

### Posting

```{r}
model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_post)
```

## Social media channels
### Facebook

```{r}
model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_fb)
```

### Twitter

```{r}
model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_tw)

# Bayesian Approach
# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp
# Also throws a warning
# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp, control=list(adapt_delta=0.82))
# Currently doesn't work, likely for lack of memory
```

Interestingly, lme4 throws a warning, likely because there is too little variance across waves, while little variation measure itself. Let's inspect the raw means.

```{r}
soc_med_tw_m <- d_long_imp %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
soc_med_tw_m
```

Inspecting the means shows that the data are intact, only that the estimation didn't fully work.

### Instagram

```{r}
model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_ig)
```

### WhatsApp

```{r}
model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_wa)
```

### YouTube

```{r}
model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_imp)
summary(model_soc_med_yt)
```

## Locus of control

The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.

```{r}
model <- "
loc_cntrl_int =~ a1*loc_cntrl_int_1 + a2*loc_cntrl_int_2 + a3*loc_cntrl_int_3 + a4*loc_cntrl_int_4
# loc_cntrl_int_1 ~~ loc_cntrl_int_2
loc_cntrl_int_3 ~~ loc_cntrl_int_4
"
cfa_loc_cntrl_int <- cfa(model, filter(d_long_imp, wave != 20), group = "wave")
summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
```

The data fit the model very well, `#td::fit_txt(cfa_loc_cntrl_int)`.

## Figure

Display the results in a combined figure.

```{r}
fig_desc_dat <- data.frame(
  rbind(
    data.frame(type = "Life satisfaction", dimension = "Life satisfaction", get_dat(model_life_sat)),
    data.frame(type = "Affect", dimension = "Positive", get_dat(model_aff_pos)),
    data.frame(type = "Affect", dimension = "Negative", get_dat(model_aff_neg)),
    data.frame(type = "Social media use", dimension = "Reading", get_dat(model_soc_med_read)),
    data.frame(type = "Social media use", dimension = "Liking & Sharing", get_dat(model_soc_med_like_share)),
    data.frame(type = "Social media use", dimension = "Posting", get_dat(model_soc_med_post)),
    data.frame(type = "Social media channel", dimension = "Facebook", get_dat(model_soc_med_fb)),
    data.frame(type = "Social media channel", dimension = "Twitter", get_dat(model_soc_med_tw)),
    data.frame(type = "Social media channel", dimension = "Instagram", get_dat(model_soc_med_ig)),
    data.frame(type = "Social media channel", dimension = "WhatsApp", get_dat(model_soc_med_wa)),
    data.frame(type = "Social media channel", dimension = "YouTube", get_dat(model_soc_med_yt))
  ) %>% 
    mutate(
      type = factor(.$type, levels = c("Life satisfaction", "Affect", "Social media use", "Social media channel")),
      dimension = factor(.$dimension, levels = c("Life satisfaction", "Positive", "Negative", "Reading", "Liking & Sharing", "Posting", "Facebook", "Twitter", "Instagram", "WhatsApp", "YouTube"))
      )
)

fig_desc_life_sat <- make_graph(
  fig_desc_dat %>% filter(type == "Life satisfaction"), 
  title = "Life satisfaction", 
  ll = 0, ul = 10, 
  lmer = FALSE, 
  line = TRUE,
  legend = FALSE
  )

fig_desc_aff <- make_graph(
  fig_desc_dat %>% filter(type == "Affect"), 
  title = "Affect", 
  ll = 1, ul = 5, 
  lmer = FALSE, 
  line = TRUE
  )

fig_desc_soc_med_use <- make_graph(
  fig_desc_dat %>% filter(type == "Social media use"), 
  title = "Social media use", 
  ll = 1, ul = 5, 
  lmer = FALSE, 
  line = TRUE
  )

fig_desc_soc_med_channel <- make_graph(
  fig_desc_dat %>% filter(type == "Social media channel"), 
  title = "Social media channel", 
  ll = 1, ul = 5, 
  lmer = FALSE, 
  line = TRUE
  )

fig_desc <- grid.arrange(fig_desc_life_sat, fig_desc_aff, 
                         fig_desc_soc_med_use, fig_desc_soc_med_channel,
                         nrow = 2, ncol = 2)

ggsave("figures/fig_descriptives.pdf", 
       width = 10, height = 6,
       plot = fig_desc)
```

## Table

```{r}
tab_desc_dat <- rbind(
    "Life satisfaction" = get_specs(model_life_sat),
    "Positive affect" = get_specs(model_aff_pos),
    "Negative affect" = get_specs(model_aff_neg),
    "Read" = get_specs(model_soc_med_read),
    "Like & share" = get_specs(model_soc_med_like_share),
    "Posting" = get_specs(model_soc_med_post),
    "Facebook" = get_specs(model_soc_med_fb),
    "Twitter" = c(sd = get_specs(model_soc_med_tw)$sd,
                  min = min(soc_med_tw_m$value, na.rm = TRUE),
                  max = max(soc_med_tw_m$value, na.rm = TRUE),
                  mean = mean(soc_med_tw_m$value, na.rm = TRUE)
                  ),
    "Instagram" = get_specs(model_soc_med_ig),
    "WhatsApp" = get_specs(model_soc_med_wa),
    "YouTube" = get_specs(model_soc_med_yt)
    )
kable(tab_desc_dat)
```

# Analyses
## Life satisfaction

First, without control variables.

```{r}
model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
                            (1 | id) + (1 | wave) + 
                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
                data = d_long_imp)
summary(model_life_sat_lmer_nco)
```

Now with control variables.

```{r}
model_life_sat_lmer <- lmerTest::lmer(life_sat ~
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        # hh_income_b + # large dropout (900 obs)
                                        # hh_adults + # too much dropout
                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
                                        # home_owner + # too much dropout
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_life_sat_lmer)
```

## Positive Affect

First, without control variables.

```{r}
model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
                                      , data = d_long_imp)
summary(model_aff_pos_lmer_nco)
```

Now with control variables.

```{r}
model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        # hh_income_b + # large dropout (900 obs)
                                        # hh_adults + # too much dropout
                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
                                        # home_owner + # too much dropout
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_pos_lmer)
```

## Negative Affect

First, without control variables.

```{r}
model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
                                      , data = d_long_imp)
summary(model_aff_neg_lmer_nco)
```

Then, with control variables.

```{r}
model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        # hh_income_b + # large dropout (900 obs)
                                        # hh_adults + # too much dropout
                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
                                        # home_owner + # too much dropout
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_neg_lmer)
```

# Figures

First, for activities.

```{r}
# get data
dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity")
dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity")

# make figure
fig_results_activity <- make_graph_res(
  data_1 = dat_fig_results_activity,
  data_2 = dat_fig_results_activity_nco,
  sesoi = "est"
  )
fig_results_activity

# safe figure
ggsave("figures/fig_results_activity.pdf", 
       width = 7, height = 4,
       plot = fig_results_activity)
```

Then for channels.

```{r}
# get data
dat_fig_results_channel <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channel")
dat_fig_results_channel_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channel")

# make figure
fig_results_channel <- make_graph_res(
  data_1 = dat_fig_results_channel,
  data_2 = dat_fig_results_channel_nco,
  sesoi = "est"
  )
fig_results_channel

# save figure
ggsave("figures/fig_results_channel.pdf", 
       width = 7, height = 4,
       plot = fig_results_channel)
```

Select control variables and make graph.

```{r}
# make figure
fig_results_control <- make_graph_res(
  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "control")
  )
fig_results_control

# save figure
ggsave("figures/fig_results_control.pdf", 
       width = 7, height = 4,
       plot = fig_results_control)
```

```{r}
save.image("data/workspace_1.RData")
```

### 2021-11-30 analyses_additional.Rmd
- Implemented several changes after ICA - Imputed missing data

---
title: "Additional analyses"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    # code_folding: hide
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
  # pdf_document:
  #   toc: TRUE
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
options(digits = 3, width = 200, scipen=999)
```

# Set-up

Load packages, and make repo replicable.

```{r r-setup, results='hide'}
# install packages
# devtools::install_github("https://github.com/tdienlin/td")

# define packages
packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
              "gridExtra", "haven", "kableExtra", "knitr", 
              "lavaan", "lme4", "magrittr", "mice", "mvnormalTest", "PerFit", 
              "psych", "quanteda.textstats",
              "pwr", "purrr", "semPlot", "semTools", "tidyverse")

# renv::install(packages)
# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install

# load packages
lapply(c(packages, "td"), library, character.only = TRUE)

# load workspace
load("data/workspace_1.RData")
```

# Standardized results

First, create standardized dataset.

```{r}
d_long_std <- d_long_imp %>% 
  mutate(across(c(-id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), scale))
```

Then let's rerun the analyses.

## Life satisfaction

```{r}
model_life_sat_lmer_std <- lmerTest::lmer(life_sat ~
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_std)
summary(model_life_sat_lmer_std)
```

## Positive affect

```{r}
model_aff_pos_lmer_std <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_pos_lmer_std)
```

## Negative Affect

```{r}
model_aff_neg_lmer_std <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_neg_lmer_std)
```

## Figures

First, for activities.

```{r}
# make figure
fig_results_activity_std <- make_graph_res(
  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "activity"),
  sesoi = "std"
  )
fig_results_activity_std

# safe figure
ggsave("figures/fig_results_activity_std.pdf", 
       width = 7, height = 4,
       plot = fig_results_activity_std)
```

Then for channels.

```{r}
# make figure
fig_results_channel_std <- make_graph_res(
  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "channel"),
  sesoi = "std"
  )
fig_results_channel_std

# save figure
ggsave("figures/fig_results_channel_std.pdf", 
       width = 7, height = 4,
       plot = fig_results_channel_std)
```

Select control variables and make graph.

```{r}
# make figure
fig_results_control_std <- make_graph_res(
  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "control"),
  sesoi = "std"
  )
fig_results_control_std

# save figure
ggsave("figures/fig_results_control_std.pdf", 
       width = 7, height = 4,
       plot = fig_results_control_std)
```

# Results with mean scores

Instead of using factor scores, in what follows I report the results with mean scores of affect. Note that the results for life satisfaction were the same, as it was measured with a single item.

## Positive affect

```{r}
model_aff_pos_lmer_m <- lmerTest::lmer(aff_pos_m ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_pos_lmer_std)
```

The results differed only slightly, and all inferences remained the same.

## Negative Affect

```{r}
model_aff_neg_lmer_m <- lmerTest::lmer(aff_neg_m ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long_imp)
summary(model_aff_neg_lmer_std)
```

The results differed only slightly, and all inferences remained the same.

# Results without imputed data

We first need to export factor scores for variables without imputed data

```{r}
model <- "
aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
"
cfa_aff_pos <- cfa(model, d_long, group = "wave", estimator = "MLM")
d_long$aff_pos_fs <- get_fs(cfa_aff_pos)

model <- "
aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
"
cfa_aff_neg <- cfa(model, d_long, group = "wave", estimator = "MLM")
d_long$aff_neg_fs <- get_fs(cfa_aff_neg)
```

## Life satisfaction

First, without control variables.

```{r}
model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
                            (1 | id) + (1 | wave) + 
                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
                data = d_long)
summary(model_life_sat_lmer_nco)
```

Now with control variables.

```{r}
model_life_sat_lmer <- lmerTest::lmer(life_sat ~
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long)
summary(model_life_sat_lmer)
```

## Positive Affect

First, without control variables.

```{r}
model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
                                      , data = d_long)
summary(model_aff_pos_lmer_nco)
```

Now with control variables.

```{r}
model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long)
summary(model_aff_pos_lmer)
```

## Negative Affect

First, without control variables.

```{r}
model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
                                      , data = d_long)
summary(model_aff_neg_lmer_nco)
```

Then, with control variables.

```{r}
model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
                                        (1 | id) + (1 | wave) + 
                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
                                        res_vienna + acc_bal + acc_gar + home_sqm + 
                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
                                        med_vid_orf_b + med_vid_pri_b + 
                                        risk_prop_b + loc_cntrl_int_m_b + 
                                        act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
                                        act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
                                        sat_dem_w + sat_dem_b 
                                      , data = d_long)
summary(model_aff_neg_lmer)
```

```{r}
save.image("data/workspace_2.RData")
```

### 2021-11-30 index.Rmd
- Implemented several changes after ICA - Imputed missing data

---
title: "No meaningful effects of COVID-19 related social media use on well-being"
---

On this website, you can find the companion material for the paper _No meaningful effects of COVID-19 related social media use on well-being_.

You can find the following information here:

- The paper, which you can read [in your browser](manuscript.html) or download as [pdf](manuscript.pdf). 
- The detailed code and summaries of the [analyses](analyses.html)
- [Additional analyses](analyses_additional.html).
- A list of the [items translated into English](items_english.html)
- A list of the [original items in German](items_german.html)
- The [preregistration](preregistration.html) 
- A list where I needed to [deviate](preregistration_changes.Rmd) from the preregistration.

This website is produced directly from the project's [github repository](https://github.com/tdienlin/Austrian_Corona_Panel_Project). You can download the material and the data, file an issue, or submit a pull request.

### 2021-11-30 items_english.rmd
- Implemented several changes after ICA - Imputed missing data

---
title: "Items: English"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

Here you can find a quick overview of all items that were measured.

## Life Satisfaction

"Taken everything together, how satisfied are you currently with your life?"

(0 = extremely unsatisfied, 10 = extremely satisfied)

## Positive Affect (scale)

"In the last week, how often did you feel"

 - a: ... calm and relaxed
 - e: ... happy
 - i: ... full of energy

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

## Negative Affect (scale)

"In the last week, how often did you feel"

 - b: ... lonely
 - c: ... aggravated
 - d: ... so depressed, that nothing could lift you up
 - f: ... very nervous
 - g: ... anxious
 - h: ... glum and sad

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

## Social media use

"How often during the last week have you engaged in the following activities on social media?"

(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)

### Reading

 - "Reading the posts of others with content on the Coronavirus."

### Liking & Sharing

 - "When seeing Posts on the Coronavirus, I clicked "like", "share" or "retweet".

### Active Postings

 - "I myself wrote posts on the Coronavirus on Social Media."

### Specific platforms

"How often in the last week have you followed information related to the Corona-crisis in the following social media?"

 - Facebook 
 - Twitter
 - Instagram
 - Youtube
 - WhatsApp

(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)

## Gender

"What is your biological sex?"

(a: Male; b Female; c: Diverse)

## Age

"In which year were you born?"

(year)

## Education

"What is the highest educational degree you have earned?"

 1. No degree
 1. Middle school
 1. Vocational school
 1. Technical school
 1. High school
 1. Applied high school
 1. State college
 1. Bachelor
 1. Master
 1. PhD

## Austria country of birth

"Is Austria you country of birth?"

  - Yes
  - NO

## Austria country of birth of parents

 "Is Austria your parents' country of birth?"

 - a. Yes, one parent.
 - b. Yes, both parents.
 - c. No.

## News consumption: Text-based

"Below you can find a list of different newspapers. How often during the last week have you informed yourself politically using on the these outlets (both online and offline)?"

 - Kronen Zeitung or www.krone.at
 - Der Standard or derstandard.at
 - Die Presse or diepresse.com
 - Österreich or oe24.at
 - Kurier or kurier.at
 - Salzburger Nachrichten or salzburg.at
 - Other Austrian papers

## News consumption: Video-based

"Below you can find a list of different news shows in television. How often during the last week have you informed yourself politically using on the these shows?"

 - "News shows on ORF (e.g., Zeit im Bild at 19:30h or ZiB 2 at 22:00h)"
 - "News shows in private television (e.g., ATV aktuell, Puls 4/ Pro7/ Sat1 Austria)"

(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)

## Residency Vienna

"In which county do you live currently?"

 - a. Vorarlberg
 - b. Tirol
 - c. Salzburg
 - d. Steiermark
 - e. Kärnten
 - f. Oberösterreich
 - g. Niederösterreich
 - h. Wien
 - i. Burgenland

(Was transformed to Vienna: yes/no)

## Household size

"Please include yourself into the following calculation. If one groups not present in your household, indicate "0"."

 - a. ...Adults older than 18 years:
 - b. ...Teenager between 15-17 years:
 - c. ...Kids between 6-14 years:
 - d. ...Kids between 3-5 years:
 - e. ...Kids between 0-2 years:

## Health

"Let's next turn to some questions concerning health. How would you describe your overall health conditionas? Would you say your healt is ..."

 - a. 1 = Sehr gut
 - b. 2 = Gut
 - c. 3 = Mittelmäßig
 - d. 4 = Schlecht
 - e. 5 = Sehr schlecht

## Living space

"How many square meteres does your appartment have?"

(Number, min. 0 – max. 600, voluntary.)

## Access garden

"Have you got access to a private area outside?"

 - (Yes, garden)

## Access balcony

"Have you got access to a private area outside?"

 - (Yes, balcony)

## Employment

In Februrary 2020, what was your job situation?

 1. Unemployed
 1. Industrie
 1. Public service
 1. Self-employed
 1. Retired
 1. Housekeeping
 1. Student
 1. Incapacitated
 1. Parental Leave

## Work hours per week

"How many hours per week (including extra hours) do you currently work in your main job? If you don't know, please guess.

(hours)

## Home-Office Yes/No

"Has your current job situation changed due to the Corona crisis? Please select the appropriate answer."

 - a. I'm in home-office

## Household income

"How much income does your household currently have per month?"

 - a. less than 1.100 Euro
 - b. between 1.100 and 1.500 Euro 
 - c. between 1.500 and 1.800 Euro 
 - d. between 1.800 and 2.200 Euro 
 - e. between 2.200 and 2.700 Euro 
 - f. between 2.700 and 3.100 Euro 
 - g. between 3.100 and 3.700 Euro 
 - h. between 3.700 and 4.300 Euro 
 - i. between 4.300 and 5.500 Euro 
 - j. 5.500 Euro and more

## Outdoor activities

"Have you left your home during the last week for the following activity?"

 - a: Work
 - b: Sport
 - c: Meeting friends or relatives
 - e: Shopping
 - f: Pets

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Satisfaction Democracy

"How satisfied or dissatisfied are you all in all with the democracy how it's currently working in Austria?"

 - a. 1 = very satisfied
 - b. 2 = rather satisfied
 - c. 3 = neither nor
 - d. 4 = rather dissatisfied 
 - e. 5 = very dissatisfied

## Risk-taking

"How willing are you, in general, to take risks?"

(0: not at all willing; 10: very willing)

## Locus of Control

 - "I'm in control of my own life."
 - "If I try hard, I'll have success."
 - "Bei it in my private or professional life: My life is primarily determined by others."
 - "My plans for life are determined by fate."

(1 = Agree completely; 2 = Agree somewhat; 3 = Disgagree somewhat; 4 = Disagree completely)

### 2021-11-30 items_german.Rmd
- Implemented several changes after ICA - Imputed missing data

---
title: "Items: German"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

Here you can find a quick overview of all items that were measured.

## Life Satisfaction

"Alles in allem betrachtet, wie zufrieden sind Sie derzeit mit Ihrem Leben?"

(0 = äußerst unzufrieden, 10 = äußerst zufrieden)

## Positive Affect (scale)

"Wie often waren Sie in der letzten Woche"

 - a: ... ruhig und gelassen
 - e: ... glücklich
 - i: ... voller Energie

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Negative Affect (scale)

"Wie often waren Sie in der letzten Woche"

 - b: ... einsam
 - c: ... ärgerlich
 - d: ... so niedergeschlagen, dass Sie nichts aufheitern konnte
 - f: ... sehr nervös
 - g: ... ängstlich
 - h: ... bedrückt und traurig

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Social media use: Reading

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Postings anderer lesen, in denen es um Inhalte zum Coronavirus geht"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

## Social media use: Liking & Sharing

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Bei Postings anderer zum Coronavirus auf den „Gefällt mir“ Button drücken, „teilen“ oder „retweeten“"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Social media use: Active Postings

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Selbst Postings zum Coronavirus in Sozialen Netzwerken verfasst"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Social media use: Specific platforms

"Wie oft in der letzten Woche haben Sie Informationen zur Corona-Krise in den folgenden Sozialen Medien verfolgt?"

 - Facebook 
 - Twitter
 - Instagram
 - Youtube
 - WhatsApp

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

## Gender

"Welches Geschlecht haben Sie?"

(a: Männlich; b Weiblich; c: Divers)

## Age

"In welchem Jahr wurden Sie geboren?"

(Jahr)

## Education

"Was ist der höchste Schul- oder Bildungsabschluss, den Sie erreicht haben?"

 - a. Volksschule oder weniger
 - b. Hauptschule oder AHS Unterstufe
 - c. Polytechnikum, BMS (Fachschule, z.B. HASCH)
 - d. Lehre, Berufsschule
 - e. AHS mit Matura
 - f. BHS mit Matura (z.B. HTL, HAK, HBLA etc.) 
 - g. Hochschulverwandte Lehranstalt oder Kolleg 
 - h. Bachelor
 - i. Magister / Master / Diplomingenieur / Fachhochschule 
 - j. Doktor / PhD

## Austria country of birth

"Ist Ihr Geburtsland Österreich?"

  - Ja
  - Nein

## Austria country of birth of parents

 "Wurde einer oder beide Ihrer Elternteile nicht in Österreich geboren?"

 - a. Ja, einer
 - b. Ja, beide
 - c. Nein

## News consumption: Text-based

"Nachstehend finden Sie eine Liste mit verschiedenen Tageszeitungen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in der jeweiligen Zeitung (im Internet oder gedruckt) informiert?"

 - Kronen Zeitung oder www.krone.at
 - Der Standard oder derstandard.at
 - Die Presse oder diepresse.com
 - Österreich oder oe24.at
 - Kurier oder kurier.at
 - Salzburger Nachrichten oder salzburg.at
 - Sonstige österreichische Tageszeitungen

## News consumption: Video-based

"Nachstehend finden Sie eine Liste mit verschiedenen Nachrichtensendungen im Fernsehen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in den folgenden Nachrichtensendungen informiert?"

 - "Nachrichtensendungen im ORF (z.B. Zeit im Bild um 19:30h oder ZiB 2 um 22:00h)"
 - "Nachrichtensendungen im Privatfernsehen (z.B. ATV aktuell, Puls 4/ Pro7/ Sat1 Austria)"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

## Residency Vienna

"In welchem Bundesland leben Sie aktuell?"

 - a. Vorarlberg
 - b. Tirol
 - c. Salzburg
 - d. Steiermark
 - e. Kärnten
 - f. Oberösterreich
 - g. Niederösterreich
 - h. Wien
 - i. Burgenland

(Will be transformed to Vienna: yes/no)

## Household size

"Bitte rechnen Sie hier sich selbst wieder mit ein. Wenn es eine Personengruppe nicht in Ihrem Haushalt gibt, tragen Sie bitte „0“ ein.

 - a. ...Erwachsene über 18 Jahre:
 - b. ...Teenager von 15-17 Jahren:
 - c. ...Kinder von 6-14 Jahren:
 - d. ...Kinder von 3-5 Jahren:
 - e. ...Kinder von 0-2 Jahren:

(Achtung, wenn Anzahl Erwachsene missing, 1 eintragen, da vorher Filterfrage)

## Health

"Kommen wir jetzt zu ein paar Fragen zu Ihrer Gesundheit. Wie schätzen Sie Ihren allgemeinen Gesundheitszustand ein? Würden Sie sagen, er ist..."

 - a. 1 = Sehr gut
 - b. 2 = Gut
 - c. 3 = Mittelmäßig
 - d. 4 = Schlecht
 - e. 5 = Sehr schlecht

## Living space

"Wie viele Quadratmeter hat Ihre Wohnung?"

(Einfach-Nennung, Zahleneingabe, min. 0 – max. 600, freiwillig)

## Access garden

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?

 - (Ja, Garten)

## Access balcony

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?"

 - (Ja, Balkon)

## Employment

Wenn Sie sich selbst zuordnen: Im Februar 2020, welcher der folgenden Gruppen gehören Sie vorwiegend an? (Einfach-Nennung)

 - a. Unselbstständig erwerbstätig (auch Lehrling), Privatwirtschaft
 - b. Unselbstständig erwerbstätig (auch Lehrling), Öffentlicher Dienst
 - c. Selbstständig erwerbstätig
 - d. Arbeitslos
 - e. Pensionist*in
 - f. Ausschließlich haushaltsführend
 - g. Schüler/in, Student/in, Bildungskarenz
 - h. Dauerhaft arbeitsunfähig
 - i. Präsenz-/Zivildiener
 - j. In Elternkarenz

## Work hours per week

"Wie viele Stunden (inklusive Überstunden) arbeiten Sie in Ihrer Haupttätigkeit jetzt pro Woche? Wenn Sie es nicht genau wissen, genügt Ihre beste Schätzung."

## Home-Office Yes/No

"Hat sich in Ihrer beruflichen Situation aufgrund der Corona-Krise etwas geändert? Bitte wählen Sie alles Zutreffende aus."

 - a. Ich bin im Home-Office

## Household income

"Wie viel Geld steht Ihrem Haushalt aktuell zur Verfügung (Netto-Haushaltseinkommen inklusive Sozialleistungen, Rente usw.)? Bitte rechnen Sie Überstunden mit ein, nicht aber den 13./14. Bezug."

 - a. weniger als 1.100 Euro
 - b. 1.100 bis unter 1.500 Euro 
 - c. 1.500 bis unter 1.800 Euro 
 - d. 1.800 bis unter 2.200 Euro 
 - e. 2.200 bis unter 2.700 Euro 
 - f. 2.700 bis unter 3.100 Euro 
 - g. 3.100 bis unter 3.700 Euro 
 - h. 3.700 bis unter 4.300 Euro 
 - i. 4.300 bis unter 5.500 Euro 
 - j. 5.500 Euro und mehr

## Outdoor activities

"Haben Sie wegen folgenden Gründen in der vergangenen Woche Ihr Zuhause verlassen?"

 - a: Arbeit
 - b: Sport
 - c: Freunde oder Verwandte treffen
 - e: Essen einkaufen
 - f: Haustiere ausführen

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Satisfaction Democracy

"Wie zufrieden oder unzufrieden sind Sie alles in allem mit der Demokratie, so wie sie derzeit in Österreich funktioniert?"

 - a. 1 = Sehr zufrieden
 - b. 2 = Eher zufrieden
 - c. 3 = Teils-teils
 - d. 4 = Eher unzufrieden 
 - e. 5 = Sehr unzufrieden

## Risk-taking

"Wie sehr sind Sie im Allgemeinen dazu bereit, Risiken einzugehen?"

(0: gar nicht bereit; 10: sehr bereit)

## Locus of Control

 - "Ich habe mein Leben selbst in der Hand."
 - "Wenn ich mich anstrenge, werde ich auch Erfolg haben."
 - "Egal ob privat oder im Beruf: Mein Leben wird zum großen Teil von anderen bestimmt."
 - "Meine Pläne werden oft vom Schicksal durchkreuzt."

(1 = Trifft voll und ganz zu; 2 = Trifft etwas zu; 3 = Trifft wenig zu; 4 = Trifft nicht zu)

### 2021-11-30 manuscript.Rmd
- Implemented several changes after ICA - Imputed missing data

---
title             : "No meaningful effects of COVID-19 related social media use on well-being"
shorttitle        : "COVID-19 related social media use and well-being"

author: 
  - name          : "Tobias Dienlin"
    affiliation   : "1"
    corresponding : yes
    address       : "University of Vienna, Department of Communication, 1090 Vienna, Austria"
    email         : "tobias.dienlin@univie.ac.at"

affiliation:
  - id            : "1"
    institution   : "University of Vienna"

authornote: |
  Tobias Dienlin, Department of Communication, University of Vienna.

abstract          : "In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures. To this end, many people use various types of media, and increasingly social media. However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'. In this preregistered study, I investigate whether using social media for COVID-19 related reasons might put personal well-being at risk. To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants. I ran three random effects cross lagged panel models, controlling for several stable and varying covariates. Results showed that the effects of various types of COVID-19 related social media use on well-being were very small, arguably too small to matter. The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal."

bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  - \note{\clearpage}
---

```{r analysis-preferences, include=F, cache=F}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
                      cache=T,
                      echo=F
                      # , fig.pos = 'h'
                      )
```

```{r setup, include=F, cache=F}

# make project reproducible
# devtools::install_github("rstudio/renv")  # install renv
# renv::init()  # initialize renv

# install necessary packages
packages <- c("devtools", "english", "ggplot2", "kableExtra", 
              "lavaan", "lme4", "lmerTest",
              "magrittr", "semPlot", "tidyverse")
# renv::install(packages)
# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install
# renv::install("crsh/papaja@devel")  # uncomment to install

# load packages
lapply(c(packages, "papaja", "td"), library, character.only = TRUE)

# save snapshot of installed packages
# renv::snapshot()

# load workspace created from analyses.rmd to use results
load("data/workspace_2.RData")

# export references
# r_refs(file = "bibliography/r-references.bib")
```

During the COVID-19 pandemic, 
<!-- and especially in the beginning,  -->
numerous events unfolded in quick succession.
Several open questions emerged.
How dangerous is the virus? 
Is it spreading in my region? 
How is it transmitted, and how can I protect myself?
Because for many it was (and continues to be) a matter of life or death, people aimed to stay informed regarding the latest developments.
Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high [@statistaAverageDailyTime2021].

Some people could not stop using social media to learn about COVID-19 related news.
A new phenomenon termed "doomscrolling" emerged:
Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
These concerns seem justified: 
A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p. 283], which implies negative consequences for well-being.

As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide)`.
<!-- At least 1,500 participants took part per wave, and the sample is representative of the Austrian population. -->
The panel study collected a large number of psychological and demographic variables.
<!-- Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design, this creates a unique opportunity  -->
I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.

## Defining Well-being and Media Use

The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
<!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
One of the most prominent indicators of well-being is life satisfaction.
In my view, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
<!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
Together, this should provide an encompassing perspective on potential media effects.

The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
Whereas the first four levels focus on the _channel_, the last two address the _type_ of communication.
To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.

First, I investigate how different types of communication affect well-being.
Specifically, I differentiate between active and passive use.
I distinguish (a) _reading_ (passive), _posting_ (active), and _liking and sharing_ COVID-19 related posts (both active and passive).
Second, I analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
Branded apps are separate entities with potentially different effects. 
Twitter might have a different effect as compared to WhatsApp because of their different affordances.
For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content. 
Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.

## Effects of Social Media on Well-Being

So far, there is only little empirical research on COVID-19 related social media use on well-being.
In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p. 288).
@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
In a sample of 312 respondents collected via Amazon Mechanical Turk, @choiMediatedCommunicationMatters2021 reported that people who used media to attain information were more lonely and less satisfied with their lives.
@stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
Together, the literature emphasizes potentially negative effects of social media as news use [see also @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
However, note that all of these findings represent between-person relation stemming from cross-sectional data (see below).
We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.

The question of whether and how _general_ social media use affect well-being, on the other hand, is well-researched.
This also holds true for the different types of communication such as active or passive use.
A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small [@meierComputermediatedCommunicationSocial2020a]---potentially too small to matter.
But what determines whether or not an effect is considered small or trivial?
If we refer to standardized effect sizes, according to @cohenPowerPrimer1992 small effect sizes start at _r_ = .10.
And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].

These overviews are well aligned with several individual new studies employing advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
In conclusion, most effects are likely somewhere between trivial and small.
I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.

From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.

First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
If one learns about such events, the initial reaction might be shock, fear, or dismay.
Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life.
<!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.
<!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->

There could also be _indirect_ effects.
When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
For example, during the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->

On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might be not the worst idea.
Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
<!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->

Together, the strongest argument seems to be that _in general_ the effects of social media on well-being are, on average, small at best.
Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, and the overall potential of the effects should diminish further.
Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
<!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
Therefore, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.

> Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.

# Current Study
## Smallest Effect Size of Interest

Testing this hypothesis, however, is not trivial.
First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
Effectively testing this hypothesis necessitates defining what's considered a "trivial effect size" and what's not.
Above I already referred to standardized effect sizes.
However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
Standardized effect sizes are determined by a sample's variance.[^cohensd], which is problematic: 
The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
Instead, it should depend on absolute criteria.

What could be a minimally interesting, nontrivial effect?
Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
<!-- In the end, it is a personal question. -->
However, it is still necessary and helpful to try provide such a plausible benchmark. 
I therefore suggest the following SESOI for this research question:

[^cohensd]: Consider the effect size Cohen's _d_: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation [@cohenPowerPrimer1992]. Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.

> SESOI: If a heavy user of COVID-19 related social media news suddenly _stops_ using social media altogether, this should have a _noticeable_ impact on their overall well-being.

What does this mean practically and how can it be operationalized?
In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = _never_ to 5 = _several times a day_. 
Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
But what's a noticeable change in well-being?
According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish _seven_ levels of satisfaction with health.
So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.

## Causality

The hypothesis explicitly states a causal effect.
In non-experimental studies, longitudinal designs help investigate causality.
Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements [@rohrerTheseAreNot2021].
In addition, we for example also need to control for confounding third variables, and importantly also for _varying_ third variables.

To explain, consider the following example.
Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life.
Eventually, use and life satisfaction return to prior levels.
If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.

Thus, only if we can control for _all_ confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
Obviously, we can never be entirely sure that we have included all confounders, which makes absolute statements regarding causality virtually impossible.
In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
Complicating matters, often it is unclear if a variable is a mediator or a confounder.[^collider]
However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
The aim should therefore be to collect as many varying and nonvarying confounders as possible, while knowing that absolute certainty regarding causality cannot be reached.

[^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018].

When searching for suitable candidates for confounders, we look for variables that affect both media use and well-being. 
Controlling for these factors isolates the actual effect of social media use on well-being.
We can also control for variables that affect only social media use or well-being. 
However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].

In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are very likely to do so, and which also are not mediators:
<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
I will not control for variables such as trust in institutions or trust in media, because these variables are likely influenced by social media use to a meaningful extent.

Next to including covariates, it is now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
If a specific person changes their media diet, we need to measure how this behavior affects _their own_ well-being.
Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
In this study, I hence differentiate between-person relations from within-person effects. 
And as explicated above, to test the hypothesis I thus consider only the within-person effects.

Finally, one precondition of causality is temporal order. 
The cause needs to precede the effect.
Finding the right interval between cause and effect is crucial.
For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
<!-- Finding the right interval is difficult. -->
If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are necessary.
In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].

In this study, I therefore adopt an intermediate perspective.
I analyze if when a person changes their social media diet, are there simultaneous changes in their well-being?
When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
<!-- Similar approaches were already implemented by several other studies [@johannesNoEffectDifferent2021; @scharkowHowSocialNetwork2020] and they are considered a best practice to analyze causality [@bellFixedRandomEffects2019]. -->

# Method

In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.

## Preregistration

The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) I preregistered on the Open Science Framework. 
The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
In some cases I could not execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html) (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).

## Sample

The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021].
The study was conducted between March 2020 and October 2021.
It contains 26 waves, and at the time of writing the first 24 waves were available for download. 
Each wave consists of at least 1,500 respondents.
The overall sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, and `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations were collected.
Panel mortality was compensated through a continuous acquisition of new participants. 
All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.

The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level. 
In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
The study received IRB approval from the University of Vienna.
The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.

## Inference criteria

Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
The sample is very large, and it is hence well-equipped reliably to detect also small effects.
<!-- , which is why no exact post hoc power analysis were conducted. -->
In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
On the basis of the SESOI, I then defined a null region (in Bayesian terms, this is called the region of practical equivalence, ROPE).
In what follows, I explain how I determined the SESOI and the null region.

In this study, life satisfaction was measured on an 11-point scale.
If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
In a statistical regression, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
We would therefore expect a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
For affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
Because we are agnostic as to whether the effects are positive or negative, the null region will include negative and positive effects.
Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.[^rounding] 
Together, this leads to a null region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.

[^rounding]: Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects [@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019].

<!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
To illustrate, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is only trivial is supported.
If the effects interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
If the confidence falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.

<!-- Responses were individually checked for patterns such as straight-lining or missing of inverted items. -->
<!-- X clear cases were removed. -->
<!-- Respondents who answered less than 50% of all questions were removed. -->
<!-- The remaining missing responses were imputed using predictive mean matching. -->

## Data Analysis

The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)[@bellFixedRandomEffects2019]. 
Three models were run, one for each dependent variable.
The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
Nesting in participants allowed to separate between-person relations from within-person effects.
Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
Thus, there was no need additionally to control for specific phases or measures of the lockdown.
Predictors were modeled as fixed effects.
They included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
All predictors were included simultaneously and in each of the three models.

The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
To avoid overfitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
Finally, REWB-models cannot model scales as latent variables.
To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.

For more information on the analyses, for example a complete documentation of the models and their results, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).

## Measures

In what follows, I list all the variables that I analyzed.
For the variables' means, range, and variance, see Table \@ref(tab:descriptives).
For a complete list of all items and item characteristics, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).

### Well-being

Life satisfaction was measured with the item "Taken everything together, how satisfied are you currently with your life?".
The response options ranged from 0 (_extremely unsatisfied_) to 10 (_extremely satisfied_).

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
Reliability was also very high, $\omega$ = `r rel_aff_pos %>% my_round("std")`.

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
Reliability was also high, $\omega$ = `r rel_aff_neg %>% my_round("std")`.

All three variables were measured on each wave.

### COVID-19 related social media use

COVID-19 related social media use focused on communication _types_ was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
The three items were "Reading the posts of others with content on the Coronavirus", "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'", "I myself wrote posts on the Coronavirus on social media."
Answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
The items were inverted for the analyses.

COVID-19 related social media use focused on _channels_ was measured with five variables.
The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
Again, the answer options were 1 (_several times per day_), 2 (_daily_), 3 (_several times per week_), 4 (_weekly_), 5 (_never_).
Again, the items were inverted for the analyses.

Social media use was measured for all participants on waves 1, 2, 8, 17, and 23. 
Freshly recruited respondents always answered all questions on social media use.

```{r descriptives, cache=F}
apa_table(tab_desc_dat,
          stub_indents = list(
            "Well-being" = c(1:3),
            "Social media use" = c(4:6),
            "Social media channel" = c(7:11)),
          caption = "Descriptives of the main variables."
          )
```

### Control variables

The effects of COVID-19 related social media use were controlled for the following stable variables: 
(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
I originally planned to implement other variables as varying covariates.
However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.

# Results
## Preregistered Analyses

When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.

```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic. Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F}
knitr::include_graphics("figures/fig_descriptives.pdf")
```

The study's main hypothesis was that the effects of social media use on well-being would be trivial.
Regarding the effects of different communication _types_---that is, reading vs. sharing vs. posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:res-activity)).
For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
All confidence intervals include zero; hence, all effects were also statistically non-significant.
<!-- Only one effect did not include zero. -->
<!-- If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]). -->
<!-- However, because the effect did not fall outside of our null region, it's likely not large enough to be considered meaningful. -->
As a result, the hypothesis was supported for all COVID-19 related types of social media communication.

```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being. The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables. The SESOI was |0.30| for life satisfaction and |0.15| for affect. Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
knitr::include_graphics("figures/fig_results_activity.pdf")
```

Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
<!-- First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]). -->
<!-- However, note that this effect still fell completely within the null-region. -->
<!-- Hence, although positive the effect was considered too small to matter practically. -->
Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
Interestingly, Respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
However, note that the effect were still completely inside of the null region, hence not considered large enough to be practically relevant.

Note that when comparing the results with and without control variables, the results differed.
For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).

```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being. The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables. The SESOI was |0.30| for life satisfaction and |0.15| for affect. Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
knitr::include_graphics("figures/fig_results_channel.pdf")
```

Regarding the COVID-19 related use of social media _channels_, the results were comparable (see Figure \@ref(fig:res-channels)).
Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
Only two effects differed substantially from zero.
Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.

In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
Only one relation did not include zero, was hence statistically significant. 
Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
However, please note that this effect again was not considered large enough to be practically relevant.

Again, note that when comparing the results with and without control variables, the results differed.
Especially on the between-person level, five effects stopped being significant if they were controlled for additional variables.
For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).

```{r tab-aff-neg}
# apa_print(model_aff_neg_lmer)
# tab <- broom.mixed::tidy(model_aff_neg_lmer, conf.int = T) %>% 
#  select(term, estimate, conf.low, conf.high, p.value)
#  kable()
```

## Exploratory Analyses

In what follows, I briefly report some exploratory analyses that I did not preregister.
First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected covariates.
Because each variable had different response options, we would need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
Therefore, I report the results of the standardized scales,
which allows for a better comparison across variables.
In addition, as a rough estimate for the SESOI, we can use on the typical convention that small effects start at _r_ = |.10|.
The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
This includes for example internal locus of control, health, satisfaction with democracy, or excercising.
For an overview, see Figure \@ref(fig:res-control).

```{r res-control, fig.cap="Results of selected covariates. All variables were standardize except 'Male'  and ‘Employed in public service' because there were binary.", out.width = "\\textwidth", cache=F}
knitr::include_graphics("figures/fig_results_control_std.pdf")
```

To find out whether my inferences were robust across legitimate alternative analyses, I reran the analyses also using standardized estimates,  mean scores instead of factor scores, and with a data set where missing data were not imputed.
The results were virtually the same.
For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of $\beta$ = |.10|.
The additional analyses are reported in the [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).

# Discussion

In this study I analyzed the effects of COVID-19 related social media use on well-being.
The data come from a panel study with 24 waves and are representative of the Austrian population.
In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates.
The results showed that within-person effects were trivial.
People who used social media more than usual to learn about COVID-19 did not show meaningful changes in their well-being.

The results imply that COVID-19 related social media use does not seem to be particularly relevant for well-being.
Other factors among the third variables that were measured revealed larger effects or relations, showing that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or excercising.
According to this study, popular fears that "doomscrolling" or overusing social media during times of crises don't seem to be justified.

On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle. 
These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.

If anything, two preliminary and subtle trends can be observed.
First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
On the other hand, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so the results are ambivalent.
Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
Instagram, on the other hand, was related to slightly increased simultaneous levels of life satisfaction.
The positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
That said, all these effects were still very small and arguably too small to matter. 
But future research might elaborate on these specific relations to probe their stability and relevance.

Another last interesting observation is that life satisfaction was remarkably stable.
Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019].

## Limitations

The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
Together, this allows for a good perspective on potential causality.
That said, causality necessitates temporal order, and the cause needs to precede the effect.
Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
However, even then we don't know for certain if we actually measured the right interval.
Effects depend on the intensity of use or the length of the interval, and to borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.

Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
Media use is only one aspect of several factors that simultaneously affect well-being.
Is it realistic to expect that extremely changing only _one_ of these aspects should already manifest in a detectable change in well-being?
Or would it make more sense to expect that thoroughly committing to _two_ activities (e.g. regularly exercising _and_ establishing a reading habit) should then cause detectable improvements in well-being?
<!-- In other words, the beneficial effect of a particular activity is large enough if people feel a difference when implementing two of those activities.  -->
Practically, this would imply a SESOI half as large as I have defined here, that is _b_ = |.15| for well-being and _b_ = |.075| for affect.
In the case of this study, however, reducing the SESOI would not even make a difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
However, at all events future research needs to start a thorough conversation on what effect sizes are considered meaningful and what not.
With this study I hope to provide some first input and guidelines.

Both media use and well-being were measured using self-reports.
Measuring well-being with self-reports is adequate, because it by definition requires introspection.
However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
Unfortunately, such procedures introduce other problems, especially related to privacy.
Hence, for this type of research question it still seems necessary to use self-reported measures.

Because the data were collected in a single country, the generalizability of the results is limited.
The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
That said, because this is a comparatively large study representative of an entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.

<!-- Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one. -->
<!-- If treated as an ordinal one, it would have been necessary to analyze four different contrasts for each media measure, which plus the differentiation between between and within factor would have produced eight different measures, we would have made the model exceedingly complex. -->

## Conclusion

In this study, COVID-19 related social media use did not meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
If it's the aim to improve well-being, it might hence be fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.

\newpage

# References

<!-- \begingroup -->
<!-- \setlength{\parindent}{-0.5in} -->
<!-- \setlength{\leftskip}{0.5in} -->
<div id = "refs"></div>
<!-- \endgroup -->

\newpage

# Competing Interests

I declare no competing interests.

# Supplementary Material

All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).

# Data Accessibility Statement

The data are shared on AUSSDA, see https://doi.org/10.11587/28KQNS.
The data can only be used for scientific purposes.

### 2021-11-30 manuscript.tex
- Implemented several changes after ICA - Imputed missing data

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,mask,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={No meaningful effects of COVID-19 related social media use on well-being},
  pdflang={en-EN},
  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\setlength{\parskip}{0em}
\raggedbottom
\note{\clearpage}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{No meaningful effects of COVID-19 related social media use on well-being}
\author{Tobias Dienlin\textsuperscript{1}}
\date{}

\shorttitle{COVID-19 related social media use and well-being}

\authornote{

Tobias Dienlin, Department of Communication, University of Vienna.

Correspondence concerning this article should be addressed to Tobias Dienlin, University of Vienna, Department of Communication, 1090 Vienna, Austria. E-mail: \href{mailto:tobias.dienlin@univie.ac.at}{\nolinkurl{tobias.dienlin@univie.ac.at}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Vienna}

\abstract{
In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures. To this end, many people use various types of media, and increasingly social media. However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related reasons might put personal well-being at risk. To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants. I ran three random effects cross lagged panel models, controlling for several stable and varying covariates. Results showed that the effects of various types of COVID-19 related social media use on well-being were very small, arguably too small to matter. The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded.
}

\begin{document}
\maketitle

During the COVID-19 pandemic,
numerous events unfolded in quick succession.
Several open questions emerged.
How dangerous is the virus?
Is it spreading in my region?
How is it transmitted, and how can I protect myself?
Because for many it was (and continues to be) a matter of life or death, people aimed to stay informed regarding the latest developments.
Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).

Some people could not stop using social media to learn about COVID-19 related news.
A new phenomenon termed ``doomscrolling'' emerged:
Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
These concerns seem justified:
A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p. 283), which implies negative consequences for well-being.

As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
The panel consists of 24 waves and has an overall sample size of 3018.
The panel study collected a large number of psychological and demographic variables.
I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.

\hypertarget{defining-well-being-and-media-use}{%
\subsection{Defining Well-being and Media Use}\label{defining-well-being-and-media-use}}

The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).
According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
One of the most prominent indicators of well-being is life satisfaction.
In my view, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
Together, this should provide an encompassing perspective on potential media effects.

The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
Whereas the first four levels focus on the \emph{channel}, the last two address the \emph{type} of communication.
To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.

First, I investigate how different types of communication affect well-being.
Specifically, I differentiate between active and passive use.
I distinguish (a) \emph{reading} (passive), \emph{posting} (active), and \emph{liking and sharing} COVID-19 related posts (both active and passive).
Second, I analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
Branded apps are separate entities with potentially different effects.
Twitter might have a different effect as compared to WhatsApp because of their different affordances.
For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.
Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.

\hypertarget{effects-of-social-media-on-well-being}{%
\subsection{Effects of Social Media on Well-Being}\label{effects-of-social-media-on-well-being}}

So far, there is only little empirical research on COVID-19 related social media use on well-being.
In their study on the relations between media use and mental health during the pandemic, Bendau et al. (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
Together, the literature emphasizes potentially negative effects of social media as news use (see also Liu \& Tong, 2020; Riehm et al., 2020).
However, note that all of these findings represent between-person relation stemming from cross-sectional data (see below).
We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.

The question of whether and how \emph{general} social media use affect well-being, on the other hand, is well-researched.
This also holds true for the different types of communication such as active or passive use.
A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
But what determines whether or not an effect is considered small or trivial?
If we refer to standardized effect sizes, according to Cohen (1992) small effect sizes start at \emph{r} = .10.
And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).

These overviews are well aligned with several individual new studies employing advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
In conclusion, most effects are likely somewhere between trivial and small.
I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.

From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.

First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
If one learns about such events, the initial reaction might be shock, fear, or dismay.
Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life.
That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.

There could also be \emph{indirect} effects.
When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
For example, during the pandemic social media use was at an all-time high in the US (Statista, 2021).
In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.

On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might be not the worst idea.
Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.

Together, the strongest argument seems to be that \emph{in general} the effects of social media on well-being are, on average, small at best.
Because this study only looks at \emph{one part} of social media use---namely, COVID-19 related interactions---it is very focused, and the overall potential of the effects should diminish further.
Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
Therefore, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.

\begin{quote}
Hypothesis: The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators---while controlling for several stable and varying covariates such as sociodemographic variables and psychological dispositions---will be trivial.
\end{quote}

\hypertarget{current-study}{%
\section{Current Study}\label{current-study}}

\hypertarget{smallest-effect-size-of-interest}{%
\subsection{Smallest Effect Size of Interest}\label{smallest-effect-size-of-interest}}

Testing this hypothesis, however, is not trivial.
First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
Above I already referred to standardized effect sizes.
However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
Standardized effect sizes are determined by a sample's variance.\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992). Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.}, which is problematic:
The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
Instead, it should depend on absolute criteria.

What could be a minimally interesting, nontrivial effect?
Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
However, it is still necessary and helpful to try provide such a plausible benchmark.
I therefore suggest the following SESOI for this research question:

\begin{quote}
SESOI: If a heavy user of COVID-19 related social media news suddenly \emph{stops} using social media altogether, this should have a \emph{noticeable} impact on their overall well-being.
\end{quote}

What does this mean practically and how can it be operationalized?
In this study, COVID-19 related social media use was measured on a 5-point scale, ranging from 1 = \emph{never} to 5 = \emph{several times a day}.
Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
But what's a noticeable change in well-being?
According to Norman, Sloan, and Wyrwich (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.

\hypertarget{causality}{%
\subsection{Causality}\label{causality}}

The hypothesis explicitly states a causal effect.
In non-experimental studies, longitudinal designs help investigate causality.
Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
In addition, we for example also need to control for confounding third variables, and importantly also for \emph{varying} third variables.

To explain, consider the following example.
Imagine that a person suddenly starts using social media much more than usual, and then after some time becomes less satisfied with their life.
Eventually, use and life satisfaction return to prior levels.
If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.

Thus, only if we can control for \emph{all} confounders, can we correctly estimate causality without bias (Rohrer, 2018).
Obviously, we can never be entirely sure that we have included all confounders, which makes absolute statements regarding causality virtually impossible.
In addition, when determining the overall causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
Complicating matters, often it is unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
The aim should therefore be to collect as many varying and nonvarying confounders as possible, while knowing that absolute certainty regarding causality cannot be reached.

When searching for suitable candidates for confounders, we look for variables that affect both media use and well-being.
Controlling for these factors isolates the actual effect of social media use on well-being.
We can also control for variables that affect only social media use or well-being.
However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).

In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are very likely to do so, and which also are not mediators:
gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
I will not control for variables such as trust in institutions or trust in media, because these variables are likely influenced by social media use to a meaningful extent.

Next to including covariates, it is now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
In this study, I hence differentiate between-person relations from within-person effects.
And as explicated above, to test the hypothesis I thus consider only the within-person effects.

Finally, one precondition of causality is temporal order.
The cause needs to precede the effect.
Finding the right interval between cause and effect is crucial.
For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are necessary.
In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).

In this study, I therefore adopt an intermediate perspective.
I analyze if when a person changes their social media diet, are there simultaneous changes in their well-being?
When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.

\hypertarget{method}{%
\section{Method}\label{method}}

In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.

\hypertarget{preregistration}{%
\subsection{Preregistration}\label{preregistration}}

The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) I preregistered on the Open Science Framework.
The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
In some cases I could not execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website} (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

The data come from the Austrian Corona Panel Project (Kittel et al., 2021).
The study was conducted between March 2020 and October 2021.
It contains 26 waves, and at the time of writing the first 24 waves were available for download.
Each wave consists of at least 1,500 respondents.
The overall sample size was \emph{N} = 3,018, and 72,432 observations were collected.
Panel mortality was compensated through a continuous acquisition of new participants.
All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.

The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
The study received IRB approval from the University of Vienna.
The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.

\hypertarget{inference-criteria}{%
\subsection{Inference criteria}\label{inference-criteria}}

Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
The sample is very large, and it is hence well-equipped reliably to detect also small effects.
In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
To this end, I adopted the interval testing approach as proposed by Dienes (2014).
On the basis of the SESOI, I then defined a null region (in Bayesian terms, this is called the region of practical equivalence, ROPE).
In what follows, I explain how I determined the SESOI and the null region.

In this study, life satisfaction was measured on an 11-point scale.
If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
In a statistical regression, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
We would therefore expect a SESOI of \emph{b} = 1.57 / 4 = 0.39.
For affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
Because we are agnostic as to whether the effects are positive or negative, the null region will include negative and positive effects.
Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).}
Together, this leads to a null region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.

To illustrate, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is only trivial is supported.
If the effects interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
If the confidence falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell, Fairbrother, \& Jones, 2019).
Three models were run, one for each dependent variable.
The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
Nesting in participants allowed to separate between-person relations from within-person effects.
Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
Thus, there was no need additionally to control for specific phases or measures of the lockdown.
Predictors were modeled as fixed effects.
They included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
All predictors were included simultaneously and in each of the three models.

The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
Because Mardia's test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
Finally, REWB-models cannot model scales as latent variables.
To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.

For more information on the analyses, for example a complete documentation of the models and their results, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.

\hypertarget{measures}{%
\subsection{Measures}\label{measures}}

In what follows, I list all the variables that I analyzed.
For the variables' means, range, and variance, see Table \ref{tab:descriptives}.
For a complete list of all items and item characteristics, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.

\hypertarget{well-being}{%
\subsubsection{Well-being}\label{well-being}}

Life satisfaction was measured with the item ``Taken everything together, how satisfied are you currently with your life?''
The response options ranged from 0 (\emph{extremely unsatisfied}) to 10 (\emph{extremely satisfied}).

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \textit{p} = .032, cfi = 1.00, rmsea = .02, 90\% CI {[}.01, .03{]}, srmr = .01.
Reliability was also very high, \(\omega\) = .85.

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \textit{p} \textless{} .001, cfi = .97, rmsea = .08, 90\% CI {[}.07, .08{]}, srmr = .03.
Reliability was also high, \(\omega\) = .89.

All three variables were measured on each wave.

\hypertarget{covid-19-related-social-media-use}{%
\subsubsection{COVID-19 related social media use}\label{covid-19-related-social-media-use}}

COVID-19 related social media use focused on communication \emph{types} was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
The three items were ``Reading the posts of others with content on the Coronavirus,'' ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet','' ``I myself wrote posts on the Coronavirus on social media.''
Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
The items were inverted for the analyses.

COVID-19 related social media use focused on \emph{channels} was measured with five variables.
The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis on the following social media?''
The five items were (a) Facebook, (b) Twitter, (c) Instagram, (d) Youtube, and (e) WhatsApp.
Again, the answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
Again, the items were inverted for the analyses.

Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
Freshly recruited respondents always answered all questions on social media use.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:descriptives}Descriptives of the main variables.}

\begin{tabular}{lllll}
\toprule
 & \multicolumn{1}{c}{sd} & \multicolumn{1}{c}{min} & \multicolumn{1}{c}{max} & \multicolumn{1}{c}{mean}\\
\midrule
Well-being &  &  &  & \\
\ \ \ Life satisfaction & 1.68 & 6.38 & 6.81 & 6.59\\
\ \ \ Positive affect & 0.57 & 3.05 & 3.28 & 3.15\\
\ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
Social media use &  &  &  & \\
\ \ \ Read & 1.03 & 2.10 & 2.92 & 2.43\\
\ \ \ Like \& share & 0.87 & 1.61 & 1.92 & 1.78\\
\ \ \ Posting & 0.63 & 1.33 & 1.46 & 1.39\\
Social media channel &  &  &  & \\
\ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.37\\
\ \ \ Instagram & 0.83 & 1.84 & 2.66 & 2.09\\
\ \ \ WhatsApp & 1.23 & 2.28 & 2.62 & 2.45\\
\ \ \ YouTube & 0.88 & 1.77 & 2.30 & 2.00\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{control-variables}{%
\subsubsection{Control variables}\label{control-variables}}

The effects of COVID-19 related social media use were controlled for the following stable variables:
(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
I originally planned to implement other variables as varying covariates.
However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{preregistered-analyses}{%
\subsection{Preregistered Analyses}\label{preregistered-analyses}}

When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.

\begin{figure}
\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic. Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-desc}
\end{figure}

The study's main hypothesis was that the effects of social media use on well-being would be trivial.
Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
All confidence intervals include zero; hence, all effects were also statistically non-significant.
As a result, the hypothesis was supported for all COVID-19 related types of social media communication.

\begin{figure}
\includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being. The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables. The SESOI was |0.30| for life satisfaction and |0.15| for affect. Hence, all of the reported effects are considered not meaningful.}\label{fig:res-activity}
\end{figure}

Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
Interestingly, Respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
However, note that the effect were still completely inside of the null region, hence not considered large enough to be practically relevant.

Note that when comparing the results with and without control variables, the results differed.
For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).

\begin{figure}
\includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being. The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables. The SESOI was |0.30| for life satisfaction and |0.15| for affect. Hence, all of the reported effects are considered not meaningful.}\label{fig:res-channels}
\end{figure}

Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
Only two effects differed substantially from zero.
Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} simultaneous levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} simultaneous levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.

In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
Only one relation did not include zero, was hence statistically significant.
Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
However, please note that this effect again was not considered large enough to be practically relevant.

Again, note that when comparing the results with and without control variables, the results differed.
Especially on the between-person level, five effects stopped being significant if they were controlled for additional variables.
For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).

\hypertarget{exploratory-analyses}{%
\subsection{Exploratory Analyses}\label{exploratory-analyses}}

In what follows, I briefly report some exploratory analyses that I did not preregister.
First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected covariates.
Because each variable had different response options, we would need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
Therefore, I report the results of the standardized scales,
which allows for a better comparison across variables.
In addition, as a rough estimate for the SESOI, we can use on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
This includes for example internal locus of control, health, satisfaction with democracy, or excercising.
For an overview, see Figure \ref{fig:res-control}.

\begin{figure}
\includegraphics[width=\textwidth]{figures/fig_results_control_std} \caption{Results of selected covariates. All variables were standardize except 'Male'  and ‘Employed in public service' because there were binary.}\label{fig:res-control}
\end{figure}

To find out whether my inferences were robust across legitimate alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
The results were virtually the same.
For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
The additional analyses are reported in the \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In this study I analyzed the effects of COVID-19 related social media use on well-being.
The data come from a panel study with 24 waves and are representative of the Austrian population.
In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates.
The results showed that within-person effects were trivial.
People who used social media more than usual to learn about COVID-19 did not show meaningful changes in their well-being.

The results imply that COVID-19 related social media use does not seem to be particularly relevant for well-being.
Other factors among the third variables that were measured revealed larger effects or relations, showing that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or excercising.
According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.

On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being.
However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.

If anything, two preliminary and subtle trends can be observed.
First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
On the other hand, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so the results are ambivalent.
Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo, Baumgartner, Peter, \& Valkenburg, 2018), which might help explain the results.
Instagram, on the other hand, was related to slightly increased simultaneous levels of life satisfaction.
The positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
That said, all these effects were still very small and arguably too small to matter.
But future research might elaborate on these specific relations to probe their stability and relevance.

Another last interesting observation is that life satisfaction was remarkably stable.
Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
Together, this allows for a good perspective on potential causality.
That said, causality necessitates temporal order, and the cause needs to precede the effect.
Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
However, even then we don't know for certain if we actually measured the right interval.
Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.

Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
Media use is only one aspect of several factors that simultaneously affect well-being.
Is it realistic to expect that extremely changing only \emph{one} of these aspects should already manifest in a detectable change in well-being?
Or would it make more sense to expect that thoroughly committing to \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause detectable improvements in well-being?
Practically, this would imply a SESOI half as large as I have defined here, that is \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
In the case of this study, however, reducing the SESOI would not even make a difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
However, at all events future research needs to start a thorough conversation on what effect sizes are considered meaningful and what not.
With this study I hope to provide some first input and guidelines.

Both media use and well-being were measured using self-reports.
Measuring well-being with self-reports is adequate, because it by definition requires introspection.
However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
Unfortunately, such procedures introduce other problems, especially related to privacy.
Hence, for this type of research question it still seems necessary to use self-reported measures.

Because the data were collected in a single country, the generalizability of the results is limited.
The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
That said, because this is a comparatively large study representative of an entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

In this study, COVID-19 related social media use did not meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
If it's the aim to improve well-being, it might hence be fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}%
Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}

\leavevmode\hypertarget{ref-bellFixedRandomEffects2019}{}%
Bell, A., Fairbrother, M., \& Jones, K. (2019). Fixed and random effects models: Making an informed choice. \emph{Quality \& Quantity}, \emph{53}(2), 1051--1074. \url{https://doi.org/10.1007/s11135-018-0802-x}

\leavevmode\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}%
Bendau, A., Petzold, M. B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J. (2021). Associations between {COVID}-19 related media consumption and symptoms of anxiety, depression and {COVID}-19 related fear in the general population in {Germany}. \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291. \url{https://doi.org/10.1007/s00406-020-01171-6}

\leavevmode\hypertarget{ref-beyensSocialMediaUse2021}{}%
Beyens, I., Pouwels, J. L., Driel, I. I. van, Keijsers, L., \& Valkenburg, P. M. (2021). Social media use and adolescents' well-being: {Developing} a typology of person-specific effect patterns. \emph{Communication Research}. \url{https://doi.org/10.31234/osf.io/ftygp}

\leavevmode\hypertarget{ref-brownEasyHappinessPie2019}{}%
Brown, N. J. L., \& Rohrer, J. M. (2019). Easy as (happiness) pie? {A} critical evaluation of a popular model of the determinants of well-being. \emph{Journal of Happiness Studies}. \url{https://doi.org/10.1007/s10902-019-00128-4}

\leavevmode\hypertarget{ref-choiMediatedCommunicationMatters2021}{}%
Choi, M., \& Choung, H. (2021). Mediated communication matters during the {COVID}-19 pandemic: {The} use of interpersonal and masspersonal media and psychological well-being. \emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418. \url{https://doi.org/10.1177/02654075211029378}

\leavevmode\hypertarget{ref-cohenPowerPrimer1992}{}%
Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}

\leavevmode\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}%
Diener, E., Lucas, R. E., \& Oishi, S. (2018). Advances and open questions in the science of subjective well-being. \emph{Collabra: Psychology}, \emph{4}(1), 15. \url{https://doi.org/10.1525/collabra.115}

\leavevmode\hypertarget{ref-dienesUsingBayesGet2014}{}%
Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}

\leavevmode\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}%
Dienlin, T., \& Johannes, N. (2020). The impact of digital technology use on adolescent well-being. \emph{Dialogues in Clinical Neuroscience}, \emph{22}(2), 135--142. \url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}

\leavevmode\hypertarget{ref-edenMediaCopingCOVID192020}{}%
Eden, A. L., Johnson, B. K., Reinecke, L., \& Grady, S. M. (2020). Media for coping during {COVID}-19 social distancing: {Stress}, anxiety, and psychological well-being. \emph{Frontiers in Psychology}, \emph{11}, 577639. \url{https://doi.org/10.3389/fpsyg.2020.577639}

\leavevmode\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}%
Ferguson, C. J., Kaye, L. K., Branley-Bell, D., Markey, P., Ivory, J. D., Klisanin, D., \ldots{} Wilson, J. (2021). Like this meta-analysis: {Screen} media and mental health. \emph{Professional Psychology: Research and Practice}. \url{https://doi.org/10.1037/pro0000426}

\leavevmode\hypertarget{ref-funderEvaluatingEffectSize2019}{}%
Funder, D. C., \& Ozer, D. J. (2019). Evaluating effect size in psychological research: {Sense} and nonsense. \emph{Advances in Methods and Practices in Psychological Science}, \emph{2}(2), 156--168. \url{https://doi.org/10.1177/2515245919847202}

\leavevmode\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}%
Greenspoon, P. J., \& Saklofske, D. H. (2001). Toward an integration of subjective well-being and psychopathology. \emph{Social Indicators Research}, \emph{54}(1), 81--108. \url{https://doi.org/10.1023/A:1007219227883}

\leavevmode\hypertarget{ref-hamakerWhyResearchersShould2014}{}%
Hamaker, E. L. (2014). Why researchers should think "within-person": {A} paradigmatic rationale. In M. R. Mehl, T. S. Conner, \& M. Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.). New York, NY: Guilford.

\leavevmode\hypertarget{ref-huangTimeSpentSocial2017}{}%
Huang, C. (2017). Time spent on social network sites and psychological well-being: {A} meta-analysis. \emph{Cyberpsychology, Behavior and Social Networking}, \emph{20}(6), 346--354. \url{https://doi.org/10.1089/cyber.2016.0758}

\leavevmode\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}%
Keresteš, G., \& Štulhofer, A. (2020). Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach. \emph{Computers in Human Behavior}, \emph{104}, 106187. \url{https://doi.org/10.1016/j.chb.2019.106187}

\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2020}{}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L. (2020). \emph{Austrian {Corona} {Panel} {Project} ({SUF} edition)}. AUSSDA. \url{https://doi.org/10.11587/28KQNS}

\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2021}{}%
Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L. (2021). The {Austrian} {Corona} {Panel} {Project}: Monitoring individual and societal dynamics amidst the {COVID}-19 crisis. \emph{European Political Science}, \emph{20}(2), 318--344. \url{https://doi.org/10.1057/s41304-020-00294-7}

\leavevmode\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}%
Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. Retrieved from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}

\leavevmode\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}%
Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). New York, NY: The Guilford Press.

\leavevmode\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}%
Liu, J. C. J., \& Tong, E. M. W. (2020). The relation between official {WhatsApp}-distributed {COVID}-19 news exposure and psychological symptoms: Cross-sectional survey study. \emph{Journal of Medical Internet Research}, \emph{22}(9), e22142. \url{https://doi.org/10.2196/22142}

\leavevmode\hypertarget{ref-mcelreathYesterdayClass2021}{}%
McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}. Retrieved from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}

\leavevmode\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}%
Meier, A., \& Reinecke, L. (2020). Computer-mediated communication, social media, and mental health: {A} conceptual and empirical meta-review. \emph{Communication Research}, 009365022095822. \url{https://doi.org/10.1177/0093650220958224}

\leavevmode\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}%
Norman, G., Sloan, J., \& Wyrwich, K. (2003). Interpretation of changes in health-related quality of life: {The} remarkable universality of half a standard deviation. \emph{Medical Care}, \emph{41}(5), 582--592. Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}

\leavevmode\hypertarget{ref-orbenTeenagersScreensSocial2020}{}%
Orben, A. (2020). Teenagers, screens and social media: A narrative review of reviews and key studies. \emph{Social Psychiatry and Psychiatric Epidemiology}, \emph{55}(4), 407--414. \url{https://doi.org/10.1007/s00127-019-01825-4}

\leavevmode\hypertarget{ref-orbenSocialMediaEnduring2019}{}%
Orben, A., Dienlin, T., \& Przybylski, A. K. (2019). Social media's enduring effect on adolescent life satisfaction. \emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228. \url{https://doi.org/10.1073/pnas.1902058116}

\leavevmode\hypertarget{ref-przybylskiDoesTakingShort2021a}{}%
Przybylski, A. K., Nguyen, T. T., Law, W., \& Weinstein, N. (2021). Does taking a short break from social media have a positive effect on well-being? {Evidence} from three preregistered field experiments. \emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514. \url{https://doi.org/10.1007/s41347-020-00189-w}

\leavevmode\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}%
Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}

\leavevmode\hypertarget{ref-riehmAssociationsMediaExposure2020}{}%
Riehm, K. E., Holingue, C., Kalb, L. G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J. (2020). Associations between media exposure and mental distress among {U}.{S}. Adults at the beginning of the {COVID}-19 pandemic. \emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638. \url{https://doi.org/10.1016/j.amepre.2020.06.008}

\leavevmode\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}%
Rohrer, J. M. (2018). Thinking clearly about correlations and causation: {Graphical} causal models for observational data. \emph{Advances in Methods and Practices in Psychological Science}, \emph{24}(2), 251524591774562. \url{https://doi.org/10.1177/2515245917745629}

\leavevmode\hypertarget{ref-rohrerTheseAreNot2021}{}%
Rohrer, J. M., \& Murayama, K. (2021). \emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} {[}Preprint{]}. PsyArXiv. \url{https://doi.org/10.31234/osf.io/tg4vj}

\leavevmode\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}%
Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S. (2021). Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place. Retrieved from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}

\leavevmode\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}%
Scharkow, M. (2016). The accuracy of self-reported {Internet} use---{A} validation study using client log data. \emph{Communication Methods and Measures}, \emph{10}(1), 13--27. \url{https://doi.org/10.1080/19312458.2015.1118446}

\leavevmode\hypertarget{ref-schemerImpactInternetSocial2021}{}%
Schemer, C., Masur, P. K., Geiß, S., Müller, P., \& Schäfer, S. (2021). The {Impact} of {Internet} and {Social} {Media} {Use} on {Well}-{Being}: {A} {Longitudinal} {Analysis} of {Adolescents} {Across} {Nine} {Years}. \emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21. \url{https://doi.org/10.1093/jcmc/zmaa014}

\leavevmode\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}%
Schnauber-Stockmann, A., \& Karnowski, V. (2020). Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs. \emph{Communication Methods and Measures}, \emph{14}(3), 145--164. \url{https://doi.org/10.1080/19312458.2020.1784402}

\leavevmode\hypertarget{ref-stainbackCOVID1924News2020}{}%
Stainback, K., Hearne, B. N., \& Trieu, M. M. (2020). {COVID}-19 and the 24/7 {News} {Cycle}: {Does} {COVID}-19 {News} {Exposure} {Affect} {Mental} {Health}? \emph{Socius}, \emph{6}, 2378023120969339. \url{https://doi.org/10.1177/2378023120969339}

\leavevmode\hypertarget{ref-statistaAverageDailyTime2021}{}%
Statista. (2021). \emph{Average daily time spent on social networks by users in the {United} {States} from 2018 to 2022}. Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}

\leavevmode\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}%
Waterloo, S. F., Baumgartner, S. E., Peter, J., \& Valkenburg, P. M. (2018). Norms of online expressions of emotion: {Comparing} {Facebook}, {Twitter}, {Instagram}, and {WhatsApp}. \emph{New Media \& Society}, \emph{20}(5), 1813--1831. \url{https://doi.org/10.1177/1461444817707349}

\end{CSLReferences}

\newpage

\hypertarget{competing-interests}{%
\section{Competing Interests}\label{competing-interests}}

I declare no competing interests.

\hypertarget{supplementary-material}{%
\section{Supplementary Material}\label{supplementary-material}}

All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).

\hypertarget{data-accessibility-statement}{%
\section{Data Accessibility Statement}\label{data-accessibility-statement}}

The data are shared on AUSSDA, see \url{https://doi.org/10.11587/28KQNS}.
The data can only be used for scientific purposes.

\end{document}

### 2021-11-30 preregistration.Rmd
- Implemented several changes after ICA - Imputed missing data

---
title: "Preregistration"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

The actual preregistration can be found [here](https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3). I have also pasted it below.

## Hypotheses

The within-person effects of all types of COVID-19 related social media use on all types of well-being indicators will be trivial, while controlling for several stable and varying covariates such as  sociodemographic variables and psychological dispositions.

## Design Plan
#### Study type

Observational Study - Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, “natural experiments,” and regression discontinuity designs.

#### Blinding

No blinding is involved in this study.

#### Is there any additional blinding in this study?

No response

## Study design

Secondary data analysis of a questionnaire-based panel study with 24 waves that consist of at least 1,500 participants per wave. A large number of questions was collected each, and a subset will be used for this study.

#### Randomization

No response

#### Sampling Plan

Existing Data

Registration prior to accessing the data

#### Explanation of existing data

I will analyze the Austrian Corona Panel. The data are posted on AUSSDA (https://doi.org/10.11587/28KQNS). I have not yet accessed the data. I have used the study description online to design the preregistration. After finishing the preregistration, I plan to create a login on AUSSDA. I will then download the scientific use file. In case the use file should not contain some of the variables I preselected, or if they are recoded such that the variables cannot be used, I will directly contact the study authors to ask for access to the original data set. In case this won't be possible, I will slightly change the analyses so that I can use the scientific use file.

#### Data collection procedures

I will use the already existing data set. Information on how the data was collected can be found here: https://viecer.univie.ac.at/coronapanel/austrian-corona-panel-data/method-report/

#### Sample size

The sample consists of at least 1,500 respondents per wave. Panel attrition was remedied by collecting new respondents. The overall and final sample size is not reported online, and will know it once I have downloaded the final data set.

#### Sample size rationale

The sample size was predetermined by the original study authors.

#### Stopping rule

Data collection is already finished.

## Variables
#### Manipulated variables

No variables were manipulated.

#### Measured variables

(Note: To improve readability, compared to the actual [preregistration](https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3) the following part was reformatted and partially translated. No actual changes were introduced.)

In what follows, I list all the variables that I will use for the analyses (which also means that they cannot be dropped later). Note that the data set contains many more variables that I will not use. I provide the variable name in English, and the response options, answer format, and items in German.

###### Life Satisfaction

"Taken everything together, how satisfied are you currently with your life?"

(0 = extremely unsatisfied, 10 = extremely satisfied)

###### Positive Affect (scale)

"In the last week, how often did you feel"

 - a: ... calm and relaxed
 - e: ... happy
 - i: ... full of energy

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

###### Negative Affect (scale)

"In the last week, how often did you feel"

 - b: ... lonely
 - c: ... aggravated
 - d: ... so depressed, that nothing could lift you up
 - f: ... very nervous
 - g: ... anxious
 - h: ... glum and sad

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

###### Social media use: Reading

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Postings anderer lesen, in denen es um Inhalte zum Coronavirus geht"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

###### Social media use: Liking & Sharing

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Bei Postings anderer zum Coronavirus auf den „Gefällt mir“ Button drücken, „teilen“ oder „retweeten“"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

###### Social media use: Active Postings

"Wie oft haben Sie letzte Woche die folgenden Aktivitäten in Sozialen Netzwerken gesetzt?"

 - "Selbst Postings zum Coronavirus in Sozialen Netzwerken verfasst"

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

###### Social media use: Specific platforms

"Wie oft in der letzten Woche haben Sie Informationen zur Corona-Krise in den folgenden Sozialen Medien verfolgt?"

 - Facebook 
 - Twitter
 - Instagram
 - Youtube
 - WhatsApp

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

###### Gender

"Welches Geschlecht haben Sie?"

(a: Männlich; b Weiblich; c: Divers)

###### Age

"In welchem Jahr wurden Sie geboren?"

(Jahr)

###### Education

"Was ist der höchste Schul- oder Bildungsabschluss, den Sie erreicht haben?"

 - a. Volksschule oder weniger
 - b. Hauptschule oder AHS Unterstufe
 - c. Polytechnikum, BMS (Fachschule, z.B. HASCH)
 - d. Lehre, Berufsschule
 - e. AHS mit Matura
 - f. BHS mit Matura (z.B. HTL, HAK, HBLA etc.) 
 - g. Hochschulverwandte Lehranstalt oder Kolleg 
 - h. Bachelor
 - i. Magister / Master / Diplomingenieur / Fachhochschule 
 - j. Doktor / PhD

###### Austria country of birth

"Ist Ihr Geburtsland Österreich?"

  - Ja
  - Nein

###### Austria country of birth of parents

 "Wurde einer oder beide Ihrer Elternteile nicht in Österreich geboren?"

 - a. Ja, einer
 - b. Ja, beide
 - c. Nein

###### News consumption: Text-based

"Nachstehend finden Sie eine Liste mit verschiedenen Tageszeitungen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in der jeweiligen Zeitung (im Internet oder gedruckt) informiert?"

 - Kronen Zeitung oder www.krone.at
 - Der Standard oder derstandard.at
 - Die Presse oder diepresse.com
 - Österreich oder oe24.at
 - Kurier oder kurier.at
 - Salzburger Nachrichten oder salzburg.at
 - Sonstige österreichische Tageszeitungen

###### News consumption: Video-based

"Nachstehend finden Sie eine Liste mit verschiedenen Nachrichtensendungen im Fernsehen. Wie oft haben Sie sich letzte Woche über das politische Geschehen in den folgenden Nachrichtensendungen informiert?"

 - "Nachrichtensendungen im ORF (z.B. Zeit im Bild um 19:30h oder ZiB 2 um 22:00h)"
 - "Nachrichtensendungen im Privatfernsehen (z.B. ATV aktuell, Puls 4/ Pro7/ Sat1 Austria)"

(1 = Mehrmals täglich; 2 = Einmal täglich; 3 = Mehrmals in der Woche; 4 = Einmal in der Woche; 5 = Gar nicht/nie)

###### Residency Vienna

"In welchem Bundesland leben Sie aktuell?"

 - a. Vorarlberg
 - b. Tirol
 - c. Salzburg
 - d. Steiermark
 - e. Kärnten
 - f. Oberösterreich
 - g. Niederösterreich
 - h. Wien
 - i. Burgenland

(Will be transformed to Vienna: yes/no)

###### Household size

"Bitte rechnen Sie hier sich selbst wieder mit ein. Wenn es eine Personengruppe nicht in Ihrem Haushalt gibt, tragen Sie bitte „0“ ein.

 - a. ...Erwachsene über 18 Jahre:
 - b. ...Teenager von 15-17 Jahren:
 - c. ...Kinder von 6-14 Jahren:
 - d. ...Kinder von 3-5 Jahren:
 - e. ...Kinder von 0-2 Jahren:

(Achtung, wenn Anzahl Erwachsene missing, 1 eintragen, da vorher Filterfrage)

###### Health

"Kommen wir jetzt zu ein paar Fragen zu Ihrer Gesundheit. Wie schätzen Sie Ihren allgemeinen Gesundheitszustand ein? Würden Sie sagen, er ist..."

 - a. 1 = Sehr gut
 - b. 2 = Gut
 - c. 3 = Mittelmäßig
 - d. 4 = Schlecht
 - e. 5 = Sehr schlecht

###### Living space

"Wie viele Quadratmeter hat Ihre Wohnung?"

(Einfach-Nennung, Zahleneingabe, min. 0 – max. 600, freiwillig)

###### Access garden

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?

 - (Ja, Garten)

###### Access balcony

"Haben Sie Zugang zu einer privaten Freifläche (Balkon, Garten)?"

 - (Ja, Balkon)

###### Employment

Wenn Sie sich selbst zuordnen: Im Februar 2020, welcher der folgenden Gruppen gehören Sie vorwiegend an? (Einfach-Nennung)

 - a. Unselbstständig erwerbstätig (auch Lehrling), Privatwirtschaft
 - b. Unselbstständig erwerbstätig (auch Lehrling), Öffentlicher Dienst
 - c. Selbstständig erwerbstätig
 - d. Arbeitslos
 - e. Pensionist*in
 - f. Ausschließlich haushaltsführend
 - g. Schüler/in, Student/in, Bildungskarenz
 - h. Dauerhaft arbeitsunfähig
 - i. Präsenz-/Zivildiener
 - j. In Elternkarenz

###### Work hours per week

"Wie viele Stunden (inklusive Überstunden) arbeiten Sie in Ihrer Haupttätigkeit jetzt pro Woche? Wenn Sie es nicht genau wissen, genügt Ihre beste Schätzung."

###### Home-Office Yes/No

"Hat sich in Ihrer beruflichen Situation aufgrund der Corona-Krise etwas geändert? Bitte wählen Sie alles Zutreffende aus."

 - a. Ich bin im Home-Office

###### Household income

"Wie viel Geld steht Ihrem Haushalt aktuell zur Verfügung (Netto-Haushaltseinkommen inklusive Sozialleistungen, Rente usw.)? Bitte rechnen Sie Überstunden mit ein, nicht aber den 13./14. Bezug."

 - a. weniger als 1.100 Euro
 - b. 1.100 bis unter 1.500 Euro 
 - c. 1.500 bis unter 1.800 Euro 
 - d. 1.800 bis unter 2.200 Euro 
 - e. 2.200 bis unter 2.700 Euro 
 - f. 2.700 bis unter 3.100 Euro 
 - g. 3.100 bis unter 3.700 Euro 
 - h. 3.700 bis unter 4.300 Euro 
 - i. 4.300 bis unter 5.500 Euro 
 - j. 5.500 Euro und mehr

###### Outdoor activities

"Haben Sie wegen folgenden Gründen in der vergangenen Woche Ihr Zuhause verlassen?"

 - a: Arbeit
 - b: Sport
 - c: Freunde oder Verwandte treffen
 - e: Essen einkaufen
 - f: Haustiere ausführen

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

###### Satisfaction Democracy

"Wie zufrieden oder unzufrieden sind Sie alles in allem mit der Demokratie, so wie sie derzeit in Österreich funktioniert?"

 - a. 1 = Sehr zufrieden
 - b. 2 = Eher zufrieden
 - c. 3 = Teils-teils
 - d. 4 = Eher unzufrieden 
 - e. 5 = Sehr unzufrieden

###### Risk-taking

"Wie sehr sind Sie im Allgemeinen dazu bereit, Risiken einzugehen?"

(0: gar nicht bereit; 10: sehr bereit)

###### Locus of Control

 - "Ich habe mein Leben selbst in der Hand."
 - "Wenn ich mich anstrenge, werde ich auch Erfolg haben."
 - "Egal ob privat oder im Beruf: Mein Leben wird zum großen Teil von anderen bestimmt."
 - "Meine Pläne werden oft vom Schicksal durchkreuzt."

(1 = Trifft voll und ganz zu; 2 = Trifft etwas zu; 3 = Trifft wenig zu; 4 = Trifft nicht zu)
#### Indices

All the measures above that include the label (scale) will be combined into a scale. These variables will be analyzed using CFAs. If, and only if, model fit is below the recommended thresholds, the models will be adapted to achieve sufficient fit (see below). To avoid overfitting, I will use more liberal model fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10). Of these final models I will export factor scores, which will be used for the subsequent analyses.

## Analysis Plan
#### Statistical models

The data will be analyzed using random effects within between models. The independent variables (all types of social media use) will be analyzed on a between person level (person average) and a within-person level (deviation from person average for each wave).

Because the models are highly complex it is likely that they will not converge immediately. I plan the following analysis pipeline.

Analysis (proceed if doesn't converge):
- REWB Frequentist Model: Random Slope for Social Media Use & Covars
	- REWB Frequentist Model: Random Slope for Social Media Use & Fixed Slope for Covars
		- REWB Frequentist Model: Fixed Slope for Social Media Use  & Fixed Slope for Covars
			- REWB Frequentist Model: Fixed Slope for Social Media Use & Fixed Slope for Covars with standardized effects
				- REWB Bayesian Model
					- REWB Bayesian Model with standardized effects

CFA:
The factorial validity of the measures and the hypotheses will be tested with structural equation modeling (SEM). If Mardia’s test shows that the assumption of multivariate normality is violated, I will use the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator. I will test each scale in a confirmatory factor analysis. To avoid overfitting, I will use more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) (Kline 2016). If model fit is below the criteria, I will first inspect modification indices, potentially allowing covariance or cross-loadings if theoretically plausible. If these changes do not yield sufficient fit, I will drop malfunctioning items. If fit is still subpar, I will conduct exploratory factor analyses (EFA) to assess the underlying factor structure. EFAs will be run using maximum likelihood estimation and oblimin rotation (Osborne and Costello 2004, 7). If more than one dimension will be revealed, I will implement bifactor model solutions. Bifactor models retain a general measure of the variable, and make it unnecessary to introduce novel (and potentially overfitted) subdimensions. If no adequate bifactor model can be found, I will proceed by deleting items with low loadings on the general factor and/or the specific factors. If also after deletion of individual items no bifactor solution should emerge, I will use a subset of the items to extract a single factor with sufficient factorial validity.

#### Transformations

See list of variables.

#### Inference criteria

Definition of trivial effect size (SESOI)

- b < .3 (life-satisfaction), b < .15 (positive and negative affect)
	- If I need to use standardized effects (see above): beta < .1
- in addition, p < 5%

#### Data exclusion

I will individually check responses for patterns such as straight-lining or missing of inverted items, making sure to remove only clear cases. I will exclude respondents if they answer less than 50% of all questions.

#### Missing data

Missing responses will be imputed using predictive mean matching.

#### Exploratory analysis

I will provide additional analyses online, such as the results without control variables. These aren't alternative results, but rather stem from procedures that I consider -- at this point -- to be inferior.

## Other

It's a deliberate decision not to statistically control the analyses for potential mediators, such as trust in politics, media, etc.

### 2021-11-30 preregistration_changes.Rmd
- Implemented several changes after ICA - Imputed missing data

---
title: "Preregistration: Changes"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

The following changes were introduced after the preregistration.

## Covariates

I originally planned to include the following variables as varying covariates. I thought they were measured at several/all waves. However, upon seeing the actual data I realized this wasn't the case. I hence included them as stable covariates.

This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.

The following variables I originally planned to include as control variables. However, their inclusion led to the deletion of too many cases/observations. I hence decided against their inclusion. This includes (a) household size, (b) work hours per week, (c) home office, (d) household income.

### 2021-11-30 analyses.RmD
- updated urls

@@
 
-Load packages, and make repo replicable.
+Load packages.
 
@@
 # install packages
-# devtools::install_github("https://github.com/tdienlin/td")
+# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
 
 # define packages
-packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
-              "gridExtra", "haven", "kableExtra", "knitr", 
-              "lavaan", "lme4", "magrittr", "mice", "mvnormalTest", "PerFit", 
-              "psych", "quanteda.textstats",
-              "pwr", "purrr", "semPlot", "semTools", "tidyverse")
-
-# renv::install(packages)
-# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
-# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install
+packages <- c("brms", "devtools", "GGally", "ggplot2", 
+              "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
+              "magrittr", "mice", "mvnormalTest", 
+              "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")
 

### 2021-11-30 manuscript.Rmd
- updated urls

@@
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
-                      cache=T,
-                      echo=F
-                      # , fig.pos = 'h'
+                      cache=T, echo=F
                       )
@@
 
-# make project reproducible
-# devtools::install_github("rstudio/renv")  # install renv
-# renv::init()  # initialize renv
+# install necessary packages
+# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")
+# devtools::install_github("https://github.com/crsh/papaja@devel")
 
-# install necessary packages
+# load packages
 packages <- c("devtools", "english", "ggplot2", "kableExtra", 
@@
               "magrittr", "semPlot", "tidyverse")
-# renv::install(packages)
-# renv::install("tdienlin/td@v.0.0.2.5")  # uncomment to install
-# renv::install("crsh/papaja@v0.1.0.9997")  # uncomment to install
-# renv::install("crsh/papaja@devel")  # uncomment to install
-
-# load packages
 lapply(c(packages, "papaja", "td"), library, character.only = TRUE)
-
-# save snapshot of installed packages
-# renv::snapshot()
 
@@
 load("data/workspace_2.RData")
-
-# export references
-# r_refs(file = "bibliography/r-references.bib")
 ```
@@
 During the COVID-19 pandemic, 
-<!-- and especially in the beginning,  -->
 numerous events unfolded in quick succession.
@@
 In some cases I could not execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html) (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html) (https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
 
@@
 
-For more information on the analyses, for example a complete documentation of the models and their results, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+For more information on the analyses, for example a complete documentation of the models and their results, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
 
@@
 For the variables' means, range, and variance, see Table \@ref(tab:descriptives).
-For a complete list of all items and item characteristics, see [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+For a complete list of all items and item characteristics, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
 
@@
 For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of $\beta$ = |.10|.
-The additional analyses are reported in the [companion website](https://xmtra.github.io/Austrian_Corona_Panel_Project/analyses_additional.html).
+The additional analyses are reported in the [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/analyses_additional.html).
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
 

### 2021-11-30 manuscript.tex
- updated urls

@@
 In some cases I could not execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website} (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website} (\url{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}).
 
@@
 
-For more information on the analyses, for example a complete documentation of the models and their results, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
+For more information on the analyses, for example a complete documentation of the models and their results, see \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website}.
 
@@
 For the variables' means, range, and variance, see Table \ref{tab:descriptives}.
-For a complete list of all items and item characteristics, see \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}{companion website}.
+For a complete list of all items and item characteristics, see \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website}.
 
@@
 For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
-The additional analyses are reported in the \href{https://xmtra.github.io/Austrian_Corona_Panel_Project/analyses_additional.html}{companion website}.
+The additional analyses are reported in the \href{https://tdienlin.github.io/Austrian_Corona_Panel/analyses_additional.html}{companion website}.
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://xmtra.github.io/Austrian_Corona_Panel_Project/index.html}).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}).
 

### 2021-12-01 manuscript.Rmd
- updated language

@@
 
-abstract          : "In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
+abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
@@
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that the effects of various types of COVID-19 related social media use on well-being were very small, arguably too small to matter.
+I ran three random effects within between models, controlling for several stable and varying confounders Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
@@
 How is it transmitted, and how can I protect myself?
-Because for many it was (and continues to be) a matter of life or death, people aimed to stay informed regarding the latest developments.
+Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
 Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
@@
 These concerns seem justified: 
-A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
-283], which implies negative consequences for well-being.
+A study with 6,233 people from Germany that was conducted during the pandemic found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
+283].
 
@@
 <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
-To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
+To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
 Together, this should provide an encompassing perspective on potential media effects.
@@
 The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
+First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
 Whereas the first four levels focus on the _channel_, the last two address the _type_ of communication.
@@
 I distinguish (a) _reading_ (passive), _posting_ (active), and _liking and sharing_ COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially different effects. 
-Twitter might have a different effect as compared to WhatsApp because of their different affordances.
+Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
+Branded apps are separate entities with potentially divergent effects. 
+Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
 For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
@@
 288).
-@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
+@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
 In a sample of 312 respondents collected via Amazon Mechanical Turk, @choiMediatedCommunicationMatters2021 reported that people who used media to attain information were more lonely and less satisfied with their lives.
@@
 Together, the literature emphasizes potentially negative effects of social media as news use [see also @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
-However, note that all of these findings represent between-person relation stemming from cross-sectional data (see below).
+However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
 We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
 
-The question of whether and how _general_ social media use affect well-being, on the other hand, is well-researched.
+The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched.
 This also holds true for the different types of communication such as active or passive use.
 A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small [@meierComputermediatedCommunicationSocial2020a]---potentially too small to matter.
-But what determines whether or not an effect is considered small or trivial?
-If we refer to standardized effect sizes, according to @cohenPowerPrimer1992 small effect sizes start at _r_ = .10.
+What determines whether or not an effect is considered small or trivial?
+As a starting point, we could refer to standardized effect sizes.
+According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10.
 And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
 
-These overviews are well aligned with several individual new studies employing advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
+These overviews are well aligned with several individual studies employing advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
 In conclusion, most effects are likely somewhere between trivial and small.
@@
 If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life.
+Repeatedly consuming such news can be depressing, perhaps even changing some general perspectives on life.
 <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
@@
 
-Together, the strongest argument seems to be that _in general_ the effects of social media on well-being are, on average, small at best.
-Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, and the overall potential of the effects should diminish further.
+Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.
+Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
 Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
+First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
 Effectively testing this hypothesis necessitates defining what's considered a "trivial effect size" and what's not.
@@
 However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
-Standardized effect sizes are determined by a sample's variance.[^cohensd], which is problematic: 
+Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic: 
 The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
@@
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction. 
+(For more information, see Method section "Inference criteria.")
 
@@
 
-To explain, consider the following example.
+To illustrate, consider the following example.
 Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life.
@@
 
-Thus, only if we can control for _all_ confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
-Obviously, we can never be entirely sure that we have included all confounders, which makes absolute statements regarding causality virtually impossible.
+Thus, only when controlling for _all_ relevant confounders, we can correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
+Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
 In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
-Complicating matters, often it is unclear if a variable is a mediator or a confounder.[^collider]
-However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible, while knowing that absolute certainty regarding causality cannot be reached.
+Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider]
+However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
+The aim should therefore be to collect as many varying and nonvarying confounders as possible (which it seems is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
 
@@
 
-When searching for suitable candidates for confounders, we look for variables that affect both media use and well-being. 
+When searching for suitable candidates for confounders, we should look out for variables that affect both media use and well-being. 
 Controlling for these factors isolates the actual effect of social media use on well-being.
@@
 
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are very likely to do so, and which also are not mediators:
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also are not mediators:
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
 gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables are likely influenced by social media use to a meaningful extent.
+I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
 
@@
 <!-- Finding the right interval is difficult. -->
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are necessary.
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
 In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
@@
 
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) I preregistered on the Open Science Framework. 
+The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework. 
 The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
@@
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is very large, and it is hence well-equipped reliably to detect also small effects.
+The sample is large, and it is hence well-equipped reliably to detect also small effects.
 <!-- , which is why no exact post hoc power analysis were conducted. -->
@@
 To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
-On the basis of the SESOI, I then defined a null region (in Bayesian terms, this is called the region of practical equivalence, ROPE).
+On the basis of the SESOI, I then defined a null region.
 In what follows, I explain how I determined the SESOI and the null region.
@@
 Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
-In a statistical regression, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
+In a statistical regression analysis, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
 We would therefore expect a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
 For affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
-Because we are agnostic as to whether the effects are positive or negative, the null region will include negative and positive effects.
+Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
 Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.[^rounding] 
@@
 <!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
-To illustrate, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is only trivial is supported.
+To illustrate what this means in practice, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is trivial is supported.
 If the effects interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
@@
 <!-- X clear cases were removed. -->
-<!-- Respondents who answered less than 50% of all questions were removed. -->
-<!-- The remaining missing responses were imputed using predictive mean matching. -->
 
@@
 Three models were run, one for each dependent variable.
-The data were hierarchical, and responses were separately nested in participants and waves (put differently, participants and waves were implemented as random effects).
+The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
 Nesting in participants allowed to separate between-person relations from within-person effects.
-Nesting in waves allowed to control for general exogenous developments, such as general decrease in well-being in the population, for example due to lockdown measures.
+Nesting in waves allowed to control for general exogenous developments, such as general decreases in well-being in the population, for example due to lockdown measures.
 Thus, there was no need additionally to control for specific phases or measures of the lockdown.
 Predictors were modeled as fixed effects.
-They included all types and channels of social media communication separated into within and between-person factors, as well as stable and varying covariates.
+They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
 All predictors were included simultaneously and in each of the three models.
@@
 To avoid overfitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
-Finally, REWB-models cannot model scales as latent variables.
+Finally, REWB-models cannot model latent variables.
 To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
+Respondents who answered less than 50% of all questions were removed.
+The remaining missing responses were imputed using predictive mean matching.
 
-For more information on the analyses, for example a complete documentation of the models and their results, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
+For more information on the analyses, a complete documentation of the models and results, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
 
@@
 The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
-Reliability was also very high, $\omega$ = `r rel_aff_pos %>% my_round("std")`.
+Reliability was high, $\omega$ = `r rel_aff_pos %>% my_round("std")`.
 
@@
 The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
-Reliability was also high, $\omega$ = `r rel_aff_neg %>% my_round("std")`.
+Reliability was high, $\omega$ = `r rel_aff_neg %>% my_round("std")`.
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement other variables as varying covariates.
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement additional variables as varying covariates.
 However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
@@
 # Results
-## Preregistered Analyses
 
-When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 
+## Preregistered Analyses
+
 The study's main hypothesis was that the effects of social media use on well-being would be trivial.
@@
 For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
-All confidence intervals include zero; hence, all effects were also statistically non-significant.
-<!-- Only one effect did not include zero. -->
-<!-- If people posted more about COVID-19 related aspects than they usually did, life satisfaction increased (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Posting" & type == "within") %>% select(conf.high) %>% round(2)`]).
--->
-<!-- However, because the effect did not fall outside of our null region, it's likely not large enough to be considered meaningful. -->
+All confidence intervals included zero; hence, all effects were also statistically non-significant.
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
@@
 ```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
-The SESOI was |0.30| for life satisfaction and |0.15| for affect.
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was \\textit{b} = |0.30| for life satisfaction and \\textit{b} = |0.15| for affect.
 Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
@@
 
-Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
-<!-- First, respondents who across all waves used social media more frequently than others to read about COVID-19 reported higher levels of positive affect than others (_b_ = `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% round(2)`]).
--->
-<!-- However, note that this effect still fell completely within the null-region. -->
-<!-- Hence, although positive the effect was considered too small to matter practically. -->
+Regarding between-person relations, about which no hypotheses were formulated, only three effects didn't include zero.
 Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
 Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-Interestingly, Respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-However, note that the effect were still completely inside of the null region, hence not considered large enough to be practically relevant.
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
+However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
 
@@
 Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
+However, when controlling for potential confounders, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
 
@@
 The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
-The SESOI was |0.30| for life satisfaction and |0.15| for affect.
+The SESOI was \\textit{b} = |0.30| for life satisfaction and \\textit{b} = |0.15| for affect.
 Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
@@
 Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ simultaneous levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
 In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
@@
 Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
-However, please note that this effect again was not considered large enough to be practically relevant.
+However, please note that this effect again was not large enough to be considered practically relevant.
 
 Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, five effects stopped being significant if they were controlled for additional variables.
+Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
 For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
@@
 
-In what follows, I briefly report some exploratory analyses that I did not preregister.
-First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected covariates.
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected cherry-picked covariates.
 Because each variable had different response options, we would need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
 Therefore, I report the results of the standardized scales,
-which allows for a better comparison across variables.
-In addition, as a rough estimate for the SESOI, we can use on the typical convention that small effects start at _r_ = |.10|.
+which allows for a better comparison across the differently scaled variables.
+For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
 The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or excercising.
+This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
 For an overview, see Figure \@ref(fig:res-control).
@@
 
-To find out whether my inferences were robust across legitimate alternative analyses, I reran the analyses also using standardized estimates,  mean scores instead of factor scores, and with a data set where missing data were not imputed.
+To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
 The results were virtually the same.
@@
 The data come from a panel study with 24 waves and are representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates.
+In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, aiming to assess causality.
 The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 did not show meaningful changes in their well-being.
+People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
 
 The results imply that COVID-19 related social media use does not seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, showing that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or excercising.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
 According to this study, popular fears that "doomscrolling" or overusing social media during times of crises don't seem to be justified.
@@
 First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so the results are ambivalent.
+On the other hand, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
 Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
@@
 Instagram, on the other hand, was related to slightly increased simultaneous levels of life satisfaction.
-The positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
+To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
 That said, all these effects were still very small and arguably too small to matter. 
@@
 
-Another last interesting observation is that life satisfaction was remarkably stable.
+Finally, another interesting observation is that life satisfaction was remarkably stable.
 Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
@@
 The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
-Together, this allows for a good perspective on potential causality.
+Together, this allows for an improved perspective on assessing causality.
 That said, causality necessitates temporal order, and the cause needs to precede the effect.
@@
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic to expect that extremely changing only _one_ of these aspects should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to _two_ activities (e.g.
-regularly exercising _and_ establishing a reading habit) should then cause detectable improvements in well-being?
-<!-- In other words, the beneficial effect of a particular activity is large enough if people feel a difference when implementing two of those activities.  -->
-Practically, this would imply a SESOI half as large as I have defined here, that is _b_ = |.15| for well-being and _b_ = |.075| for affect.
-In the case of this study, however, reducing the SESOI would not even make a difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+Is it really realistic to expect that extremely changing only _one_ of these aspects should manifest in a detectable change in well-being?
+Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g.
+regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
+Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for well-being and _b_ = |.075| for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
 However, at all events future research needs to start a thorough conversation on what effect sizes are considered meaningful and what not.
-With this study I hope to provide some first input and guidelines.
+With this study I again hope to provide some first input and guidelines.
 
@@
 The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of an entire population, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
 
@@
 However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's the aim to improve well-being, it might hence be fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+If it's the aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 

### 2021-12-01 manuscript.tex
- updated language

@@
 \abstract{
-In times of crisis such as the Corona pandemic, it is important that citizens stay informed about recent events, the latest political decisions, or mandatory protection measures.
+In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
@@
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects cross lagged panel models, controlling for several stable and varying covariates.
-Results showed that the effects of various types of COVID-19 related social media use on well-being were very small, arguably too small to matter.
+I ran three random effects within between models, controlling for several stable and varying confounders Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded.
@@
 These concerns seem justified:
-A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
-283), which implies negative consequences for well-being.
+A study with 6,233 people from Germany that was conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
+283).
 
@@
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
-To capture well-being in this study, I thus build on life satisfaction, positive affect, and negative affect.
+To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
 Together, this should provide an encompassing perspective on potential media effects.
@@
 The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
+First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
 Whereas the first four levels focus on the \emph{channel}, the last two address the \emph{type} of communication.
@@
 I distinguish (a) \emph{reading} (passive), \emph{posting} (active), and \emph{liking and sharing} COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affect well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially different effects.
-Twitter might have a different effect as compared to WhatsApp because of their different affordances.
+Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
+Branded apps are separate entities with potentially divergent effects.
+Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
 For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
@@
 (2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, and found both positive and negative relations with well-being.
+Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
 In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
@@
 Together, the literature emphasizes potentially negative effects of social media as news use (see also Liu \& Tong, 2020; Riehm et al., 2020).
-However, note that all of these findings represent between-person relation stemming from cross-sectional data (see below).
+However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
 We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
 
-The question of whether and how \emph{general} social media use affect well-being, on the other hand, is well-researched.
+The question of whether and how social media use affects well-being \emph{in general}, on the other hand, is well-researched.
 This also holds true for the different types of communication such as active or passive use.
 A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
-But what determines whether or not an effect is considered small or trivial?
-If we refer to standardized effect sizes, according to Cohen (1992) small effect sizes start at \emph{r} = .10.
+What determines whether or not an effect is considered small or trivial?
+As a starting point, we could refer to standardized effect sizes.
+According to Cohen (1992), small effect sizes start at \emph{r} = .10.
 And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
 
-These overviews are well aligned with several individual new studies employing advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the majority the effects were neutral.
+These overviews are well aligned with several individual studies employing advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
+For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
 In conclusion, most effects are likely somewhere between trivial and small.
@@
 If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news might be depressing, perhaps even changing some general perspectives on life.
+Repeatedly consuming such news can be depressing, perhaps even changing some general perspectives on life.
 That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.
@@
 
-Together, the strongest argument seems to be that \emph{in general} the effects of social media on well-being are, on average, small at best.
-Because this study only looks at \emph{one part} of social media use---namely, COVID-19 related interactions---it is very focused, and the overall potential of the effects should diminish further.
+Together, the strongest argument to me is that \emph{in general} the effects of social media on well-being are, on average, small at best.
+Because this study only looks at \emph{one part} of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
 Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
@@
 Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences, it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
+First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
 Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
@@
 However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
-Standardized effect sizes are determined by a sample's variance.\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.}, which is problematic:
+Standardized effect sizes are determined by a sample's variance,\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
+Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.} which is problematic:
 The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
@@
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
+(For more information, see Method section ``Inference criteria.'')
 
@@
 
-To explain, consider the following example.
+To illustrate, consider the following example.
 Imagine that a person suddenly starts using social media much more than usual, and then after some time becomes less satisfied with their life.
@@
 
-Thus, only if we can control for \emph{all} confounders, can we correctly estimate causality without bias (Rohrer, 2018).
-Obviously, we can never be entirely sure that we have included all confounders, which makes absolute statements regarding causality virtually impossible.
+Thus, only when controlling for \emph{all} relevant confounders, we can correctly estimate causality without bias (Rohrer, 2018).
+Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
 In addition, when determining the overall causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
-Complicating matters, often it is unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
-However, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible, while knowing that absolute certainty regarding causality cannot be reached.
-
-When searching for suitable candidates for confounders, we look for variables that affect both media use and well-being.
+Complicating matters further, it is often unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
+However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
+The aim should therefore be to collect as many varying and nonvarying confounders as possible (which it seems is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
+
+When searching for suitable candidates for confounders, we should look out for variables that affect both media use and well-being.
 Controlling for these factors isolates the actual effect of social media use on well-being.
@@
 
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are very likely to do so, and which also are not mediators:
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also are not mediators:
 gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables are likely influenced by social media use to a meaningful extent.
+I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
 
@@
 For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are necessary.
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
 In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
@@
 
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) I preregistered on the Open Science Framework.
+The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework.
 The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
@@
 Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is very large, and it is hence well-equipped reliably to detect also small effects.
+The sample is large, and it is hence well-equipped reliably to detect also small effects.
 In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
 To this end, I adopted the interval testing approach as proposed by Dienes (2014).
-On the basis of the SESOI, I then defined a null region (in Bayesian terms, this is called the region of practical equivalence, ROPE).
+On the basis of the SESOI, I then defined a null region.
 In what follows, I explain how I determined the SESOI and the null region.
@@
 Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
-In a statistical regression, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
+In a statistical regression analysis, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
 We would therefore expect a SESOI of \emph{b} = 1.57 / 4 = 0.39.
 For affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
-Because we are agnostic as to whether the effects are positive or negative, the null region will include negative and positive effects.
+Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
 Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).}
@@
 
-To illustrate, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is only trivial is supported.
+To illustrate what this means in practice, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
 If the effects interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
@@
 To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
-Finally, REWB-models cannot model scales as latent variables.
+Finally, REWB-models cannot model latent variables.
 To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-
-For more information on the analyses, for example a complete documentation of the models and their results, see \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website}.
+Respondents who answered less than 50\% of all questions were removed.
+The remaining missing responses were imputed using predictive mean matching.
+
+For more information on the analyses, a complete documentation of the models and results, see \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website}.
 
@@
 The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \textit{p} = .032, cfi = 1.00, rmsea = .02, 90\% CI {[}.01, .03{]}, srmr = .01.
-Reliability was also very high, \(\omega\) = .85.
+Reliability was high, \(\omega\) = .85.
 
@@
 The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \textit{p} \textless{} .001, cfi = .97, rmsea = .08, 90\% CI {[}.07, .08{]}, srmr = .03.
-Reliability was also high, \(\omega\) = .89.
+Reliability was high, \(\omega\) = .89.
 
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (answer options: female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement other variables as varying covariates.
+(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
+I originally planned to implement additional variables as varying covariates.
 However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
@@
 
-\hypertarget{preregistered-analyses}{%
-\subsection{Preregistered Analyses}\label{preregistered-analyses}}
-
-When looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
+First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
 COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
@@
 
+\hypertarget{preregistered-analyses}{%
+\subsection{Preregistered Analyses}\label{preregistered-analyses}}
+
 The study's main hypothesis was that the effects of social media use on well-being would be trivial.
@@
 For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
-All confidence intervals include zero; hence, all effects were also statistically non-significant.
+All confidence intervals included zero; hence, all effects were also statistically non-significant.
 As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
@@
 \includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
-The SESOI was |0.30| for life satisfaction and |0.15| for affect.
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was \textit{b} = |0.30| for life satisfaction and \textit{b} = |0.15| for affect.
 Hence, all of the reported effects are considered not meaningful.}\label{fig:res-activity}
@@
 
-Regarding between-person relations, about which no hypotheses were formulated, only two effects didn't include zero.
+Regarding between-person relations, about which no hypotheses were formulated, only three effects didn't include zero.
 Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
 Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
-Interestingly, Respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
-However, note that the effect were still completely inside of the null region, hence not considered large enough to be practically relevant.
+Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
+However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
 
@@
 Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
+However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
 
@@
 The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
-The SESOI was |0.30| for life satisfaction and |0.15| for affect.
+The SESOI was \textit{b} = |0.30| for life satisfaction and \textit{b} = |0.15| for affect.
 Hence, all of the reported effects are considered not meaningful.}\label{fig:res-channels}
@@
 Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} simultaneous levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} simultaneous levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
-However, both effects were still completely inside of the null region, hence not considered large enough to be meaningful.
+Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
+However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
 In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
@@
 Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
-However, please note that this effect again was not considered large enough to be practically relevant.
+However, please note that this effect again was not large enough to be considered practically relevant.
 
 Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, five effects stopped being significant if they were controlled for additional variables.
+Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
 For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
@@
 
-In what follows, I briefly report some exploratory analyses that I did not preregister.
-First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected covariates.
+In what follows, I briefly report some exploratory analyses that weren't preregistered.
+First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected cherry-picked covariates.
 Because each variable had different response options, we would need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
 Therefore, I report the results of the standardized scales,
-which allows for a better comparison across variables.
-In addition, as a rough estimate for the SESOI, we can use on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
+which allows for a better comparison across the differently scaled variables.
+For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
 The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or excercising.
+This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
 For an overview, see Figure \ref{fig:res-control}.
@@
 
-To find out whether my inferences were robust across legitimate alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
+To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
 The results were virtually the same.
@@
 The data come from a panel study with 24 waves and are representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates.
+In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, aiming to assess causality.
 The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 did not show meaningful changes in their well-being.
+People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
 
 The results imply that COVID-19 related social media use does not seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, showing that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or excercising.
+Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
 According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
@@
 Instagram, on the other hand, was related to slightly increased simultaneous levels of life satisfaction.
-The positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
+To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
 That said, all these effects were still very small and arguably too small to matter.
@@
 The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
-Together, this allows for a good perspective on potential causality.
+Together, this allows for an improved perspective on assessing causality.
 That said, causality necessitates temporal order, and the cause needs to precede the effect.
@@
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it realistic to expect that extremely changing only \emph{one} of these aspects should already manifest in a detectable change in well-being?
-Or would it make more sense to expect that thoroughly committing to \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause detectable improvements in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, that is \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, reducing the SESOI would not even make a difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
+Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
+Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
+Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
+In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
 However, at all events future research needs to start a thorough conversation on what effect sizes are considered meaningful and what not.
-With this study I hope to provide some first input and guidelines.
+With this study I again hope to provide some first input and guidelines.
 

### 2021-12-01 manuscript.Rmd
- updated website
- changed fig caption

@@
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was \\textit{b} = |0.30| for life satisfaction and \\textit{b} = |0.15| for affect.
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
 Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
@@
 The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
-The SESOI was \\textit{b} = |0.30| for life satisfaction and \\textit{b} = |0.15| for affect.
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
 Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
@@
 However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
-
-```{r tab-aff-neg}
-# apa_print(model_aff_neg_lmer)
-# tab <- broom.mixed::tidy(model_aff_neg_lmer, conf.int = T) %>% 
-#  select(term, estimate, conf.low, conf.high, p.value)
-#  kable()
-```
 

### 2021-12-01 manuscript.Rmd
- language updates

@@
 However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
-In this preregistered study, I investigate whether using social media for COVID-19 related reasons might put personal well-being at risk.
+In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects within between models, controlling for several stable and varying confounders Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
+I ran three random effects within between models, controlling for several stable and varying confounders.
+Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
@@
 During the COVID-19 pandemic, 
-numerous events unfolded in quick succession.
-Several open questions emerged.
+numerous events unfolded in quick succession and several open questions emerged.
 How dangerous is the virus? 
@@
 
-Some people could not stop using social media to learn about COVID-19 related news.
-A new phenomenon termed "doomscrolling" emerged:
+Some people actually couldn't stop using social media to learn about COVID-19 related news.
+A new phenomenon emerged, termed "doomscrolling":
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
@@
 These concerns seem justified: 
-A study with 6,233 people from Germany that was conducted during the pandemic found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
+A study with 6,233 people from Germany conducted during the pandemic found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
 283].
@@
 To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
-The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide)`.
+The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants.
 <!-- At least 1,500 participants took part per wave, and the sample is representative of the Austrian population. -->
@@
 One of the most prominent indicators of well-being is life satisfaction.
-In my view, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
+In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
@@
 First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
-Whereas the first four levels focus on the _channel_, the last two address the _type_ of communication.
+Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
 To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
@@
 Specifically, I differentiate between active and passive use.
-I distinguish (a) _reading_ (passive), _posting_ (active), and _liking and sharing_ COVID-19 related posts (both active and passive).
+I distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
 Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
@@
 In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
-In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.
+In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
 
-First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
+First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
 Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
@@
 Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might be not the worst idea.
+In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
 Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
@@
 <!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-Therefore, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
+Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction. 
-(For more information, see Method section "Inference criteria.")
+(For more information, see Methods section "Inference Criteria.")
 
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs help investigate causality.
+In non-experimental studies, longitudinal designs can help investigate causality.
 Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements [@rohrerTheseAreNot2021].
@@
 However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which it seems is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
+The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
 
@@
 
-When searching for suitable candidates for confounders, we should look out for variables that affect both media use and well-being. 
+When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being. 
 Controlling for these factors isolates the actual effect of social media use on well-being.
@@
 
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also are not mediators:
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
 <!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
@@
 
-Next to including covariates, it is now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
+Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
 If a specific person changes their media diet, we need to measure how this behavior affects _their own_ well-being.
@@
 Finding the right interval between cause and effect is crucial.
-For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
+For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
 <!-- Finding the right interval is difficult. -->
@@
 In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there simultaneous changes in their well-being?
+I analyze if when a person changes their social media diet, are there _simultaneous_ changes in their well-being?
 When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
@@
 The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
-In some cases I could not execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
+In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
 The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html) (https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
@@
 
-## Inference criteria
+## Inference Criteria
 
@@
 <!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
-To illustrate what this means in practice, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is trivial is supported.
-If the effects interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+Let's briefly illustrate what this means in practice.
+If the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 
-COVID-19 related social media use focused on communication _types_ was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
 The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
@@
 
-COVID-19 related social media use focused on _channels_ was measured with five variables.
+COVID-19 related social media use focused on channels was measured with five variables.
 The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
@@
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_activity.pdf")
@@
 
-Regarding between-person relations, about which no hypotheses were formulated, only three effects didn't include zero.
+Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
 Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
@@
 ```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are considered not meaningful.", out.width = "\\textwidth", cache=F}
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_channel.pdf")
@@
 In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected cherry-picked covariates.
-Because each variable had different response options, we would need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
+First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
 Therefore, I report the results of the standardized scales,
@@
 ```{r res-control, fig.cap="Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service' because there were binary.", out.width = "\\textwidth", cache=F}
+All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F}
 knitr::include_graphics("figures/fig_results_control_std.pdf")
@@
 The data come from a panel study with 24 waves and are representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, aiming to assess causality.
+In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
 The results showed that within-person effects were trivial.
@@
 
-The results imply that COVID-19 related social media use does not seem to be particularly relevant for well-being.
+The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
 Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
@@
 If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
+First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
 Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
 Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
-Instagram, on the other hand, was related to slightly increased simultaneous levels of life satisfaction.
+Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
 To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
@@
 In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events future research needs to start a thorough conversation on what effect sizes are considered meaningful and what not.
-With this study I again hope to provide some first input and guidelines.
+However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
+Again, with this study I hope to provide some first input and guidelines.
 
@@
 
-In this study, COVID-19 related social media use did not meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
+In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
 However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's the aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 

### 2021-12-01 manuscript.tex
- language updates

@@
 To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related reasons might put personal well-being at risk.
+However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects within between models, controlling for several stable and varying confounders Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
+I ran three random effects within between models, controlling for several stable and varying confounders.
+Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded.
@@
 During the COVID-19 pandemic,
-numerous events unfolded in quick succession.
-Several open questions emerged.
+numerous events unfolded in quick succession and several open questions emerged.
 How dangerous is the virus?
@@
 
-Some people could not stop using social media to learn about COVID-19 related news.
-A new phenomenon termed ``doomscrolling'' emerged:
+Some people actually couldn't stop using social media to learn about COVID-19 related news.
+A new phenomenon emerged, termed ``doomscrolling'':
 Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
@@
 These concerns seem justified:
-A study with 6,233 people from Germany that was conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
+A study with 6,233 people from Germany conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
 283).
@@
 To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and has an overall sample size of 3018.
+The panel consists of 24 waves and has an overall sample size of 3,018 participants.
 The panel study collected a large number of psychological and demographic variables.
@@
 One of the most prominent indicators of well-being is life satisfaction.
-In my view, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being, because it represents a general appraisal of one's life.
+In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
 Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
@@
 First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
-Whereas the first four levels focus on the \emph{channel}, the last two address the \emph{type} of communication.
+Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
 To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
@@
 Specifically, I differentiate between active and passive use.
-I distinguish (a) \emph{reading} (passive), \emph{posting} (active), and \emph{liking and sharing} COVID-19 related posts (both active and passive).
+I distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
 Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
@@
 In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
-In advance, there does not seem to be a clear winner, and it's likely that both positive and negative effects are equally strong.
-
-First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which is more volatile and fluctuating.
+In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
+
+First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
 Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
@@
 Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
-Therefore, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
+Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
 
@@
 So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
-(For more information, see Method section ``Inference criteria.'')
+(For more information, see Methods section ``Inference Criteria.'')
 
@@
 The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs help investigate causality.
+In non-experimental studies, longitudinal designs can help investigate causality.
 Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
@@
 However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which it seems is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
-
-When searching for suitable candidates for confounders, we should look out for variables that affect both media use and well-being.
+The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
+
+When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being.
 Controlling for these factors isolates the actual effect of social media use on well-being.
@@
 
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also are not mediators:
+In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
 gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
@@
 
-Next to including covariates, it is now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
+Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
 If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
@@
 Finding the right interval between cause and effect is crucial.
-For example, if we want to measure the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
+For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
 If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
@@
 In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there simultaneous changes in their well-being?
+I analyze if when a person changes their social media diet, are there \emph{simultaneous} changes in their well-being?
 When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
@@
 The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases I could not execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
+In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
 The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website} (\url{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}).
@@
 \hypertarget{inference-criteria}{%
-\subsection{Inference criteria}\label{inference-criteria}}
+\subsection{Inference Criteria}\label{inference-criteria}}
 
@@
 
-To illustrate what this means in practice, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
-If the effects interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+Let's briefly illustrate what this means in practice.
+If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
 
@@
 
-COVID-19 related social media use focused on communication \emph{types} was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
+COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
 The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
@@
 
-COVID-19 related social media use focused on \emph{channels} was measured with five variables.
+COVID-19 related social media use focused on channels was measured with five variables.
 The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis on the following social media?''
@@
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was \textit{b} = |0.30| for life satisfaction and \textit{b} = |0.15| for affect.
-Hence, all of the reported effects are considered not meaningful.}\label{fig:res-activity}
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.}\label{fig:res-activity}
 \end{figure}
 
-Regarding between-person relations, about which no hypotheses were formulated, only three effects didn't include zero.
+Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
 Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
@@
 \includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text); the grey estimates are without control variables.
-The SESOI was \textit{b} = |0.30| for life satisfaction and \textit{b} = |0.15| for affect.
-Hence, all of the reported effects are considered not meaningful.}\label{fig:res-channels}
+The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
+The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
+Hence, all of the reported effects are not considered meaningful.}\label{fig:res-channels}
 \end{figure}
@@
 In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above, and to see if other variables showed meaningful effects, I also looked at the effect sizes of selected cherry-picked covariates.
-Because each variable had different response options, we would need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
+First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
+Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
 Therefore, I report the results of the standardized scales,
@@
 \includegraphics[width=\textwidth]{figures/fig_results_control_std} \caption{Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service' because there were binary.}\label{fig:res-control}
+All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}
 \end{figure}
@@
 If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
+First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
 Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
 Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo, Baumgartner, Peter, \& Valkenburg, 2018), which might help explain the results.
-Instagram, on the other hand, was related to slightly increased simultaneous levels of life satisfaction.
+Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
 To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.

### 2021-12-01 manuscript.Rmd
- unblinded ms

@@
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
-keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal."
+keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study."
 
@@
 linenumbers       : yes
-mask              : yes
+mask              : no
 draft             : no

### 2021-12-01 manuscript.tex
- unblinded ms

@@
   english,
-  man,mask,floatsintext]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
   pdftitle={No meaningful effects of COVID-19 related social media use on well-being},
+  pdfauthor={Tobias Dienlin1},
   pdflang={en-EN},
-  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.},
+  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.},
   hidelinks,
@@
 \makeatother
-\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.}
+\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal study.}
 \usepackage{lineno}

### 2021-12-13 analyses.RmD
- added three studies to lit review

@@
     toc_depth: 2
-    # code_folding: hide
+    code_folding: hide
 bibliography      : 
@@
 knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
-options(digits = 3, width = 200, scipen=999)
-```
+options(digits = 3, width = 250, scipen=999)
+```
+
+Here you can find the code and the results of all analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
 
@@
   )
-
-# rename variables
-# names(d_wide) <- sub("\\.\\.\\.", "§", names(d_wide)) # replace "..." with "_"
 ```
@@
     employment_fac = factor(.$employment, 
-                           # labels = c("Unselbststaendig erwerbstaetig (auch Lehrling), Privatwirtschaft",
-                           #            "Unselbststaendig erwerbstaetig (auch Lehrling), Oeffentlicher Dienst",
-                           #            "Selbststaendig erwerbstaetig",
-                           #            "Arbeitslos",
-                           #            "Pensionist*in",
-                           #            "Ausschliesslich haushaltsfuehrend",
-                           #            "Schueler/in, Student/in, Bildungskarenz",
-                           #            "Sonstiges (Dauerhaft arbeitsunfaehig oder Praesenz-/Zivildiener)",
-                           #            "In Elternkarenz")
                            labels = c("Unemployed",
@@
     edu_fac = factor(.$edu,
-                     # labels = c("Volksschule oder weniger", 
-                     #            "Hauptschule oder AHS Unterstufe",
-                     #            "Polytechnikum, BMS (Fachschule, z.B. HASCH)",
-                     #            "Lehre, Berufsschule",
-                     #            "AHS mit Matura",
-                     #            "BHS mit Matura (z.B. HTL, HAK, HBLA etc.)",
-                     #            "Hochschulverwandte Lehranstalt oder Kolleg",
-                     #            "Bachelor",
-                     #            "Magister / Master / Diplomingenieur / Fachhochschule",
-                     #            "Doktor / PhD"),
                      labels = c("No degree", 
@@
 
-<!-- Filter waves without data -->
-
-<!-- ```{r eval=F} -->
-<!-- d_long_test <- d_long %>% -->
-<!--   filter(!is.na(.[vars_varying])) -->
-<!-- ``` -->
-
-Recode values
+Recode values.
 
@@
 
-<!-- Create vectors with names -->
-
-<!-- ```{r} -->
-<!-- # create vector of names -->
-<!-- vars_varying <- names(select(d_long, "health_b":"loc_cntrl_int_4_w")) -->
-<!-- predictors_varying <- select(d_long, "health_b":"loc_cntrl_int_4_w") %>%  -->
-<!--   select(-starts_with("aff_"), -starts_with("life_")) %>%  -->
-<!--   names() -->
-<!-- ``` -->
-
 # Impute data
@@
 
-The data fit the model very well, `#td::fit_txt(cfa_aff_pos)`.
+The data fit the model very well, `r td::fit_txt(cfa_aff_pos)`.
 Let's next inspect reliability
@@
 
-The data fit the model very well, `#td::fit_txt(cfa_aff_neg)`.
+The data fit the model very well, `r td::fit_txt(cfa_aff_neg)`.
 
@@
 summary(model_soc_med_tw)
-
-# Bayesian Approach
-# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp
-# Also throws a warning
-# model_soc_med_tw <- brm(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp, control=list(adapt_delta=0.82))
-# Currently doesn't work, likely for lack of memory
 ```
@@
 
-The data fit the model very well, `#td::fit_txt(cfa_loc_cntrl_int)`.
+The data fit the model very well, `r td::fit_txt(cfa_loc_cntrl_int)`.
 
@@
                                         res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        # hh_income_b + # large dropout (900 obs)
-                                        # hh_adults + # too much dropout
-                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
-                                        # home_owner + # too much dropout
                                         med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 

### 2021-12-13 analyses_additional.Rmd
- added three studies to lit review

@@
 ```
+
+Here you can find the code and the results of all additional analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
 
@@
 
+Save results so that we can extract them in the manuscript.
+
 ```{r}

### 2021-12-13 manuscript.Rmd
- added three studies to lit review

@@
   Tobias Dienlin, Department of Communication, University of Vienna.
+
+note: |
+  Preprint submitted for publication. Please cite carefully.
 
@@
 
-## Defining Well-being and Media Use
+## Understanding Well-being and Media Use
 
@@
 @stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
-Together, the literature emphasizes potentially negative effects of social media as news use [see also @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
+A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations [@sewallObjectivelyMeasuredDigital2021].
+The results showed that digital technology did not have any significant effects on mental health [for a similar study with virtually identical results, see @bradleyStressMoodSmartphone2021].
+Together, the literature is mixed, with a slight focus on the negative effects of social media as news use [see also dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
 However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
@@
 If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news can be depressing, perhaps even changing some general perspectives on life.
+Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life.
 <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
@@
 On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
+This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being [but see @bradleyStressMoodSmartphone2021; or @sewallObjectivelyMeasuredDigital2021].
 However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
@@
 The data can only be used for scientific purposes.
+
+# Acknowledgements
+
+I would like to thank Niklas Johannes for providing valuable feedback on this manuscript.

### 2021-12-13 manuscript.tex
- added three studies to lit review

@@
 
+\note{
+
+Preprint submitted for publication. Please cite carefully.
+
+}
+
 \abstract{
@@
 
-\hypertarget{defining-well-being-and-media-use}{%
-\subsection{Defining Well-being and Media Use}\label{defining-well-being-and-media-use}}
+\hypertarget{understanding-well-being-and-media-use}{%
+\subsection{Understanding Well-being and Media Use}\label{understanding-well-being-and-media-use}}
 
@@
 Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
-Together, the literature emphasizes potentially negative effects of social media as news use (see also Liu \& Tong, 2020; Riehm et al., 2020).
+A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use---objectively measured via screenshots of screen-time applications---on mental health, separating within- and between-person relations (Sewall, Goldstein, \& Rosen, 2021).
+The results showed that digital technology did not have any significant effects on mental health (for a similar study with virtually identical results, see Bradley \& Howard, 2021).
+Together, the literature is mixed, with a slight focus on the negative effects of social media as news use {[}see also dornemannHowGoodBad2021; Liu and Tong (2020); Riehm et al. (2020){]}.
 However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
@@
 If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news can be depressing, perhaps even changing some general perspectives on life.
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.
+Consuming such news can be depressing (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021), perhaps even changing some general perspectives on life.
+That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects (Dörnemann, Boenisch, Schommer, Winkelhorst, \& Wingen, 2021).
 
@@
 On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being.
+This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being (but see Bradley \& Howard, 2021; or Sewall, Goldstein, \& Rosen, 2021).
 However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
@@
 \url{https://doi.org/10.31234/osf.io/ftygp}
+
+\leavevmode\hypertarget{ref-bradleyStressMoodSmartphone2021}{}%
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. OSF Preprints. \url{https://doi.org/10.31219/osf.io/frvpb}
 
@@
 \url{https://doi.org/doi:10.31887/DCNS.2020.22.2/tdienlin}
+
+\leavevmode\hypertarget{ref-dornemannHowGoodBad2021}{}%
+Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
+(2021).
+\emph{How do good and bad news impact mood during the {Covid}-19 pandemic?
+{The} role of similarity}.
+OSF Preprints.
+\url{https://doi.org/10.31219/osf.io/sy2kd}
 
@@
 
+\leavevmode\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}%
+Sewall, C.
+J.
+R., Goldstein, T.
+R., \& Rosen, D.
+(2021).
+Objectively measured digital technology use during the {COVID}-19 pandemic: {Impact} on depression, anxiety, and suicidal ideation among young adults.
+\emph{Journal of Affective Disorders}, \emph{288}, 145--147.
+\url{https://doi.org/10.1016/j.jad.2021.04.008}
+
 \leavevmode\hypertarget{ref-stainbackCOVID1924News2020}{}%
@@
 
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+I would like to thank Niklas Johannes for providing valuable feedback on this manuscript.
+
 \end{document}

### 2023-01-30 analyses.RmD
- major updates to preprint

@@
     code_folding: hide
+    df_print: paged
 bibliography      : 
@@
   - "bibliography/references.bib"
-  # pdf_document:
-  #   toc: TRUE
 ---
@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
-options(digits = 3, width = 250, scipen=999)
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache=T, rows.print=20)
+options(digits=2, width=260, scipen=999)
 ```
 
-Here you can find the code and the results of all analyses. To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
+ - Here you can find the code and the results of all main analyses. 
+ - To see the code, click on button "Code".
+ - Alternatively, you can download the rmd file from the github repo.
 
@@
 
-Load packages.
+Load packages & workspace.
 
-```{r r-setup, results='hide'}
+```{r r-setup, results='hide', cache=F}
 # install packages
@@
 # define packages
-packages <- c("brms", "devtools", "GGally", "ggplot2", 
+packages <- c("broom.mixed", "brms", "corrr", "devtools", "GGally", "ggplot2", 
               "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
-              "magrittr", "mice", "mvnormalTest", 
-              "PerFit", "psych", "quanteda.textstats", "semTools", "tidyverse")
+              "magrittr", "mice", "PerFit", "performance", "psych", 
+              "quanteda.textstats", "scales", "semTools", "tidyverse")
 
@@
 lapply(c(packages, "td"), library, character.only = TRUE)
-```
 
-## Custom functions
-
-Find variables in dataset.
-
-```{r}
-find_var <- function(name, data = doc)(
-  # finds the variables names for an item for each wave
-    data %>% 
-    filter(Label == name) %>% 
-    select(variable) %>% 
-    unlist() %>% 
-    set_names(sub("\\_.*", "", .))
-)
-```
-
-Extract characteristics from fitted lmer models.
-
-```{r}
-get_specs <- function(model){
-    # Get mean, max, and min values
-
-    dat <- coefficients(model)$wave
-    specs <- data.frame(
-    sd = attr(VarCorr(model), "sc"),
-    min = min(dat),
-    max = max(dat),
-    mean = mean(dat$`(Intercept)`)
-  )
-}
-```
-
-Get data from fitted lmer objects for descriptives.
-
-```{r}
-get_dat <- function(model){
-  coefficients(model)$wave %>% 
-  tibble::rownames_to_column("wave") %>% 
-  rename(value = "(Intercept)") %>% 
-  mutate(wave = as.integer(.$wave),
-         value = as.numeric(.$value)) 
-} 
-```
-
-Determine average reliability for measures across waves
-
-```{r}
-get_rel <- function(data, waves=24){
-
-  # extract average reliability from lavaan fitted cfa with several groups
-  reliability(data) %>% 
-  unlist() %>% 
-  matrix(5, waves) %>% 
-  as.data.frame() %>% 
-  set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
-  summarise(omega = rowMeans(.["omega",])) %>% 
-  return()
-}
-```
-
-Make graph of variables' development.
-
-```{r}
-make_graph <- function(model, title, ll, ul, line = FALSE, labels = FALSE, lmer=TRUE, legend=TRUE){
-
-    if(isTRUE(lmer)){
-      dat <- get_dat(model)
-    } else{
-    dat <- model
-  }
-
-  graph <- 
-    ggplot(dat, aes(wave, value, color = dimension)) +
-    {if(legend) theme(
-      legend.position = c(0.85, 0.8),
-      legend.title = element_blank()
-      )} + 
-    {if(!legend) theme(
-      legend.position = "none"
-    )} +
-    theme(axis.title.x=element_blank()) +
-    coord_cartesian(ylim = c(ll, ul)) +
-    ggtitle(title) +
-    {if(line) geom_smooth(se = FALSE, method = 'loess')} +
-    # geom_point() +
-    scale_color_manual(values=c("dodgerblue3", "deepskyblue2", "magenta2", "green2", "red"))
-  graph
-}
-```
-
-Get data of lmer objects for results.
-
-```{r}
-# get data
-get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, type){
-
-  dat_fig_results <- 
-  broom.mixed::tidy(data_aff_neg, conf.int = T) %>% 
-  mutate(dv = "aff_neg") %>% 
-  rbind(
-    broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
-    mutate(dv = "aff_pos")
-  ) %>% 
-  rbind(
-    broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
-    mutate(dv = "life_sat")
-  ) %>% 
-  mutate(
-    type = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
-    iv = gsub("_(w|b)\\>", "", .$term)
-         ) %>%
-  mutate(
-    type = factor(.$type, 
-                  levels = c("within", "between")),
-    dv = factor(.$dv, 
-                levels = c("life_sat", "aff_pos", "aff_neg"),
-                labels = c("Life satisfaction", "Positive affect", "Negative affect"))
-  ) %>% 
-  select(dv, iv, type, estimate, conf.low, conf.high, p.value)
-
-  # select Social Media type of activity
-  if(type == "activity") {
-    dat_fig_results %<>%
-    filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
-    mutate(
-      iv = factor(.$iv, 
-                  levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
-                  labels = c("Posting", "Liking & Sharing", "Reading"))
-    )
-  } else if(type == "channel"){
-    dat_fig_results %<>%
-    filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
-    mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
-  } else if(type == "control"){
-    dat_fig_results %<>%
-      filter(iv %in% c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem")) %>% 
-      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m", "employment_facPublic service", "act_spo", "sat_dem"), 
-                         labels = c("Male", "Health", "Internal locus of control", "Employed public service", "Sport", "Satisfaction democracy")))
-  }
-  return(dat_fig_results)
-}
-
-```
-
-Make graph of effects
-
-```{r}
-make_graph_res <- function(data_1, data_2 = NULL, sesoi = NULL){
-  ggplot(data_1, aes(x = estimate, y = iv)) + 
-  scale_color_manual(values = c("grey40", "deepskyblue")) +
-  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
-  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
-                lwd = .75, height = .2) +
-  geom_point(size = 2) +
-  {if(!is.null(data_2)) geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), data_2,
-                lwd = .75, height = .2, color = "grey",
-                position = position_nudge(y = -.3)
-                )} +
-  {if(!is.null(data_2)) geom_point(data = data_2, size = 2, color = "grey",
-             position = position_nudge(y = -.3)
-             )} +
-  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
-    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
-  {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
-    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data_1, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
-  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
-  {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   theme(
-    axis.title.y = element_blank(),
-    plot.title = element_text(hjust = .5),
-    panel.spacing = unit(.9, "lines"),
-    text = element_text(size = 12),
-    axis.text.y = element_text(hjust = 0)
-    ) +
-  facet_grid(rows = vars(type),
-             cols = vars(dv),
-             scales = "free",
-             space = "free_y") 
-}
-```
-
-Extract factor scores.
-
-```{r}
-get_fs <- function(model) {
-  # the aim is to get factor scores on original scaling
-  # hence export factor values for all items
-  # then compute average
-  dat_ov <- lavaan::lavPredict(model, type = "ov", assemble = TRUE) %>% 
-    mutate(fs = rowMeans(.[1:3]))
-  return(dat_ov$fs)
-}
-```
-
-# Data Wrangling
-
-```{r}
-doc <- read_csv("documentation/10094_vi_de_v4_0.csv")
-d_raw <- read_csv("data/10094_da_de_v4_0.csv")
-```
-
-Identify the names of each item for each wave.
-
-```{r}
-# identify variable names of items, using custom function
-health <- find_var("Gesundheitszustand") # wsn't collected at wave 21
-aff_neg_1 <- find_var("Depressivitaet: einsam")
-aff_neg_2 <- find_var("Depressivitaet: aergerlich")
-aff_neg_3 <- find_var("Depressivitaet: so niedergeschlagen")
-aff_neg_4 <- find_var("Depressivitaet: sehr nervoes")
-aff_neg_5 <- find_var("Depressivitaet: aengstlich")
-aff_neg_6 <- find_var("Depressivitaet: bedrueckt und traurig")
-aff_pos_1 <- find_var("Depressivitaet: ruhig und gelassen")
-aff_pos_2 <- find_var("Depressivitaet: gluecklich")
-aff_pos_3 <- find_var("Depressivitaet: voller Energie")
-act_wrk <- find_var("Zuhause verlassen: Arbeit")
-act_spo <- find_var("Zuhause verlassen: Sport")
-act_frn <- find_var("Zuhause verlassen: Freunde oder Verwandte treffen")
-act_sho <- find_var("Zuhause verlassen: Essen einkaufen")
-act_pet <- find_var("Zuhause verlassen: Haustier ausfuehren")
-sat_dem <- find_var("Demokratiezufriedenheit: Oesterreich")
-work_h <- find_var("Arbeitsstunden: Jetzt pro Woche")
-work_homeoff <- find_var("Aenderung berufliche Situation: Home-Office")
-hh_income <- find_var("Aktuelles Haushaltseinkommen")
-med_txt_kro <- find_var("Mediennutzung: Kronen Zeitung oder www.krone.at")
-med_txt_sta <- find_var("Mediennutzung: Der Standard oder derstandard.at")
-med_txt_pre <- find_var("Mediennutzung: Die Presse oder diepresse.com")
-med_txt_oes <- find_var("Mediennutzung: Oesterreich oder oe24.at")
-med_txt_kur <- find_var("Mediennutzung: Kurier oder kurier.at")
-med_txt_slz <- find_var("Mediennutzung: Salzburger Nachrichten oder salzburg.at")
-med_txt_son <- find_var("Mediennutzung: Sonstige oesterreichische Tageszeitungen")
-med_vid_orf <- find_var("Mediennutzung: ORF (Nachrichten)")
-med_vid_pri <- find_var("Mediennutzung: Privatfernsehen (Nachrichten)")
-soc_med_fb <- find_var("Soziale Medien: Facebook")
-soc_med_tw <- find_var("Soziale Medien: Twitter")
-soc_med_ig <- find_var("Soziale Medien: Instagram")
-soc_med_yt <- find_var("Soziale Medien: Youtube")
-soc_med_wa <- find_var("Soziale Medien: WhatsApp")
-soc_med_read <- find_var("Soziale Medien Aktivitaet: Postings zu Corona anderer lesen")
-soc_med_like_share <- find_var("Soziale Medien Aktivitaet: Postings liken, teilen oder retweeten")
-soc_med_post <- find_var("Soziale Medien Aktivitaet: selber Postings zu Corona verfassen")
-life_sat <- find_var("Lebenszufriedenheit")
-risk_prop <- find_var("Risikobereitschaft")
-loc_cntrl_int_1 <- find_var("Psychologie: habe Leben selbst in der Hand")
-loc_cntrl_int_2 <- find_var("Psychologie: Belohnung durch Anstrengung")
-loc_cntrl_int_3 <- find_var("Psychologie: Fremdbestimmung")
-loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
-```
-
-Select variables.
-
-```{r}
-d_wide <- d_raw %>% 
-  select(
-    id = RESPID,
-    gender = SD_GENDER,
-    acc_bal = SD_ACCESS_BALCONY,
-    acc_gar = SD_ACCESS_GARDEN,
-    year_birth = SD_BIRTHYEAR,
-    born_aus = SD_BORN_AUSTRIA,
-    born_aus_prnts = SD_MIGR_BACKGR,
-    county = SD_BULA,
-    edu = SD_EDU,
-    employment = SD_EMPLSTATUS_FEB2020,
-    hh_adults = SD_HH_ADULTS,
-    hh_child18 = SD_HH_CHILD18,
-    hh_child17 = SD_HH_TEENS,
-    hh_child14 = SD_HH_CHILD14,
-    hh_child5 = SD_HH_CHILD5,
-    hh_child2 = SD_HH_CHILD2,
-    hh_oldfam = SD_HH_OLDERFAM,
-    hh_outfam = SD_HH_OUTERFAM,
-    hh_partner = SD_HH_PARTNER,
-    # hh_income = SD_HHINCOME_FEB2020,
-    home_sqm = SD_HOME_SQM,
-    home_owner = SD_HOMEOWNER,
-    # work_h = SD_WORKHOURS_FEB2020,
-    health = all_of(health),
-    life_sat = all_of(life_sat),
-    aff_neg_1 = all_of(aff_neg_1),
-    aff_neg_2 = all_of(aff_neg_2),
-    aff_neg_3 = all_of(aff_neg_3),
-    aff_neg_4 = all_of(aff_neg_4),
-    aff_neg_5 = all_of(aff_neg_5),
-    aff_neg_6 = all_of(aff_neg_6),
-    aff_pos_1 = all_of(aff_pos_1),
-    aff_pos_2 = all_of(aff_pos_2),
-    aff_pos_3 = all_of(aff_pos_3),
-    act_wrk = all_of(act_wrk),
-    act_spo = all_of(act_spo),
-    act_frn = all_of(act_frn),
-    act_sho = all_of(act_sho),
-    act_pet = all_of(act_pet),
-    sat_dem = all_of(sat_dem),
-    sat_dem = all_of(sat_dem),
-    work_h = all_of(work_h),
-    work_homeoff = all_of(work_homeoff),
-    hh_income = all_of(hh_income),
-    med_txt_kro = all_of(med_txt_kro),
-    med_txt_sta = all_of(med_txt_sta),
-    med_txt_pre = all_of(med_txt_pre),
-    med_txt_oes = all_of(med_txt_oes),
-    med_txt_kur = all_of(med_txt_kur),
-    med_txt_slz = all_of(med_txt_slz),
-    med_txt_son = all_of(med_txt_son),
-    med_vid_orf = all_of(med_vid_orf),
-    med_vid_pri = all_of(med_vid_pri),
-    soc_med_fb = all_of(soc_med_fb),
-    soc_med_tw = all_of(soc_med_tw),
-    soc_med_ig = all_of(soc_med_ig),
-    soc_med_yt = all_of(soc_med_yt),
-    soc_med_wa = all_of(soc_med_wa),
-    soc_med_like_share = all_of(soc_med_like_share),
-    soc_med_read = all_of(soc_med_read),
-    soc_med_post = all_of(soc_med_post),
-    risk_prop = all_of(risk_prop),
-    loc_cntrl_int_1 = all_of(loc_cntrl_int_1),
-    loc_cntrl_int_2 = all_of(loc_cntrl_int_2),
-    loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
-    loc_cntrl_int_4 = all_of(loc_cntrl_int_4)
-  )
-```
-
-Make new documentation with selected variables.
-
-```{r}
-doc_selected <- filter(doc, variable %in% c(
-  "RESPID",
-  "SD_GENDER",
-  "SD_ACCESS_BALCONY",
-  "SD_ACCESS_GARDEN",
-  "SD_BIRTHYEAR",
-  "SD_BORN_AUSTRIA",
-  "SD_MIGR_BACKGR",
-  "SD_BULA",
-  "SD_EDU",
-  "SD_EMPLSTATUS_FEB2020",
-  "SD_HH_ADULTS",
-  "SD_HH_CHILD18",
-  "SD_HH_TEENS",
-  "SD_HH_CHILD14",
-  "SD_HH_CHILD5",
-  "SD_HH_CHILD2",
-  "SD_HH_OLDERFAM",
-  "SD_HH_OUTERFAM",
-  "SD_HH_PARTNER",
-  "SD_HHINCOME_FEB2020",
-  "SD_HOME_SQM",
-  "SD_HOMEOWNER",
-  "SD_WORKHOURS_FEB2020",
-  health[1],
-  life_sat[1],
-  aff_neg_1[1],
-  aff_neg_2[1],
-  aff_neg_3[1],
-  aff_neg_4[1],
-  aff_neg_5[1],
-  aff_neg_6[1],
-  aff_pos_1[1],
-  aff_pos_2[1],
-  aff_pos_3[1],
-  act_wrk[1],
-  act_spo[1],
-  act_frn[1],
-  act_sho[1],
-  act_pet[1],
-  sat_dem[1],
-  work_h[1],
-  work_homeoff[1],
-  hh_income[1],
-  med_txt_kro[1],
-  med_txt_sta[1],
-  med_txt_pre[1],
-  med_txt_oes[1],
-  med_txt_kur[1],
-  med_txt_slz[1],
-  med_txt_son[1],
-  med_vid_orf[1],
-  med_vid_pri[1],
-  soc_med_fb[1],
-  soc_med_tw[1],
-  soc_med_ig[1],
-  soc_med_yt[1],
-  soc_med_wa[1],
-  soc_med_like_share[1],
-  soc_med_post[1],
-  soc_med_read[1],
-  risk_prop[1],
-  loc_cntrl_int_1[1],
-  loc_cntrl_int_2[1],
-  loc_cntrl_int_3[1],
-  loc_cntrl_int_4[1]
-  ))
-```
-
-Recode Variables
-
-```{r}
-d_wide %<>%
-  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
-  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
-  mutate(
-    male = 2 - .$gender,
-    age = 2021 - .$year_birth,
-    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
-    born_aus = 2 - .$born_aus,
-    home_owner = 2 - .$home_owner,
-    employment_fac = factor(.$employment, 
-                           labels = c("Unemployed",
-                                      "Industrie",
-                                      "Public service",
-                                      "Self-employed",
-                                      "Retired",
-                                      "Housekeeping",
-                                      "Student",
-                                      "Incapacitated",
-                                      "Parental Leave"),
-                            levels = c(4, 1:3, 5:8, 10) # make unemployment reference cat
-                           ),
-    edu_fac = factor(.$edu,
-                     labels = c("No degree", 
-                                "Middle school",
-                                "Vocational school",
-                                "Technical school",
-                                "High school",
-                                "Applied high school",
-                                "State college",
-                                "Bachelor",
-                                "Master",
-                                "PhD")
-                     ))
-```
-
-Make data from wide to long.
-
-```{r}
-d_long <-
-  d_wide %>% 
-  pivot_longer(
-    cols = health...W1:loc_cntrl_int_4...W9,
-    names_to = "item",
-    values_to = "value"
-  ) %>% 
-  separate(item, c("item", "wave"), sep = "\\.\\.\\.", extra = "merge") %>% 
-  pivot_wider(names_from = "item", values_from = "value")
-```
-
-Recode values.
-
-```{r}
-# recode such that higher values imply more strength / align with wording
-d_long %<>% 
-  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem, soc_med_fb:soc_med_post), 
-            funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
-  mutate_at(vars(loc_cntrl_int_1:loc_cntrl_int_4), 
-            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L))) %>% 
-  mutate_at(vars(born_aus_prnts), 
-            funs(recode(., `3` = 0L, `2` = 2L, `1` = 1L)))
-
-# recode inverted items
-d_long %<>%
-  mutate_at(vars(loc_cntrl_int_3, loc_cntrl_int_4), 
-            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L)))
-
-# recode other
-d_long %<>%
-  mutate(
-    wave = gsub("W", "", .$wave) %>% as.integer(),
-    id = as.integer(id)
-  )
-```
-
-Order dataset. First id, then wave.
-
-```{r}
-d_long %<>%
-  arrange(id, wave)
-```
-
-# Impute data
-
-Determine amount of missingness per respondent per wave. Because calculation takes long, this chunk is not executed but loaded from memory. To execute, set eval == TRUE.
-
-```{r eval=TRUE}
-vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem")
-
-# filter respondents with more than 50% missing data
-d_long_filt <- 
-  d_long %>% 
-  mutate(na_perc = rowSums(is.na(select(., all_of(vars_used))) / ncol(select(., all_of(vars_used))))) %>% 
-  filter(na_perc < .5)
-
-# impute missing data
-# exclude social media use data, because they were measured only on selected waves and would lead to the imputation of too much missing data
-vars_excl <- c("na_perc", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
-incl_ma <- d_long_filt %>% 
-  mutate(across(.cols = everything(), .fns = is.na))
-incl_ma[vars_excl] <- FALSE
-
-# init <- mice(select(d_long_test_filt, maxit = 0))
-# pred_ma <- init$predictorMatrix
-# pred_ma[, vars_excl] <- 0
-d_long_imp <- mice(d_long_filt, method = "pmm", m = 1, maxit = 30, 
-              where = incl_ma,
-              # predictorMatrix = pred_ma, 
-              seed = 180719, print = FALSE)
-d_long_imp <- mice::complete(d_long_imp)
-
-write_csv(d_long_imp, "data/data_imputed.csv")
-```
-
-Load imputed data from disc.
-
-```{r}
-d_long_imp <- read_csv("data/data_imputed.csv")
-```
-
-# Create within-between data
-
-Create means for scales
-
-```{r}
-d_long_imp %<>%
-  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
-         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
-         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
-```
-
-Create mean values for each person (between), and unique changes per wave (within).
-
-```{r}
-d_long_imp %<>% 
-  group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
-  ungroup()
-```
-
-And because later we want to rerun the results without imputed data, now the same for data without imputation.
-
-```{r}
-d_long %<>%
-  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
-         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
-         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))
-
-d_long %<>% 
-  group_by(id) %>% 
-  mutate_at(vars(health:loc_cntrl_int_4, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
-  ungroup()
+# load workspace
+load("data/workspace_1.RData")
 ```
@@
 # Measures
+
+Let's first inspect the individual measures, how they develop over time. For positive and negative affect, we also look at their factor structure, as they're measured with multiple items.
+
 ## Well-being
@@
 
+Let's inspect the development of life satisfaction across the study. We nest reponses inside participants and waves to get better results.
+
 ```{r}
-model_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_imp)
-summary(model_life_sat)
+fit_life_sat <- lmer(life_sat ~ (1 | id) + (1 | wave), d_long_100_imp)
+summary(fit_life_sat)
+dat_fig_life_sat <- data.frame(type = "Life satisfaction", dimension = "Life satisfaction", 
+                               get_dat(fit_life_sat))
+make_graph(dat_fig_life_sat, "Life Satisfaction", 1, 10)
 ```
@@
 
-Positive affect was measured as a scale. We hence first inspect factorial validity using CFA.
+Let's next expect the development across waves.
 
-We first test assumption of multivariate normality.
-
-```{r eval=FALSE}
-mardia(d_long_imp %>% select(aff_pos_1, aff_pos_2, aff_pos_3))
+```{r}
+model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_100_imp)
+summary(model_aff_pos)
+dat_fig_aff_pos <- data.frame(type = "Affect", dimension = "Positive", 
+                              get_dat(model_aff_pos))
+make_graph(dat_fig_aff_pos, "Positive Affect", 1, 5)
 ```
 
-Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+Positive affect was measured as a scale. We hence also inspect factorial validity using CFA.
+
+We first test assumption of multivariate normality. We focus on wave 1 here.
+
+```{r}
+d_long_100_imp %>%
+  filter(wave == 1) %>% 
+  select(aff_pos_1, aff_pos_2, aff_pos_3) %>% 
+  mardia()
+```
+
+Assumption of multivariate normal distribution was violated; hence, robust estimator will be used.
 
@@
 "
-cfa_aff_pos <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+cfa_aff_pos <- cfa(model, d_long_100_imp, group = "wave", estimator = "MLM")
 summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 The data fit the model very well, `r td::fit_txt(cfa_aff_pos)`.
-Let's next inspect reliability
+Let's next inspect reliability.
 
@@
 rel_aff_pos <- get_rel(cfa_aff_pos)
-rel_aff_pos
 ```
 
-The average reliability across all waves was omega = `r rel_aff_pos`, hence good.
+The average reliability across all waves was omega = `r rel_aff_pos[[1]]`, hence good.
 
-Let's next expect the development across waves.
+Let's now export factor scores for results reported in additional analyses.
 
 ```{r}
-model_aff_pos <- lmer(aff_pos_m ~ (1 | id) + (1 | wave), d_long_imp)
-summary(model_aff_pos)
-```
+# with imputed data
+cfa_aff_pos_50 <- cfa(model, d_long_50_imp, group = "wave", estimator = "MLM")
+d_long_50_imp$aff_pos_fs <- get_fs(cfa_aff_pos_50)
 
-Let's now export factor scores.
-
-```{r}
-d_long_imp$aff_pos_fs <- get_fs(cfa_aff_pos)
+# without imputed data
+cfa_aff_pos <- cfa(model, d_long_50, group = "wave", estimator = "MLM")
+d_long_50$aff_pos_fs <- get_fs(cfa_aff_pos)
 ```
@@
 
-Negative affect was measured as a scale. We hence inspect factorial validty using CFA.
-
-We first test assumption of multivariate normality. (Warning, as these analyses take long and might require additional working memory, they're deactived currently.)
-
-```{r eval=FALSE}
-mardia(d_long_imp %>% select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6))
+```{r}
+model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_100_imp)
+summary(model_aff_neg)
+dat_fig_aff_neg <- data.frame(type = "Affect", dimension = "Negative", get_dat(model_aff_neg))
+make_graph(dat_fig_aff_neg, "Negative Affect", 1, 5)
 ```
 
-Assumption of multivariate normal distributino was violated; hence, robust estimator will be used.
+Negative affect was measured as a scale. We hence inspect factorial validity using CFA.
+
+We first test assumption of multivariate normality. We'll focus on wave 1 here.
+
+```{r eval=T}
+d_long_100_imp %>%
+  filter(wave == 1) %>% 
+  select(aff_neg_1, aff_neg_2, aff_neg_3, aff_neg_4, aff_neg_5, aff_neg_6) %>% 
+  mardia()
+```
+
+Assumption of multivariate normal distribution was violated; hence, robust estimator will be used.
 
@@
 "
-cfa_aff_neg <- cfa(model, d_long_imp, group = "wave", estimator = "MLM")
+cfa_aff_neg <- cfa(model, d_long_100_imp, group = "wave", estimator = "MLM")
 summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 rel_aff_neg <- get_rel(cfa_aff_neg)
-rel_aff_neg
 ```
 
-The average reliability across all waves was omega = `r rel_aff_neg`, hence good.
+The average reliability across all waves was omega = `r rel_aff_neg[[1]]`, hence good.
 
 ```{r}
-model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_imp)
+model_aff_neg <- lmer(aff_neg_m ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_aff_neg)
@@
 
-Let's now export factor scores.
+Let's now export factor scores, necessary for results reported in additional analyses.
 
 ```{r}
-d_long_imp$aff_neg_fs <- get_fs(cfa_aff_neg)
+# with imputed data
+cfa_aff_neg_50 <- cfa(model, d_long_50_imp, group = "wave", estimator = "MLM")
+d_long_50_imp$aff_neg_fs <- get_fs(cfa_aff_neg_50)
+
+# without imputed data
+cfa_aff_neg <- cfa(model, d_long_50, group = "wave", estimator = "MLM")
+d_long_50$aff_neg_fs <- get_fs(cfa_aff_neg)
 ```
@@
 ## Social media use
+
+Social media use (and channels) were measured at waves 1, 2, 8, 17, 23, 28, and for everyone who was newly recruited during the study at the first wave.
+
 ### Reading
@@
 ```{r}
-model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_read <- lmer(soc_med_read ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_read)
+dat_fig_soc_med_read <- data.frame(type = "Social media use", dimension = "Reading", 
+                                   get_dat(model_soc_med_read))
+make_graph(dat_fig_soc_med_read, "Social Media Reading", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_like_share <- lmer(soc_med_like_share ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_like_share)
+dat_fig_soc_med_like_share <- data.frame(type = "Social media use", dimension = "Liking & Sharing",
+                                         get_dat(model_soc_med_like_share))
+make_graph(dat_fig_soc_med_like_share, "Social Media Liking and Sharing", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_post <- lmer(soc_med_post ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_post)
+dat_fig_soc_med_post <- data.frame(type = "Social media use", dimension = "Posting", 
+                                   get_dat(model_soc_med_post))
+make_graph(dat_fig_soc_med_post, "Social Media Posting", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_fb <- lmer(soc_med_fb ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_fb)
+dat_fig_soc_med_fb <- data.frame(type = "Social media channel", dimension = "Facebook", 
+                                 get_dat(model_soc_med_fb))
+make_graph(dat_fig_soc_med_fb, "Social Media Facebook", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_tw <- lmer(soc_med_tw ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_tw)
+dat_fig_soc_med_tw <- data.frame(type = "Social media channel", dimension = "Twitter", 
+                                 get_dat(model_soc_med_tw))
+make_graph(dat_fig_soc_med_tw, "Social Media Twitter", 1, 5)
 ```
@@
 ```{r}
-soc_med_tw_m <- d_long_imp %>% group_by(wave) %>% summarise(value = mean(soc_med_tw, na.rm = T))
+soc_med_tw_m <- 
+  d_long_100_imp %>% 
+  group_by(wave) %>% 
+  summarise(value = mean(soc_med_tw, na.rm = T))
 soc_med_tw_m
@@
 ```{r}
-model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_ig <- lmer(soc_med_ig ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_ig)
+dat_fig_soc_med_ig <- data.frame(type = "Social media channel", dimension = "Instagram", 
+                                 get_dat(model_soc_med_ig))
+make_graph(dat_fig_soc_med_ig, "Social Media Instagram", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_wa <- lmer(soc_med_wa ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_wa)
+dat_fig_soc_med_wa <- data.frame(type = "Social media channel", dimension = "WhatsApp", 
+                                 get_dat(model_soc_med_wa))
+make_graph(dat_fig_soc_med_wa, "Social Media WhatsApp", 1, 5)
 ```
@@
 ```{r}
-model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_imp)
+model_soc_med_yt <- lmer(soc_med_yt ~ (1 | id) + (1 | wave), d_long_100_imp)
 summary(model_soc_med_yt)
+dat_fig_soc_med_yt <- data.frame(type = "Social media channel", dimension = "YouTube", 
+                                 get_dat(model_soc_med_yt))
+make_graph(dat_fig_soc_med_yt, "Social Media YouTube", 1, 5)
 ```
@@
 
-The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.
+The only other variable that was measured as a scale was Locus of Control.
+Below I hence report the scale's factorial validity.
+Waves for which not a sufficient number of respondents took part were excluded.
 
-```{r}
+```{r eval=T}
 model <- "
@@
 "
-cfa_loc_cntrl_int <- cfa(model, filter(d_long_imp, wave != 20), group = "wave")
+cfa_loc_cntrl_int <- cfa(model, 
+                         filter(d_long_100_imp, wave != 11, wave != 20,  wave != 26, wave != 27, wave != 29, wave != 31, wave != 32),
+                         # d_long_100_imp,
+                         group = "wave")
 summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
@@
 
-## Figure
+## Table
 
-Display the results in a combined figure.
-
-```{r}
-fig_desc_dat <- data.frame(
-  rbind(
-    data.frame(type = "Life satisfaction", dimension = "Life satisfaction", get_dat(model_life_sat)),
-    data.frame(type = "Affect", dimension = "Positive", get_dat(model_aff_pos)),
-    data.frame(type = "Affect", dimension = "Negative", get_dat(model_aff_neg)),
-    data.frame(type = "Social media use", dimension = "Reading", get_dat(model_soc_med_read)),
-    data.frame(type = "Social media use", dimension = "Liking & Sharing", get_dat(model_soc_med_like_share)),
-    data.frame(type = "Social media use", dimension = "Posting", get_dat(model_soc_med_post)),
-    data.frame(type = "Social media channel", dimension = "Facebook", get_dat(model_soc_med_fb)),
-    data.frame(type = "Social media channel", dimension = "Twitter", get_dat(model_soc_med_tw)),
-    data.frame(type = "Social media channel", dimension = "Instagram", get_dat(model_soc_med_ig)),
-    data.frame(type = "Social media channel", dimension = "WhatsApp", get_dat(model_soc_med_wa)),
-    data.frame(type = "Social media channel", dimension = "YouTube", get_dat(model_soc_med_yt))
-  ) %>% 
-    mutate(
-      type = factor(.$type, levels = c("Life satisfaction", "Affect", "Social media use", "Social media channel")),
-      dimension = factor(.$dimension, levels = c("Life satisfaction", "Positive", "Negative", "Reading", "Liking & Sharing", "Posting", "Facebook", "Twitter", "Instagram", "WhatsApp", "YouTube"))
-      )
-)
-
-fig_desc_life_sat <- make_graph(
-  fig_desc_dat %>% filter(type == "Life satisfaction"), 
-  title = "Life satisfaction", 
-  ll = 0, ul = 10, 
-  lmer = FALSE, 
-  line = TRUE,
-  legend = FALSE
-  )
-
-fig_desc_aff <- make_graph(
-  fig_desc_dat %>% filter(type == "Affect"), 
-  title = "Affect", 
-  ll = 1, ul = 5, 
-  lmer = FALSE, 
-  line = TRUE
-  )
-
-fig_desc_soc_med_use <- make_graph(
-  fig_desc_dat %>% filter(type == "Social media use"), 
-  title = "Social media use", 
-  ll = 1, ul = 5, 
-  lmer = FALSE, 
-  line = TRUE
-  )
-
-fig_desc_soc_med_channel <- make_graph(
-  fig_desc_dat %>% filter(type == "Social media channel"), 
-  title = "Social media channel", 
-  ll = 1, ul = 5, 
-  lmer = FALSE, 
-  line = TRUE
-  )
-
-fig_desc <- grid.arrange(fig_desc_life_sat, fig_desc_aff, 
-                         fig_desc_soc_med_use, fig_desc_soc_med_channel,
-                         nrow = 2, ncol = 2)
-
-ggsave("figures/fig_descriptives.pdf", 
-       width = 10, height = 6,
-       plot = fig_desc)
-```
-
-## Table
+Table with descriptives of main variables.
 
@@
 tab_desc_dat <- rbind(
-    "Life satisfaction" = get_specs(model_life_sat),
+    "Life satisfaction" = get_specs(fit_life_sat),
     "Positive affect" = get_specs(model_aff_pos),
@@
     "YouTube" = get_specs(model_soc_med_yt)
+    ) %>%
+  as.data.frame()
+tab_desc_dat
+```
+
+## Figure
+
+Display the developments of all variables in a combined figure.
+
+```{r}
+fig_desc_dat <- data.frame(
+  rbind(
+    dat_fig_life_sat,
+    dat_fig_aff_pos,
+    dat_fig_aff_neg,
+    dat_fig_soc_med_read,
+    dat_fig_soc_med_like_share,
+    dat_fig_soc_med_post,
+    dat_fig_soc_med_fb,
+    dat_fig_soc_med_tw,
+    dat_fig_soc_med_ig,
+    dat_fig_soc_med_wa,
+    dat_fig_soc_med_yt
+  ) %>% 
+    mutate(
+      type = factor(.$type, levels = c("Life satisfaction", "Affect", "Social media use", "Social media channel")),
+      dimension = factor(.$dimension, levels = c("Life satisfaction", "Positive", "Negative", "Reading", "Liking & Sharing", "Posting", "Facebook", "Twitter", "Instagram", "WhatsApp", "YouTube"))
       )
-kable(tab_desc_dat)
+)
+
+fig_desc_life_sat <- make_graph(
+  fig_desc_dat %>% filter(type == "Life satisfaction"), 
+  title = "Life satisfaction", 
+  ll = 0, ul = 10, 
+  lmer = FALSE, 
+  line = TRUE,
+  legend = FALSE,
+  points = FALSE
+  )
+
+fig_desc_aff <- make_graph(
+  fig_desc_dat %>% filter(type == "Affect"), 
+  title = "Affect", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE,
+  points = FALSE,
+  legend = TRUE
+  )
+
+fig_desc_soc_med_use <- make_graph(
+  fig_desc_dat %>% filter(type == "Social media use"), 
+  title = "Social media use", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE,
+  points = FALSE,
+  legend = TRUE
+  )
+
+fig_desc_soc_med_channel <- make_graph(
+  fig_desc_dat %>% filter(type == "Social media channel"), 
+  title = "Social media channel", 
+  ll = 1, ul = 5, 
+  lmer = FALSE, 
+  line = TRUE,
+  points = FALSE,
+  legend = TRUE
+  )
+
+fig_desc <- grid.arrange(fig_desc_life_sat, fig_desc_aff, 
+                         fig_desc_soc_med_use, fig_desc_soc_med_channel,
+                         nrow = 2, ncol = 2)
+
+ggsave("figures/fig_descriptives.png", 
+       width = 10, height = 5,
+       plot = fig_desc)
 ```
@@
 # Analyses
-## Life satisfaction
 
-First, without control variables.
+## Multicollinearity
+
+Before running the analyses, let's briefly check zero-order correlation matrix, to get general picture and also for potential multicollinearity. We use variables from T1.
+
+Multicollinearity will then also be checked explicitly in each analysis.
+
+```{r eval=T}
+d_long_100_imp %>% 
+  filter(wave == 1) %>% 
+  select(life_sat, aff_neg_m, aff_pos_m,
+         soc_med_read, soc_med_post, soc_med_like_share, 
+         soc_med_fb, soc_med_ig, soc_med_tw, soc_med_wa, soc_med_yt,
+         health, corona_pos, work_h, work_homeoff, hh_income, med_txt_kro, med_txt_sta, med_txt_pre, med_txt_oes, med_txt_kur, med_txt_slz, med_txt_son, med_vid_orf, med_vid_pri, med_txt_kro, med_txt_sta, med_txt_pre, med_txt_oes, med_txt_kur, med_txt_slz, med_txt_son, med_vid_orf, med_vid_pri, act_wrk, act_spo, act_frn, act_sho, act_pet, risk_prop, loc_cntrl_int_m, sat_dem) %>%
+  cor(use = "pairwise.complete.obs") %>% 
+  as.data.frame()
+```
+
+## Correlations
+
+Let's also briefly look at bivariate relations between the types and channels of social media use and the well-being facets.
+
+### Table
 
 ```{r}
-model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
-                            (1 | id) + (1 | wave) + 
-                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
-                data = d_long_imp)
-summary(model_life_sat_lmer_nco)
+dat_cor <- 
+  d_long_100_mim %>% 
+  filter(wave == 1) %>%  # we can use wave 1 only, because mean values are the same across waves
+  select(`Life satis-\nfaction` = life_sat_b, 
+         `Affect\npositive` = aff_pos_m_b, 
+         `Affect\nnegative` = aff_neg_m_b,
+         `Reading` = soc_med_read_b, 
+         `Posting` = soc_med_post_b, 
+         `Like &\nshare` = soc_med_like_share_b, 
+         `Facebook` = soc_med_fb_b, 
+         `Instagram` = soc_med_ig_b, 
+         `Twitter` = soc_med_tw_b, 
+         `WhatsApp` = soc_med_wa_b, 
+         `YouTube` = soc_med_yt_b)
+  # cor() %>% 
+  # as.data.frame()
+  # correlate() %>% 
+  # fashion()
+
+tab_cor <- 
+  dat_cor %>% 
+  correlate() %>% 
+  fashion() %T>%
+  print()
 ```
 
-Now with control variables.
+### Figure
 
 ```{r}
-model_life_sat_lmer <- lmerTest::lmer(life_sat ~
+int_breaks <- function(x, n = 4) {
+  l <- pretty(x, n)
+  l[abs(l %% 1) < .Machine$double.eps ^ 0.5] 
+}
+
+fig_cor <- 
+  dat_cor %>% 
+  ggpairs(
+    upper = list(continuous = cor_plot),
+    lower = list(continuous = wrap("points", alpha = 0.3, size=0.1), 
+              combo = wrap("dot", alpha = 0.3, size=0.1)),
+    progress = FALSE
+    ) + 
+  scale_x_continuous(breaks = int_breaks) +
+  theme_bw()
+
+fig_cor
+ggsave("figures/fig_cor.png", width = 8, height = 8)
+```
+
+## Publication
+### Life satisfaction
+
+```{r}
+model_life_sat_lmer_pub <- "
+  life_sat ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        # hh_income_b + # large dropout (900 obs)
-                                        # hh_adults + # too much dropout
-                                        # hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + hh_oldfam + hh_outfam + hh_partner + 
-                                        # home_owner + # too much dropout
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_life_sat_lmer)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
 ```
 
-## Positive Affect
-
-First, without control variables.
+Let's first inspect multicollinearity.
 
 ```{r}
-model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_nco)
+check_collinearity(lmerTest::lmer(model_life_sat_lmer_pub, d_long_100_imp))
 ```
 
-Now with control variables.
+No within-person predictor shows increased values for multicollinearity. Hence results straightforward.
+
+Let's next inspect results of within-person predictors. Only parts of model shown to save space.
 
 ```{r}
-model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
+fit_life_sat_lmer_pub <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_life_sat_lmer_pub))
+fit_life_sat_lmer_pub <- summary(pool(fit_life_sat_lmer_pub), conf.int = TRUE)
+print_res(fit_life_sat_lmer_pub)
+```
+
+### Positive Affect
+
+```{r}
+model_aff_pos_lmer_pub <- "
+  aff_pos_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
 ```
 
-## Negative Affect
-
-First, without control variables.
+Let's first inspect multicollinearity.
 
 ```{r}
-model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_nco)
+check_collinearity(lmerTest::lmer(model_aff_pos_lmer_pub, d_long_100_imp))
 ```
 
-Then, with control variables.
+No within-person predictors show multicollinear relations.
+
+In what follows, the results of within-person predictors.
 
 ```{r}
-model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
+fit_aff_pos_lmer_pub <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_pos_lmer_pub))
+fit_aff_pos_lmer_pub <- summary(pool(fit_aff_pos_lmer_pub), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_pub)
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_pub <- "
+  aff_neg_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+"
 ```
 
-# Figures
-
-First, for activities.
+Let's inspect multicollinearity.
 
 ```{r}
-# get data
-dat_fig_results_activity <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "activity")
-dat_fig_results_activity_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "activity")
-
-# make figure
-fig_results_activity <- make_graph_res(
-  data_1 = dat_fig_results_activity,
-  data_2 = dat_fig_results_activity_nco,
-  sesoi = "est"
-  )
-fig_results_activity
-
-# safe figure
-ggsave("figures/fig_results_activity.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_activity)
+check_collinearity(lmerTest::lmer(model_aff_neg_lmer_pub, d_long_100_imp))
 ```
 
-Then for channels.
+No within-person predictors show multicollinear relations.
+
+Here are the results for the within-person predictors.
 
 ```{r}
-# get data
-dat_fig_results_channel <- get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "channel")
-dat_fig_results_channel_nco <- get_dat_res(model_aff_neg_lmer_nco, model_aff_pos_lmer_nco, model_life_sat_lmer_nco, type = "channel")
+fit_aff_neg_lmer_pub <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_neg_lmer_pub))
+fit_aff_neg_lmer_pub <- summary(pool(fit_aff_neg_lmer_pub), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_pub)
+```
 
-# make figure
-fig_results_channel <- make_graph_res(
-  data_1 = dat_fig_results_channel,
-  data_2 = dat_fig_results_channel_nco,
-  sesoi = "est"
+## Standardized results
+
+Then let's report also the standardized results. Helps compare effect sizes across differently scaled predictors.
+
+### Life satisfaction
+
+```{r}
+fit_life_sat_lmer_std <- with(d_long_100_mim_mice_std, exp = lmerTest::lmer(model_life_sat_lmer_pub))
+fit_life_sat_lmer_std <- summary(pool(fit_life_sat_lmer_std), conf.int = TRUE)
+print_res(fit_life_sat_lmer_std)
+```
+
+### Positive Affect
+
+```{r}
+fit_aff_pos_lmer_std <- with(d_long_100_mim_mice_std, exp = lmerTest::lmer(model_aff_pos_lmer_pub))
+fit_aff_pos_lmer_std <- summary(pool(fit_aff_pos_lmer_std), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_std)
+```
+
+### Negative Affect
+
+```{r}
+fit_aff_neg_lmer_std <- with(d_long_100_mim_mice_std, exp = lmerTest::lmer(model_aff_neg_lmer_pub))
+fit_aff_neg_lmer_std <- summary(pool(fit_aff_neg_lmer_std), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_std)
+```
+
+## Table
+
+Let's extract results for a table of within-person effects.
+
+```{r cache=F, eval=T}
+tab_within <-
+  data_tab_std %>% 
+  filter(Type %in% c("Activity", "Channels")) %>% 
+  select(std = estimate) %>% 
+  cbind(data_tab_within) %>% 
+  mutate(p.value = td::my_round(p.value, "p")) %>% 
+  arrange(dv) %>% 
+  select(Outcome = dv,
+         Predictor = iv,
+         b = estimate,
+         `Lower` = conf.low,
+         `Higher` = conf.high,
+         beta = "std",
+         p = "p.value")
+
+tab_within
+```
+
+## Figures
+
+Let's visualize results. First, results of unstandardized predictors.
+
+```{r warning=F, cache=F}
+# get dat
+data_tab_within <- rbind(
+  get_dat_res(fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+              type = "Activity", variance = "within", analysis = "Publication"),
+  get_dat_res(fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+              type = "Channels", variance = "within", analysis = "Publication")
   )
-fig_results_channel
+
+# make fig
+fig_results_within <- make_graph_res(
+  data = data_tab_within,
+  sesoi = "est",
+  legend = FALSE
+  , facet = "type"
+  # , title = "Results of selected covariates"
+  )
+fig_results_within
 
 # save figure
-ggsave("figures/fig_results_channel.pdf", 
+ggsave("figures/fig_results_within.png", 
        width = 7, height = 4,
-       plot = fig_results_channel)
+       plot = fig_results_within)
 ```
 
-Select control variables and make graph.
+Let's next visualize standardized predictors, to allow for better comparison across differently scaled variables.
 
-```{r}
+```{r warning=F, cache=F, fig.height=7}
 # make figure
-fig_results_control <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer, model_aff_pos_lmer, model_life_sat_lmer, type = "control")
+data_tab_comp <- rbind(
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Activity", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Channels", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "News\nuse", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Living\nconditions", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Outdoor\nactivities", variance = "within", analysis = "standardized"),
+  get_dat_res(fit_aff_neg_lmer_std, fit_aff_pos_lmer_std, fit_life_sat_lmer_std, 
+              type = "Psycho-\nlogy", variance = "within", analysis = "standardized")
   )
-fig_results_control
+
+fig_results_comp <- make_graph_res(
+  data = data_tab_comp,
+  sesoi = "std",
+  legend = FALSE
+  , facet = "type"
+  # , title = "Results of selected covariates"
+  )
+fig_results_comp
 
 # save figure
-ggsave("figures/fig_results_control.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_control)
+ggsave("figures/fig_results_comp.png", 
+       width = 7, height = 7,
+       plot = fig_results_control_std)
 ```
 
-```{r}
-save.image("data/workspace_1.RData")
+# Save workspace
+
+```{r cache=F}
+save.image("data/workspace_2.RData")
 ```

### 2023-01-30 analyses_additional.Rmd
- major updates to preprint

--- 2021-11-30 analyses_additional.Rmd
@@
     toc_depth: 2
-    # code_folding: hide
+    code_folding: hide
+    df_print: paged
 bibliography      : 
@@
 set.seed(170819)
-knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T)
-options(digits = 3, width = 200, scipen=999)
+knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache=T, max.row=20)
+options(digits=2, width=260, scipen=999)
 ```
+
+ - Here you can find the code and the results of all additional analyses. 
+ - To see a good overview of the results, I recommend you directly go down to the Figures section. 
+ - But if you're interested in how the analyses were executed, see below.
+ - To see the code, click on button "Code". Alternatively, you can download the rmd file from the github repo.
 
@@
 
-Load packages, and make repo replicable.
+Load packages.
 
-```{r r-setup, results='hide'}
+```{r r-setup, results='hide', cache=F}
 # install packages
@@
 # define packages
-packages <- c("brms", "devtools", "GGally", "ggplot2", 
+packages <- c("broom.mixed", "brms", "devtools", "GGally", "ggplot2", 
               "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
@@
 # load workspace
-load("data/workspace_1.RData")
+load("data/workspace_2.RData")
 ```
 
-# Standardized results
+# With additional covariates
 
-First, create standardized dataset.
+These models include additional covariates that weren't preregistered and weren't included for theoretical reasons. These are:
+
+- Trust in ORF
+- Trust in police
+- Trust in parliament
+- Trust in health sector
+- Trust in government
+- Trust in army
+
+### Life satisfaction
 
 ```{r}
-d_long_std <- d_long_imp %>% 
-  mutate(across(c(-id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), scale))
-```
-
-Then let's rerun the analyses.
-
-## Life satisfaction
-
-```{r}
-model_life_sat_lmer_std <- lmerTest::lmer(life_sat ~
+model_life_sat_lmer_add <- "
+  life_sat ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_std)
-summary(model_life_sat_lmer_std)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b +
+    trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w + 
+    trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+"
+
+fit_life_sat_lmer_add <- with(d_long_100_mim_mice, 
+                              exp = lmerTest::lmer(model_life_sat_lmer_add))
+fit_life_sat_lmer_add <- summary(pool(fit_life_sat_lmer_add), conf.int = TRUE)
+print_res(fit_life_sat_lmer_add)
 ```
 
-## Positive affect
+### Positive Affect
 
 ```{r}
-model_aff_pos_lmer_std <- lmerTest::lmer(aff_pos_fs ~ 
+model_aff_pos_lmer_add <- "
+  aff_pos_m ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_std)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b +
+    trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w + 
+    trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+"
+
+fit_aff_pos_lmer_add <- with(d_long_100_mim_mice, 
+                              exp = lmerTest::lmer(model_aff_pos_lmer_add))
+fit_aff_pos_lmer_add <- summary(pool(fit_aff_pos_lmer_add), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_add)
+```
+
+### Negative Affect
+
+```{r}
+model_aff_neg_lmer_add <- "
+  aff_neg_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+    res_vienna + acc_bal + acc_gar + home_sqm + 
+    corona_pos_b + corona_pos_w +
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
+    med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
+    med_vid_orf_b + med_vid_pri_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
+    act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
+    act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b +
+    trst_media_w + trst_police_w + trst_media_w + trst_hlthsec_w + trst_gov_w + trst_army_w + 
+    trst_media_b + trst_police_b + trst_media_b + trst_hlthsec_b + trst_gov_b + trst_army_b
+"
+
+fit_aff_neg_lmer_add <- with(d_long_100_mim_mice, 
+                              exp = lmerTest::lmer(model_aff_neg_lmer_add))
+fit_aff_neg_lmer_add <- summary(pool(fit_aff_neg_lmer_add), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_add)
+```
+
+# Without control variables
+
+Let's inspect how results change when control variables are omitted.
+
+## Life satisfaction
+
+```{r}
+model_life_sat_lmer_nco <- "
+  life_sat ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+"
+fit_life_sat_lmer_nco <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_life_sat_lmer_nco))
+fit_life_sat_lmer_nco <- summary(pool(fit_life_sat_lmer_nco), conf.int = TRUE)
+print_res(fit_life_sat_lmer_nco)
+```
+
+## Positive Affect
+
+```{r}
+model_aff_pos_lmer_nco <- "
+  aff_pos_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+"
+fit_aff_pos_lmer_nco <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_pos_lmer_nco))
+fit_aff_pos_lmer_nco <- summary(pool(fit_aff_pos_lmer_nco), conf.int = TRUE)
+print_res(fit_aff_pos_lmer_nco)
 ```
@@
 ```{r}
-model_aff_neg_lmer_std <- lmerTest::lmer(aff_neg_fs ~ 
+model_aff_neg_lmer_nco <- "
+  aff_neg_m ~ 
+    (1 | id) + (1 | wave) + 
+    soc_med_read_w + soc_med_like_share_w + soc_med_post_w  + 
+    soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
+    soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
+    soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
+"
+fit_aff_neg_lmer_nco <- with(d_long_100_mim_mice, exp = lmerTest::lmer(model_aff_neg_lmer_nco))
+fit_aff_neg_lmer_nco <- summary(pool(fit_aff_neg_lmer_nco), conf.int = TRUE)
+print_res(fit_aff_neg_lmer_nco)
+```
+
+# Preregistered
+
+Some changes were introduced because of feedback from colleagues (multiple imputation, imputation of all participants, inclusion of all variables).
+In what follows, please find the results as originally planned (but with some necessary deviations; for example not all preregistered variables could be included).
+
+ - Use data-set in which participants with >50% data were removed. 
+ - Also use factor scores (possible without multiple imputation).
+
+Note that because here no multiple imputation was used, output looks differently.
+
+## Life satisfaction
+
+```{r}
+model_life_sat_lmer_pre <- "
+  life_sat ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_std)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+    "
+fit_life_sat_lmer_pre <- lmerTest::lmer(model_life_sat_lmer_pre, data = d_long_50_imp)
+print_res(broom.mixed::tidy(fit_life_sat_lmer_pre, conf.int = T), imputation = FALSE)
 ```
 
-## Figures
-
-First, for activities.
+## Positive Affect
 
 ```{r}
-# make figure
-fig_results_activity_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "activity"),
-  sesoi = "std"
-  )
-fig_results_activity_std
-
-# safe figure
-ggsave("figures/fig_results_activity_std.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_activity_std)
-```
-
-Then for channels.
-
-```{r}
-# make figure
-fig_results_channel_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "channel"),
-  sesoi = "std"
-  )
-fig_results_channel_std
-
-# save figure
-ggsave("figures/fig_results_channel_std.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_channel_std)
-```
-
-Select control variables and make graph.
-
-```{r}
-# make figure
-fig_results_control_std <- make_graph_res(
-  get_dat_res(model_aff_neg_lmer_std, model_aff_pos_lmer_std, model_life_sat_lmer_std, type = "control"),
-  sesoi = "std"
-  )
-fig_results_control_std
-
-# save figure
-ggsave("figures/fig_results_control_std.pdf", 
-       width = 7, height = 4,
-       plot = fig_results_control_std)
-```
-
-# Results with mean scores
-
-Instead of using factor scores, in what follows I report the results with mean scores of affect. Note that the results for life satisfaction were the same, as it was measured with a single item.
-
-## Positive affect
-
-```{r}
-model_aff_pos_lmer_m <- lmerTest::lmer(aff_pos_m ~ 
+model_aff_pos_lmer_pre <- "
+  aff_pos_fs ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_pos_lmer_std)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+    "
+fit_aff_pos_lmer_pre <- lmerTest::lmer(model_aff_pos_lmer_pre, data = d_long_50_imp)
+print_res(broom.mixed::tidy(fit_aff_pos_lmer_pre, conf.int = T), imputation = FALSE)
 ```
-
-The results differed only slightly, and all inferences remained the same.
 
@@
 ```{r}
-model_aff_neg_lmer_m <- lmerTest::lmer(aff_neg_m ~ 
+model_aff_neg_lmer_pre <- "
+  aff_neg_fs ~ 
     (1 | id) + (1 | wave) + 
@@
     soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
+    age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
     res_vienna + acc_bal + acc_gar + home_sqm + 
+    work_h_b + work_h_w +
+    work_homeoff_b +  work_homeoff_w +
+    hh_income_b + hh_income_w +
+    hh_adults + hh_child18 + hh_child17 + hh_child14 + hh_child5 + hh_child2 + 
+    hh_oldfam + hh_outfam + hh_partner +
+    home_owner +
     med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
     med_vid_orf_b + med_vid_pri_b + 
-                                        risk_prop_b + loc_cntrl_int_m_b + 
+    med_txt_kro_w + med_txt_sta_w + med_txt_pre_w + med_txt_oes_w + med_txt_kur_w + med_txt_slz_w + med_txt_son_w + 
+    med_vid_orf_w + med_vid_pri_w + 
+    risk_prop_b + risk_prop_w + 
     act_wrk_w + act_spo_w + act_frn_w + act_sho_w + act_pet_w + 
     act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long_imp)
-summary(model_aff_neg_lmer_std)
+    sat_dem_w + sat_dem_b + 
+    health_w + health_b +
+    loc_cntrl_int_m_w + loc_cntrl_int_m_b
+    "
+fit_aff_neg_lmer_pre <- lmerTest::lmer(model_aff_neg_lmer_pre, data = d_long_50_imp)
+print_res(broom.mixed::tidy(fit_aff_neg_lmer_pre, conf.int = T), imputation = FALSE)
 ```
 
-The results differed only slightly, and all inferences remained the same.
+# Preregistered without imputation
 
-# Results without imputed data
-
-We first need to export factor scores for variables without imputed data
-
-```{r}
-model <- "
-aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
-"
-cfa_aff_pos <- cfa(model, d_long, group = "wave", estimator = "MLM")
-d_long$aff_pos_fs <- get_fs(cfa_aff_pos)
-
-model <- "
-aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
-"
-cfa_aff_neg <- cfa(model, d_long, group = "wave", estimator = "MLM")
-d_long$aff_neg_fs <- get_fs(cfa_aff_neg)
-```
+Same as before, but now don't use any imputed data whatsoever. Was preregistered as additional analysis to provide comparison.
 
@@
 
-First, without control variables.
-
 ```{r}
-model_life_sat_lmer_nco <- lmerTest::lmer(life_sat ~ 
-                            (1 | id) + (1 | wave) + 
-                            soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                            soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                            soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                            soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b,
-                data = d_long)
-summary(model_life_sat_lmer_nco)
-```
-
-Now with control variables.
-
-```{r}
-model_life_sat_lmer <- lmerTest::lmer(life_sat ~
+model_life_sat_lmer_noi <- "
+  life_sat ~
   (1 | id) + (1 | wave) + 
@@
   soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b +  
-                                        age + male + born_aus + born_aus_prnts + edu_fac + employment_fac + health_b + 
-                                        res_vienna + acc_bal + acc_gar + home_sqm + 
-                                        med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + 
-                                        med_vid_orf_b + med_vid_pri_b + 
+  age + male + born_aus + born_aus_prnts + edu_fac + employment_fac +
+  res_vienna + acc_bal + acc_gar + home_sqm + corona_pos +
+  med_txt_kro_b + med_txt_sta_b + med_txt_pre_b + med_txt_oes_b + med_txt_kur_b + med_txt_slz_b + med_txt_son_b + med_vid_orf_b + med_vid_pri_b + 
   risk_prop_b + loc_cntrl_int_m_b + 
@@
   act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
-summary(model_life_sat_lmer)
+  sat_dem_w + sat_dem_b + 
+  health_w + health_b
+"
+fit_life_sat_lmer_noi <- lmerTest::lmer(model_life_sat_lmer_noi, data = d_long_50)
+print_res(broom.mixed::tidy(fit_life_sat_lmer_noi, conf.int = T), imputation = FALSE)
 ```
@@
 
-First, without control variables.
-
 ```{r}
-model_aff_pos_lmer_nco <- lmerTest::lmer(aff_pos_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long)
-summary(model_aff_pos_lmer_nco)
-```
-
-Now with control variables.
-
-```{r}
-model_aff_pos_lmer <- lmerTest::lmer(aff_pos_fs ~ 
+model_aff_pos_lmer_noi <- "
+  aff_pos_fs ~ 
   (1 | id) + (1 | wave) + 
@@
   act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
-summary(model_aff_pos_lmer)
+  sat_dem_w + sat_dem_b + 
+  health_w + health_b
+"
+fit_aff_pos_lmer_noi <- lmerTest::lmer(model_aff_pos_lmer_noi, data = d_long_50)
+print_res(broom.mixed::tidy(fit_aff_pos_lmer_noi, conf.int = T), imputation = FALSE)
 ```
@@
 
-First, without control variables.
-
 ```{r}
-model_aff_neg_lmer_nco <- lmerTest::lmer(aff_neg_fs ~ 
-                                        (1 | id) + (1 | wave) + 
-                                        soc_med_read_w + soc_med_like_share_w + soc_med_post_w + 
-                                        soc_med_fb_w + soc_med_ig_w + soc_med_wa_w + soc_med_yt_w + soc_med_tw_w +
-                                        soc_med_read_b + soc_med_like_share_b + soc_med_post_b + 
-                                        soc_med_fb_b + soc_med_ig_b + soc_med_wa_b + soc_med_yt_b + soc_med_tw_b
-                                      , data = d_long)
-summary(model_aff_neg_lmer_nco)
-```
-
-Then, with control variables.
-
-```{r}
-model_aff_neg_lmer <- lmerTest::lmer(aff_neg_fs ~ 
+model_aff_neg_lmer_noi <- "
+  aff_neg_fs ~ 
   (1 | id) + (1 | wave) + 
@@
   act_wrk_b + act_spo_b + act_frn_b + act_sho_b + act_pet_b +
-                                        sat_dem_w + sat_dem_b 
-                                      , data = d_long)
-summary(model_aff_neg_lmer)
+  sat_dem_w + sat_dem_b + 
+  health_w + health_b
+"
+fit_aff_neg_lmer_noi <- lmerTest::lmer(model_aff_neg_lmer_noi, data = d_long_50)
+print_res(broom.mixed::tidy(fit_aff_neg_lmer_noi, conf.int = T), imputation = FALSE)
 ```
 
-```{r}
+# Figures
+
+ - In what follows, see figure with all results combined. 
+ - Results show that different models are comparable. 
+ - The model without control variables stands out a bit, showing that controlling for additional varying variables is important.
+ - The preregistered model without imputation also has larger confidence intervals, which was to be expected given fewer data.
+
+```{r warning=F, fig.height=6}
+# get data
+dat_fig_results_activity_pub <- get_dat_res(
+  fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+  type = "Activity", analysis = "1. Published", variance = "within")
+
+dat_fig_results_channel_pub <- get_dat_res(
+  fit_aff_neg_lmer_pub, fit_aff_pos_lmer_pub, fit_life_sat_lmer_pub, 
+  type = "Channels", analysis = "1. Published", variance = "within")
+
+dat_fig_results_activity_add <- get_dat_res(
+  fit_aff_neg_lmer_add, fit_aff_pos_lmer_add, fit_life_sat_lmer_add, 
+  type = "Activity", analysis = "2. Additional covars", variance = "within")
+
+dat_fig_results_channel_add <- get_dat_res(
+  fit_aff_neg_lmer_add, fit_aff_pos_lmer_add, fit_life_sat_lmer_add, 
+  type = "Channels", analysis = "2. Additional covars", variance = "within")
+
+dat_fig_results_activity_nco <- get_dat_res(
+  fit_aff_neg_lmer_nco, fit_aff_pos_lmer_nco, fit_life_sat_lmer_nco, 
+  type = "Activity", analysis = "3. No covars", variance = "within")
+
+dat_fig_results_channel_nco <- get_dat_res(
+  fit_aff_neg_lmer_nco, fit_aff_pos_lmer_nco, fit_life_sat_lmer_nco, 
+  type = "Channels", analysis = "3. No covars", variance = "within")
+
+dat_fig_results_activity_pre <- get_dat_res(
+  fit_aff_neg_lmer_pre, fit_aff_pos_lmer_pre, fit_life_sat_lmer_pre, 
+  type = "Activity", analysis = "4. Preregistered", variance = "within")
+
+dat_fig_results_channel_pre <- get_dat_res(
+  fit_aff_neg_lmer_pre, fit_aff_pos_lmer_pre, fit_life_sat_lmer_pre, 
+  type = "Channels", analysis = "4. Preregistered", variance = "within")
+
+dat_fig_results_activity_noi <- get_dat_res(
+  fit_aff_neg_lmer_noi, fit_aff_pos_lmer_noi, fit_life_sat_lmer_noi, 
+  type = "Activity", analysis = "5. No imputation", variance = "within")
+
+dat_fig_results_channel_noi <- get_dat_res(
+  fit_aff_neg_lmer_noi, fit_aff_pos_lmer_noi, fit_life_sat_lmer_noi, 
+  type = "Channels", analysis = "5. No imputation", variance = "within")
+
+dat_fig_results_comp <- rbind(
+  dat_fig_results_activity_pub, 
+  dat_fig_results_channel_pub,
+  dat_fig_results_activity_nco, 
+  dat_fig_results_channel_nco,
+  dat_fig_results_activity_add, 
+  dat_fig_results_channel_add,
+  dat_fig_results_activity_pre, 
+  dat_fig_results_channel_pre,
+  dat_fig_results_activity_noi, 
+  dat_fig_results_channel_noi
+)
+
+# make figure
+fig_results_comp <- make_graph_res(
+  dat_fig_results_comp,
+  sesoi = "est",
+  facet = "type"
+  )
+fig_results_comp
+
+# safe figure
+ggsave("figures/fig_results_comparison.pdf", 
+       width = 7, height = 5,
+       plot = fig_results_within_comp)
+```
+
+```{r eval=T}
 save.image("data/workspace_2.RData")

### 2023-01-30 analyses_setup.Rmd
- major updates to preprint

---
title: "Analyses Setup"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
    df_print: paged
bibliography      : 
  - "bibliography/r-references.bib"
  - "bibliography/references.bib"
---

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache=T, rows.print=20)
options(digits=2, width=260, scipen=999)
```

- Here you can find the set-up for the analyses and the results.
- Includes custom functions, data-wrangling, and data imputation. 
- To see the code, click on button "Code".
- Alternatively, you can download the rmd file from the github repo.
- The data are hosted on aussda. Please download them separately, to rerun the analyses (https://doi.org/10.11587/28KQNS). 
- Note that this paper uses a scientific use file, which is at the time of writing not officially published.

# Set-up

Load packages.

```{r r-setup, results='hide', cache=F}
# install packages
# devtools::install_github("https://github.com/tdienlin/td@v.0.0.2.5")

# define packages
packages <- c("broom.mixed", "brms", "devtools", "GGally", "ggplot2", 
              "gridExtra", "kableExtra", "knitr", "lavaan", "lme4", 
              "magrittr", "mice", #"mvnormalTest", 
              "PerFit", "performance", "psych", "quanteda.textstats", "semTools", "tidyverse")

# load packages
lapply(c(packages, "td"), library, character.only = TRUE)
```

## Session info

Provide session info to make results reproducible.

```{r}
sessionInfo()
```

# Custom functions

## find_var

Find variables in dataset.

```{r}
find_var <- function(name, data = doc)(
  # finds the variables names for an item for each wave
  data %>%
    filter(Label == name) %>%
    select(Variable) %>%
    unlist() %>%
    set_names(sub("\\_.*", "", .))
)
```

## get_specs

Extract characteristics from fitted lmer models.

```{r}
get_specs <- function(model){
  # Get mean, max, and min values

  dat <- coefficients(model)$wave
  specs <- data.frame(
    sd = attr(VarCorr(model), "sc"),
    min = min(dat),
    max = max(dat),
    mean = mean(dat$`(Intercept)`)
  )
}
```

## get_dat

Get data from fitted lmer objects for descriptives.

```{r}
get_dat <- function(model){
  coefficients(model)$wave %>% 
    tibble::rownames_to_column("wave") %>% 
    rename(value = "(Intercept)") %>% 
    mutate(wave = as.integer(.$wave),
           value = as.numeric(.$value)) 
} 
```

## get_rel

Determine average reliability for measures across waves

```{r}
get_rel <- function(data, waves=24){

  # extract average reliability from lavaan fitted cfa with several groups
  reliability(data) %>% 
    unlist() %>% 
    matrix(5, waves) %>% 
    as.data.frame() %>% 
    set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
    summarise(omega = rowMeans(.["omega",])) %>% 
    return()
}
```

## make_graph

Make graph of variables' development.

```{r}
make_graph <- function(model, title, ll, ul, line = TRUE, points = TRUE, labels = FALSE, lmer=FALSE, legend=FALSE){

    if(isTRUE(lmer)){
      dat <- get_dat(model)
    } else{
    dat <- model
  }

  graph <- 
    ggplot(dat, aes(wave, value, color = dimension)) +
    {if(legend) theme(
      legend.position = c(0.85, 0.8),
      legend.title = element_blank()
      )} + 
    {if(!legend) theme(
      legend.position = "none"
    )} +
    theme(axis.title.x=element_blank()) +
    coord_cartesian(ylim = c(ll, ul)) +
    ggtitle(title) +
    {if(line) geom_smooth(se = FALSE, method = 'loess')} +
    {if(points) geom_point()} +
    scale_color_manual(values=c("dodgerblue3", "deepskyblue2", "magenta2", "green2", "red"))
  graph
}
```

## print_res

```{r}
# print results for online material
print_res <- function(object, imputation = TRUE){

  if(isTRUE(imputation)){
    object %>% 
      as.data.frame %>% 
      select(term, estimate, `2.5 %`, `97.5 %`, p.value) %>% 
      mutate(p.value = td::my_round(p.value, "p"))
    }
  else{
    object %>% 
      as.data.frame %>% 
      select(term, estimate, `2.5 %` = conf.low, `97.5 %` = conf.high, p.value) %>% 
      mutate(p.value = td::my_round(p.value, "p"))
  }
}
```

## get_dat_res

Get data of lmer objects for specific results.

```{r}
# get data
get_dat_res <- function(data_aff_neg, data_aff_pos, data_life_sat, variance = "within", type = "channels", analysis = "Standardized"){

  if(isTRUE(class(data_aff_neg) == "lmerModLmerTest")) {
         dat_fig_results <- 
           broom.mixed::tidy(data_aff_neg, conf.int = T) %>% 
           mutate(dv = "aff_neg") %>% 
           rbind(
             broom.mixed::tidy(data_aff_pos, conf.int = T) %>% 
               mutate(dv = "aff_pos")) %>% 
           rbind(
             broom.mixed::tidy(data_life_sat, conf.int = T) %>% 
               mutate(dv = "life_sat"))
         } else{ 
           dat_fig_results <- 
             data_aff_neg %>% 
             mutate(dv = "aff_neg") %>% 
             rbind(data_aff_pos %>% 
                   mutate(dv = "aff_pos")) %>% 
             rbind(data_life_sat %>% 
                     mutate(dv = "life_sat")) %>% 
           rename(conf.low = `2.5 %`, conf.high = `97.5 %`)
           }
  dat_fig_results %<>% 
  mutate(
    Variance = ifelse(grepl(".*_w\\>", .$term), "within", "between"),
    iv = gsub("_(w|b)\\>", "", .$term)
         ) %>%
  mutate(
    Variance = factor(.$Variance, 
                  levels = c("within", "between")),
    dv = factor(.$dv, 
                levels = c("life_sat", "aff_pos", "aff_neg"),
                labels = c("Life satisfaction", "Positive affect", "Negative affect")),
    Type = type,
    Analysis = analysis
  ) %>% 
  select(dv, iv, Variance, Type, Analysis, estimate, conf.low, conf.high, p.value) %>% 
    filter(Variance == variance)

  # select Social Media type of activity
  if(type == "Activity") {
    dat_fig_results %<>%
    filter(iv %in% c("soc_med_read", "soc_med_like_share", "soc_med_post")) %>% 
    mutate(
      iv = factor(.$iv, 
                  levels = c("soc_med_post", "soc_med_like_share", "soc_med_read"), 
                  labels = c("Posting", "Liking & Sharing", "Reading"))
    )
  } else if(type == "Channels"){
    dat_fig_results %<>%
    filter(iv %in% c("soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")) %>% 
    mutate(iv = factor(.$iv, levels = c("soc_med_yt", "soc_med_wa", "soc_med_ig", "soc_med_tw", "soc_med_fb"), labels = c("YouTube", "WhatsApp", "Instagram", "Twitter", "Facebook")))
  } else if(type == "Controls"){
    dat_fig_results %<>%
      filter(iv %in% c("male", "health", "loc_cntrl_int_m",
                       "act_spo", "sat_dem", "corona_pos")) %>% 
      mutate(iv = factor(.$iv, levels = c("male", "health", "loc_cntrl_int_m",
                                          "act_spo", "sat_dem", "corona_pos"), 
                         labels = c("Male", "Health", "Internal locus of control",
                                    "Sport", "Satisfaction democracy", "Corona positive")))
  } else if(type == "Living\nconditions"){
    dat_fig_results %<>%
      filter(iv %in% c("health", "corona_pos", "work_h", "work_homeoff", "hh_income")) %>% 
      mutate(iv = factor(.$iv, levels = c("health", "corona_pos", "work_h", "work_homeoff", "hh_income"),
                         labels = c("Health", "Corona positive", "Work hours", "Homeoffice", "Household income")))
  } else if(type == "News\nuse"){
    dat_fig_results %<>%
      filter(iv %in% c("med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri")) %>% 
      mutate(iv = factor(.$iv, levels = c("med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri"),
                         labels = c("Krone", "Der Standard", "Die Presse", "Österreich", "Kurier", "Salzb. Nachrichten", "Other", "ORF", "Private news")))
  } else if(type == "Outdoor\nactivities"){
    dat_fig_results %<>%
      filter(iv %in% c("act_wrk", "act_spo", "act_frn", "act_sho", "act_pet")) %>% 
      mutate(iv = factor(.$iv, levels = c("act_wrk", "act_spo", "act_frn", "act_sho", "act_pet"),
                         labels = c("Working", "Sports", "Friends", "Shopping", "With pets")))
  } else if(type == "Psycho-\nlogy"){
    dat_fig_results %<>%
      filter(iv %in% c("risk_prop", "loc_cntrl_int_m", "sat_dem")) %>% 
      mutate(iv = factor(.$iv, levels = c("risk_prop", "loc_cntrl_int_m", "sat_dem"),
                         labels = c("Risk taking", "Internal locus of control", "Satisfaction democracy")))
  }
  return(dat_fig_results)
}
```

## make_graph_res

Make graph of effects

```{r}
make_graph_res <- function(data, sesoi = NULL, legend = TRUE, facet = "type", title = NULL){
  ggplot(data, aes(x = estimate, y = iv)) + 
  scale_color_manual(values = c("black", "grey75", "darkcyan", "deepskyblue", "cornflowerblue", "darkcyan", "aquamarine")) +
  geom_vline(xintercept = 0, lwd = .75, colour = "darkgrey") + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high, color = Analysis),
                lwd = .75, height = .2, position = position_dodge(-.7)) +
  geom_point(aes(color = Analysis), size = 1.5, position = position_dodge(-.7)) +
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=-.3), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv=="Life satisfaction"), aes(xintercept=.3), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=-.15), colour="darkgrey", linetype = "dashed")} +
    {if(isTRUE(sesoi == "est")) geom_vline(data=filter(data, dv!="Life satisfaction"), aes(xintercept=.15), colour="darkgrey", linetype = "dashed")} +
    {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=.1), colour="darkgrey", linetype = "dashed")} + 
    {if(isTRUE(sesoi == "std")) geom_vline(aes(xintercept=-.1), colour="darkgrey", linetype = "dashed")} +   
  theme(
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = .5),
    panel.spacing = unit(.9, "lines"),
    text = element_text(size = 12),
    axis.text.y = element_text(hjust = 1)
    ) +
  {if(!is.null(title)) ggtitle(title)} +
  {if(!isTRUE(legend)) theme(legend.position = "none")} + 
  # guides(colour = guide_legend(reverse=T)) +
    {if(facet == "type")
      facet_grid(rows = vars(Type),
             cols = vars(dv),
             scales = "free",
             space = "free_y")
    } + 
    {if(facet == "analysis")
      facet_grid(rows = vars(Variance),
             cols = vars(dv),
             scales = "free",
             space = "free_y")
    }
}
```

## get_fs

Extract factor scores.

```{r}
get_fs <- function(model) {
  # the aim is to get factor scores on original scaling
  # hence export factor values for all items
  # then compute average
  dat_ov <- lavaan::lavPredict(model, type = "ov", assemble = TRUE) %>% 
    mutate(fs = rowMeans(.[1:3]))
  return(dat_ov$fs)
}
```

# Data Wrangling
## Load data

```{r}
doc <- read_csv2("data/Variable_List_w1-34.csv")
d_raw <- read_csv("data/W1-32_ACPP_V20220601.csv")
```

## Find names

Identify the names of each item for each wave.

```{r}
# identify variable names of items, using custom function
health <- find_var("Gesundheitszustand") # wsn't collected at wave 21
aff_neg_1 <- find_var("Depressivitaet: einsam")
aff_neg_2 <- find_var("Depressivitaet: aergerlich")
aff_neg_3 <- find_var("Depressivitaet: so niedergeschlagen")
aff_neg_4 <- find_var("Depressivitaet: sehr nervoes")
aff_neg_5 <- find_var("Depressivitaet: aengstlich")
aff_neg_6 <- find_var("Depressivitaet: bedrueckt und traurig")
aff_pos_1 <- find_var("Depressivitaet: ruhig und gelassen")
aff_pos_2 <- find_var("Depressivitaet: gluecklich")
aff_pos_3 <- find_var("Depressivitaet: voller Energie")
act_wrk <- find_var("Zuhause verlassen: Arbeit")
act_spo <- find_var("Zuhause verlassen: Sport")
act_frn <- find_var("Zuhause verlassen: Freunde oder Verwandte treffen")
act_sho <- find_var("Zuhause verlassen: Essen einkaufen")
act_pet <- find_var("Zuhause verlassen: Haustier ausfuehren")
sat_dem <- find_var("Demokratiezufriedenheit: Oesterreich")
work_h <- find_var("Arbeitsstunden: Jetzt pro Woche")
work_homeoff <- find_var("Aenderung berufliche Situation: Home-Office")
hh_income <- find_var("Aktuelles Haushaltseinkommen")
med_txt_kro <- find_var("Mediennutzung: Kronen Zeitung oder www.krone.at")
med_txt_sta <- find_var("Mediennutzung: Der Standard oder derstandard.at")
med_txt_pre <- find_var("Mediennutzung: Die Presse oder diepresse.com")
med_txt_oes <- find_var("Mediennutzung: Oesterreich oder oe24.at")
med_txt_kur <- find_var("Mediennutzung: Kurier oder kurier.at")
med_txt_slz <- find_var("Mediennutzung: Salzburger Nachrichten oder salzburg.at")
med_txt_son <- find_var("Mediennutzung: Sonstige oesterreichische Tageszeitungen")
med_vid_orf <- find_var("Mediennutzung: ORF (Nachrichten)")
med_vid_pri <- find_var("Mediennutzung: Privatfernsehen (Nachrichten)")
soc_med_fb <- find_var("Soziale Medien: Facebook")
soc_med_tw <- find_var("Soziale Medien: Twitter")
soc_med_ig <- find_var("Soziale Medien: Instagram")
soc_med_yt <- find_var("Soziale Medien: Youtube")
soc_med_wa <- find_var("Soziale Medien: WhatsApp")
soc_med_read <- find_var("Soziale Medien Aktivitaet: Postings zu Corona anderer lesen")
soc_med_like_share <- find_var("Soziale Medien Aktivitaet: Postings liken, teilen oder retweeten")
soc_med_post <- find_var("Soziale Medien Aktivitaet: selber Postings zu Corona verfassen")
life_sat <- find_var("Lebenszufriedenheit")
risk_prop <- find_var("Risikobereitschaft")
loc_cntrl_int_1 <- find_var("Psychologie: habe Leben selbst in der Hand")
loc_cntrl_int_2 <- find_var("Psychologie: Belohnung durch Anstrengung")
loc_cntrl_int_3 <- find_var("Psychologie: Fremdbestimmung")
loc_cntrl_int_4 <- find_var("Psychologie: Schicksal")
trst_media <- find_var("Vertrauen: ORF")
trst_police <- find_var("Vertrauen: Polizei")
trst_media <- find_var("Vertrauen: Parlament")
trst_hlthsec <- find_var("Vertrauen: Gesundheitswesen")
trst_gov <- find_var("Vertrauen: Bundesregierung")
trst_army <- find_var("Vertrauen: Bundesheer")
corona_pos <- c(
  find_var("Corona-Diagnose: Respondent"),
  find_var("Corona-Diagnose: Monat")
)
```

## Select variables

```{r}
d_wide <- d_raw %>% 
  select(
    id = RESPID,
    gender = SD_GENDER,
    acc_bal = SD_ACCESS_BALCONY,
    acc_gar = SD_ACCESS_GARDEN,
    year_birth = SD_BIRTHYEAR,
    born_aus = SD_BORN_AUSTRIA,
    born_aus_prnts = SD_MIGR_BACKGR,
    county = SD_BULA,
    edu = SD_EDU,
    employment = SD_EMPLSTATUS_FEB2020,
    hh_adults = SD_HH_ADULTS,
    hh_child18 = SD_HH_CHILD18,
    hh_child17 = SD_HH_TEENS,
    hh_child14 = SD_HH_CHILD14,
    hh_child5 = SD_HH_CHILD5,
    hh_child2 = SD_HH_CHILD2,
    hh_oldfam = SD_HH_OLDERFAM,
    hh_outfam = SD_HH_OUTERFAM,
    hh_partner = SD_HH_PARTNER,
    # hh_income = SD_HHINCOME_FEB2020,
    home_sqm = SD_HOME_SQM,
    home_owner = SD_HOMEOWNER,
    # work_h = SD_WORKHOURS_FEB2020,
    health = all_of(health),
    life_sat = all_of(life_sat),
    aff_neg_1 = all_of(aff_neg_1),
    aff_neg_2 = all_of(aff_neg_2),
    aff_neg_3 = all_of(aff_neg_3),
    aff_neg_4 = all_of(aff_neg_4),
    aff_neg_5 = all_of(aff_neg_5),
    aff_neg_6 = all_of(aff_neg_6),
    aff_pos_1 = all_of(aff_pos_1),
    aff_pos_2 = all_of(aff_pos_2),
    aff_pos_3 = all_of(aff_pos_3),
    act_wrk = all_of(act_wrk),
    act_spo = all_of(act_spo),
    act_frn = all_of(act_frn),
    act_sho = all_of(act_sho),
    act_pet = all_of(act_pet),
    sat_dem = all_of(sat_dem),
    sat_dem = all_of(sat_dem),
    work_h = all_of(work_h),
    work_homeoff = all_of(work_homeoff),
    hh_income = all_of(hh_income),
    med_txt_kro = all_of(med_txt_kro),
    med_txt_sta = all_of(med_txt_sta),
    med_txt_pre = all_of(med_txt_pre),
    med_txt_oes = all_of(med_txt_oes),
    med_txt_kur = all_of(med_txt_kur),
    med_txt_slz = all_of(med_txt_slz),
    med_txt_son = all_of(med_txt_son),
    med_vid_orf = all_of(med_vid_orf),
    med_vid_pri = all_of(med_vid_pri),
    soc_med_fb = all_of(soc_med_fb),
    soc_med_tw = all_of(soc_med_tw),
    soc_med_ig = all_of(soc_med_ig),
    soc_med_yt = all_of(soc_med_yt),
    soc_med_wa = all_of(soc_med_wa),
    soc_med_like_share = all_of(soc_med_like_share),
    soc_med_read = all_of(soc_med_read),
    soc_med_post = all_of(soc_med_post),
    risk_prop = all_of(risk_prop),
    loc_cntrl_int_1 = all_of(loc_cntrl_int_1),
    loc_cntrl_int_2 = all_of(loc_cntrl_int_2),
    loc_cntrl_int_3 = all_of(loc_cntrl_int_3),
    loc_cntrl_int_4 = all_of(loc_cntrl_int_4),
    trst_media = all_of(trst_media),
    trst_police = all_of(trst_police),
    trst_media = all_of(trst_media),
    trst_hlthsec = all_of(trst_hlthsec),
    trst_gov = all_of(trst_gov),
    trst_army = all_of(trst_army),
    corona_pos = all_of(corona_pos)
  )
```

Make new documentation with selected variables.

```{r}
doc_selected <- filter(doc, Variable %in% c(
  "RESPID",
  "SD_GENDER",
  "SD_ACCESS_BALCONY",
  "SD_ACCESS_GARDEN",
  "SD_BIRTHYEAR",
  "SD_BORN_AUSTRIA",
  "SD_MIGR_BACKGR",
  "SD_BULA",
  "SD_EDU",
  "SD_EMPLSTATUS_FEB2020",
  "SD_HH_ADULTS",
  "SD_HH_CHILD18",
  "SD_HH_TEENS",
  "SD_HH_CHILD14",
  "SD_HH_CHILD5",
  "SD_HH_CHILD2",
  "SD_HH_OLDERFAM",
  "SD_HH_OUTERFAM",
  "SD_HH_PARTNER",
  "SD_HHINCOME_FEB2020",
  "SD_HOME_SQM",
  "SD_HOMEOWNER",
  "SD_WORKHOURS_FEB2020",
  health,
  life_sat,
  aff_neg_1,
  aff_neg_2,
  aff_neg_3,
  aff_neg_4,
  aff_neg_5,
  aff_neg_6,
  aff_pos_1,
  aff_pos_2,
  aff_pos_3,
  act_wrk,
  act_spo,
  act_frn,
  act_sho,
  act_pet,
  sat_dem,
  work_h,
  work_homeoff,
  hh_income,
  med_txt_kro,
  med_txt_sta,
  med_txt_pre,
  med_txt_oes,
  med_txt_kur,
  med_txt_slz,
  med_txt_son,
  med_vid_orf,
  med_vid_pri,
  soc_med_fb,
  soc_med_tw,
  soc_med_ig,
  soc_med_yt,
  soc_med_wa,
  soc_med_like_share,
  soc_med_post,
  soc_med_read,
  risk_prop,
  loc_cntrl_int_1,
  loc_cntrl_int_2,
  loc_cntrl_int_3,
  loc_cntrl_int_4,
  trst_media,
  trst_police,
  trst_media,
  trst_hlthsec,
  trst_gov,
  trst_army,
  corona_pos
))
write.csv(doc_selected, "data/documentation.csv")
```

## Recode variables

```{r}
d_wide %<>%
  mutate_at(vars(everything(.)), funs(na_if(., 88))) %>%
  mutate_at(vars(everything(.)), funs(na_if(., 99))) %>%
  mutate(
    male = 2 - .$gender,
    age = 2021 - .$year_birth,
    res_vienna = recode(.$county, `8` = 1L, .default = 0L,),
    born_aus = 2 - .$born_aus,
    home_owner = 2 - .$home_owner,
    employment_fac = factor(.$employment, 
                            labels = c("Unemployed",
                                       "Industrie",
                                       "Public service",
                                       "Self-employed",
                                       "Retired",
                                       "Housekeeping",
                                       "Student",
                                       "Incapacitated",
                                       "Parental Leave"),
                            levels = c(4, 1:3, 5:8, 10) # make unemployment reference cat
    ),
    edu_fac = factor(.$edu,
                     labels = c("No degree", 
                                "Middle school",
                                "Vocational school",
                                "Technical school",
                                "High school",
                                "Applied high school",
                                "State college",
                                "Bachelor",
                                "Master",
                                "PhD")
    ))
```

## Make long data format

```{r cache=T}
d_long <-
  d_wide %>% 
  pivot_longer(
    cols = health...W1:corona_pos...W32,
    names_to = "item",
    values_to = "value"
  ) %>% 
  separate(item, c("item", "wave"), sep = "\\.\\.\\.", extra = "merge") %>% 
  pivot_wider(names_from = "item", values_from = "value")
```

## Recode values

```{r}
# recode such that higher values imply more strength / align with wording
d_long %<>% 
  mutate_at(vars(med_txt_kro:med_vid_pri, health, sat_dem, soc_med_fb:soc_med_post), 
            funs(recode(., `1` = 5L, `2` = 4L, `3` = 3L, `4` = 2L, `5` = 1L))) %>% 
  mutate_at(vars(loc_cntrl_int_1:loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L))) %>% 
  mutate_at(vars(born_aus_prnts), 
            funs(recode(., `3` = 0L, `2` = 2L, `1` = 1L)))

# recode inverted items
d_long %<>%
  mutate_at(vars(loc_cntrl_int_3, loc_cntrl_int_4), 
            funs(recode(., `1` = 4L, `2` = 3L, `3` = 2L, `4` = 1L)))

# recode other
d_long %<>%
  mutate(
    wave = gsub("W", "", .$wave) %>% as.integer(),
    id = as.integer(id)
  )
```

## Order dataset

```{r}
d_long %<>%
  arrange(id, wave)
```

# Impute data

Next we impute data. Note that the actual imputation below is deactivated and loaded from memory to save time.

## Prep

Determine amount of missingness per respondent per wave.

```{r}
vars_used <- c("life_sat", "aff_pos_1", "aff_pos_2", "aff_pos_3", "aff_neg_1", "aff_neg_2", "aff_neg_3", "aff_neg_4", "aff_neg_5", "aff_neg_6", "soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw", "age", "male", "born_aus", "born_aus_prnts", "edu_fac", "employment_fac", "health", "res_vienna", "acc_bal", "acc_gar", "home_sqm", "med_txt_kro", "med_txt_sta", "med_txt_pre", "med_txt_oes", "med_txt_kur", "med_txt_slz", "med_txt_son", "med_vid_orf", "med_vid_pri", "risk_prop", "loc_cntrl_int_1", "loc_cntrl_int_2", "loc_cntrl_int_3", "loc_cntrl_int_4", "act_wrk", "act_spo", "act_frn", "act_sho", "act_pet", "sat_dem", "trst_media", "trst_police", "trst_media", "trst_hlthsec", "trst_gov", "trst_army", "corona_pos") # only include vars measured at _all_ points
```

Filter respondents with more than 50% missing data -- only needed for analyses as originally preregistered (see additional analyses).

```{r eval=T}
d_long_50 <- 
  d_long %>% 
  mutate(na_perc = rowSums(is.na(select(., all_of(vars_used))) / ncol(select(., all_of(vars_used))))) %>% 
  filter(na_perc < .5)
write_csv(d_long_50, "data/data_50.csv")
```

Exclude social media use data, because they were measured only on selected waves.

```{r}
vars_excl <- c("soc_med_read", "soc_med_like_share", "soc_med_post", "soc_med_fb", "soc_med_ig", "soc_med_wa", "soc_med_yt", "soc_med_tw")
incl_ma <- d_long %>% 
  mutate(across(.cols = everything(), .fns = is.na))
incl_ma[vars_excl] <- FALSE

# now also for data where 50% missing excluded
incl_ma_50 <- d_long_50 %>% 
  mutate(across(.cols = everything(), .fns = is.na))
incl_ma_50[vars_excl] <- FALSE
```

## Single imputation

For analyses as originally preregistered (see additional analyses), hence only participants who provided more than 50% of all data.

```{r imputation-single-50, eval=F, cache=T}
d_long_50_imp <- mice(d_long_50, 
                      method = "pmm", # use predictive mean matching
                      m = 1, maxit = 30, # only 1 imputation
                      # where = incl_ma_50,
                      seed = 180719, print = FALSE)
d_long_50_imp <- mice::complete(d_long_50_imp)
write_csv(d_long_50_imp, "data/data_50_imputed.csv")
```

Because CFAs aren't evaluated with multiple imputation, impute single data-set. Missing data was imputed for everyone.

```{r imputation-single-100, eval=F, cache=T}
d_long_100_imp <- mice(d_long, 
                   method = "pmm", # use predictive mean matching
                   m = 1, maxit = 30, # only 1 imputation
                   # where = incl_ma,
                   seed = 180719, print = FALSE)
d_long_100_imp <- mice::complete(d_long_100_imp)
write_csv(d_long_100_imp, "data/data_100_imputed.csv")
```

## Multiple Imputation

Now, multiple imputation of data-sets (5 data sets, 5 iterations), used for the final analyses.

```{r imputation-mult, eval=F, cache=T}
# impute missing data with multiple imputation
d_long_100_mim_mice <- mice(d_long, 
                   method = "pmm", # use predictive mean matching
                   m = 5, maxit = 5, # 5 imputations, 5 iterations
                   # where = incl_ma,
                   seed = 180719, print = FALSE)
d_long_100_mim <- mice::complete(d_long_100_mim_mice, action = "long", include = TRUE)
write_csv(d_long_100_mim, "data/data_100_mim.csv")
```

## Load data

To increase speed, you can also load the imputed data from the directory

```{r eval=T}
d_long_50_imp <- read_csv("data/data_50_imputed.csv") # data where people with >50% missing data were filtered
d_long_100_imp <- read_csv("data/data_100_imputed.csv") # data where all people were included
d_long_100_mim <- read_csv("data/data_100_mim.csv") # data where all people were included
```

# Create within-between data

First, create means for scales. Then, create mean values for each person (between), and unique changes per wave (within).

```{r}
d_long_100_mim <-
  d_long_100_mim %>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_100_mim %<>%
  group_by(id) %>% 
  mutate_at(vars(health:loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()

d_long_100_mim_mice <- as.mids(d_long_100_mim)
```

```{r}
d_long_100_imp <-
  d_long_100_imp %>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_100_imp %<>%
  group_by(id) %>% 
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

And because later we want to rerun the results as originally preregistered, we'll now do the same for analyses without filtered data.

```{r}
d_long_50_imp <-
  d_long_50_imp %>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_50_imp %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), 
            funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

And because later we want to rerun the results without imputed data, now the same for data without imputation.

```{r}
d_long_50 %<>% 
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_50 %<>% 
  group_by(id) %>% 
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>% 
  ungroup()
```

And now the same for the standardized data set.

```{r}
d_long_100_mim_std <-
  d_long_100_mim %>% 
  group_by(.imp) %>% 
  mutate_at(vars(gender:corona_pos, -id, -wave, -gender, -male, -born_aus, -born_aus_prnts, -edu_fac, -employment_fac, -res_vienna, -acc_bal, -acc_gar), ~ c(scale(.))) %>% 
  ungroup()

d_long_100_mim_std %<>%
  mutate(aff_pos_m = rowMeans(select(., aff_pos_1 : aff_pos_3)),
         aff_neg_m = rowMeans(select(., aff_neg_1 : aff_neg_6)),
         loc_cntrl_int_m = rowMeans(select(., loc_cntrl_int_1 : loc_cntrl_int_4)))

d_long_100_mim_std %<>%
  group_by(id) %>%
  mutate_at(vars(health:corona_pos, loc_cntrl_int_m), funs(b = mean(., na.rm = TRUE), w = . - mean(., na.rm = TRUE))) %>%
  ungroup()

d_long_100_mim_mice_std <- as.mids(d_long_100_mim_std)
```

# Save workspace

```{r cache=F}
save.image("data/workspace_1.RData")
```

### 2023-01-30 index.Rmd
- updated urls
- major updates to preprint

@@
 ---
-title: "No meaningful effects of COVID-19 related social media use on well-being"
+title: "The effects of COVID-19 related social media use on well-being"
 ---
 
-On this website, you can find the companion material for the paper _No meaningful effects of COVID-19 related social media use on well-being_.
+On this website, you can find the companion material for the paper _The effects of COVID-19 related social media use on well-being_.
 
@@
 - The paper, which you can read [in your browser](manuscript.html) or download as [pdf](manuscript.pdf). 
+- The [set-up](analyses_setup.html) for the analyses, including session information, custom functions, data wrangling, and data imputation.
 - The detailed code and summaries of the [analyses](analyses.html)
@@
 
-This website is produced directly from the project's [github repository](https://github.com/tdienlin/Austrian_Corona_Panel_Project).
+This website is produced directly from the project's [github repository](https://github.com/XMtRa/Austrian_Corona_Panel_Project).
 You can download the material and the data, file an issue, or submit a pull request.

### 2023-01-30 manuscript.Rmd
- major updates to preprint

--- 2021-12-01 manuscript.Rmd
@@
 
+note: |
+  Preprint submitted for publication. Please cite carefully.
+
 abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
@@
 The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
-keywords          : "COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal."
+keywords          : "COVID-19, well-being, social media, news use, panel study."
 
@@
 floatsintext      : yes
+endnotes          : yes
 figurelist        : no
@@
 linenumbers       : yes
-mask              : yes
+mask              : no
 draft             : no
 
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
-# output            : papaja::apa6_word
 header-includes:
-  - \setlength{\parskip}{0em}
-  - \raggedbottom
-  - \note{\clearpage}
+  - |
+    \makeatletter
+    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-1em}%
+      {\normalfont\normalsize\bfseries\typesectitle}}
+
+    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+      {-\z@\relax}%
+      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+    \makeatother
+    \usepackage{endnotes}
+    \setlength{\parskip}{0em}
+    \raggedbottom
+    \note{\clearpage}
+
+csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+documentclass     : "apa7"
+output            :
+  papaja::apa6_pdf:
+    latex_engine  : xelatex
+# output            : papaja::apa6_docx
+# output            : html_document
 ---
@@
 ```{r analysis-preferences, include=F, cache=F}
-# Seed for random number generation
-set.seed(42)
+set.seed(170819) # Seed for random number generation
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed,
-                      cache=T, echo=F
-                      )
+                      cache=T, echo=F)
 ```
@@
 # load packages
-packages <- c("devtools", "english", "ggplot2", "kableExtra", 
+packages <- c("devtools", "english", "GGally", "ggplot2", "gridExtra", "kableExtra", 
               "lavaan", "lme4", "lmerTest",
-              "magrittr", "semPlot", "tidyverse")
-lapply(c(packages, "papaja", "td"), library, character.only = TRUE)
+              "magrittr", "papaja", "tidyverse")
+lapply(c(packages, "td"), library, character.only = TRUE)
 
@@
 
-During the COVID-19 pandemic, 
-numerous events unfolded in quick succession and several open questions emerged.
+During the COVID-19 pandemic it was critical to stay informed regarding the latest developments.
 How dangerous is the virus? 
-Is it spreading in my region? 
-How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high [@statistaAverageDailyTime2021].
+In what region is it spreading? 
+How is it transmitted? What are the current safety regulations?
+To obtain relevant information, many people heavily relied on social media, which were at an all time high [@statistaAverageDailyTime2021].
 
-Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon emerged, termed "doomscrolling":
-Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children [@kleinDarklySoothingCompulsion2021].
-As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
-These concerns seem justified: 
-A study with 6,233 people from Germany conducted during the pandemic found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression" [@bendauAssociationsCOVID19Related2021, p.
-283].
+Some actually could not stop using social media to learn about COVID-19 related news.
+A new phenomenon termed "doomscrolling" emerged.
+Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children [@kleinDarklySoothingCompulsion2021].
+In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health [@sandstromDoomscrollingCOVIDNews2021].
+Although research has begun addressing this question [@bendauAssociationsCOVID19Related2021; @bradleyStressMoodSmartphone2021; @choiMediatedCommunicationMatters2021; @dornemannHowGoodBad2021; @edenMediaCopingCOVID192020; @guazziniSecondWaveAnalysis2022; @latikkaLonelinessPsychologicalDistress2022; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020; @sewallObjectivelyMeasuredDigital2021; @stainbackCOVID1924News2020; @yuePassiveSocialMedia2022], it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
+By analyzing large-scale panel data from Austria and by focusing on intra-individual causal effects, this study aims to answer to this question.
 
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
-To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2020]. 
-The panel consists of 24 waves and has an overall sample size of `r nrow(d_wide) %>% prettyNum(big.mark = ",")` participants.
-<!-- At least 1,500 participants took part per wave, and the sample is representative of the Austrian population. -->
-The panel study collected a large number of psychological and demographic variables.
-<!-- Being able to control for many confounding third variables, both stable and varying, together with the longitudinal design, this creates a unique opportunity  -->
-I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
+## Understanding Well-being and Media Use
 
-## Defining Well-being and Media Use
+<!-- Two underlying theories guided the selection of variables for this study, namely the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being. -->
+<!-- <!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
+<!-- Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018]. -->
+<!-- Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning. -->
+<!-- Subjective well-being is primarily about achieving positive and avoiding negative affect. -->
+<!-- One of the most prominent indicators of well-being is life satisfaction. -->
+<!-- In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept combining psychological and subjective well-being. -->
+<!-- Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020]. -->
+<!-- <!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
+<!-- To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect. -->
+<!-- Together, this should provide an encompassing perspective on potential media effects. -->
 
-The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health [@greenspoonIntegrationSubjectiveWellbeing2001] and the hierarchical taxonomy of computer-mediated communication [@meierComputermediatedCommunicationSocial2020a].
-According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
-<!-- Because the aim of this study is better to understand typical users and everyday contexts, my focus will be on well-being. -->
-Well-being can be differentiated into subjective and psychological well-being [@dienerAdvancesOpenQuestions2018].
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
-Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect [@dienlinImpactDigitalTechnology2020].
-<!-- Everyone can feel bad on any given day, but whether or not one feels generally comfortable in one's life rather changes across years. -->
-To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on potential media effects.
+<!-- The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology. -->
+<!-- First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_. -->
+<!-- Distinguishing different communication channels and communication types is important, because the effects on well-being likely differ across communication channels and communication types. -->
+<!-- Whereas active social media use such as chatting is routinely linked to improved well-being, passive use such as reading is often considered more negative [@dienlinImpactDigitalTechnology2020].
+-->
+<!-- Similarly, branded apps are separate communication entities with potentially divergent effects and affordances. -->
+<!-- For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. -->
+<!-- Especially during a pandemic, it makes sense to analyze if users engage with COVID-19 related content on Instagram, where communication is more positive, or on Facebook, where communication is more critical.
+-->
+<!-- First studies suggest that during the pandemic Instagram use was indeed more beneficial for well-being than Facebook use [@masciantonioDonPutAll2021]. -->
+<!-- In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective. -->
+<!-- Together, this should offer a nuanced and comprehensive understanding of communication. -->
 
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) [@meierComputermediatedCommunicationSocial2020a].
-Whereas the first four levels focus on the communication _channel_, the last two address the communication _type_.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
+This study investigates how different _facets_ of well-being are affected by different _types_ and different _channels_ of communication [@meierComputermediatedCommunicationSocial2020a].
+<!-- , namely active and passive use [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a]. -->
+<!-- Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- Reading is generally considered as passive and writing as active use, while there are also specific behaviors falling somewhere in-between such as liking or sharing content [@meierDoesPassiveSocial2022].
+-->
+<!-- In this study,  -->
+Building on the typology of subjective well-being [@dienerAdvancesOpenQuestions2018], three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
+Because effects of social media depend on how they are used [@verduynPassiveFacebookUsage2015], I further distinguish three types of use and five popular channels. 
+The types of use include reading, posting, and liking and sharing COVID-19 related posts.
+In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.
+The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
 
-First, I investigate how different types of communication affect well-being.
-Specifically, I differentiate between active and passive use.
-I distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially divergent effects. 
-Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
-For example, @waterlooNormsOnlineExpressions2018 found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram. 
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about _general_ social media use during times of COVID, but on social media use _focused_ on COVID-19 related content. 
-Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
+<!-- ## Theorizing Social Media Effects on Well-Being -->
 
-## Effects of Social Media on Well-Being
+<!-- <!-- In conclusion, most effects are likely somewhere between trivial and small. -->
+<!-- <!-- I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small. -->
 
-So far, there is only little empirical research on COVID-19 related social media use on well-being.
-In their study on the relations between media use and mental health during the pandemic, @bendauAssociationsCOVID19Related2021 found that people who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [] and significantly more specific COVID-19 related anxiety symptoms" (p.
-288).
-@edenMediaCopingCOVID192020 analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, @choiMediatedCommunicationMatters2021 reported that people who used media to attain information were more lonely and less satisfied with their lives.
-@stainbackCOVID1924News2020 analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
-Together, the literature emphasizes potentially negative effects of social media as news use [see also @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
-However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
-We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
+How could the various types and channels of COVID-19 related social media use affect well-being?
+<!-- In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent. -->
+According to the set-point theory, well-being is surprisingly stable [@lykkenHappinessWhatStudies1999].
+Although specific events such as marriage or salary can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time routinely returning to prior levels [@sheldonStabilityHappinessTheories2014].
+Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being [@lucasAdaptationSetpointModel2007].
+<!-- So although well-being can change this does not happen easily. -->
+<!-- Although the set-point theory has been criticized lately because events such as marriage, unemployment, or death can indeed lead to long-term changes, it is still widely maintained that well-being is very stable and hence hard to be influenced  -->
+Can media use be such a factor?
+Current literature overviews suggest no: 
+Although social media use on average decreases well-being, the effects are small [@meierComputermediatedCommunicationSocial2020a].
+In addition, the effects of media use differ across individuals and types of content [@valkenburgDifferentialSusceptibilityMedia2013].
+Whereas for some users social media are more beneficial, for others they are more harmful.
+For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
+This finding is aligned with the Differential Susceptibility to Media Effects Model: 
+<!-- This finding is aligned with the Differential Susceptibility to Media Effects Model:  -->
+Although there is substantial _variation_ of media effects for individual users, there are both negative and positive aspects, leading to small effects on average [@valkenburgDifferentialSusceptibilityMedia2013]---as confirmed by several literature overviews [@dienlinImpactDigitalTechnology2020; @huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020].
 
-The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched.
-This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small [@meierComputermediatedCommunicationSocial2020a]---potentially too small to matter.
-What determines whether or not an effect is considered small or trivial?
-As a starting point, we could refer to standardized effect sizes.
-According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10.
-And indeed, several if not most of the current meta-analyses find effect sizes below that threshold [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
+Why are the effects of social media use on well-being small on average?
+Using Social media is ambivalent.
+It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment [@buchiDigitalWellbeingTheory2021].
+Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes.
+Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself [e.g., @pelletierOneSizeDoesn2020].
+Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons [e.g., @meierDoesPassiveSocial2022].
+People are much more likely to use media that offer plentiful benefits [@katzUsesGratificationsResearch1973].
+Users implicitly learn what media help them regulate their mood and thereby well-being [@zillmannMoodManagementCommunication1988].
+Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
 
-These overviews are well aligned with several individual studies employing advanced methods [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
-For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-In conclusion, most effects are likely somewhere between trivial and small.
-I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.
+<!-- Two prominent media effect theories argue (mostly implicitly) against strong average negative effects. -->
+<!-- According to mood management theory [@zillmannMoodManagementCommunication1988], using media can substantially affect people's moods. -->
+<!-- Use can be stimulating or overwhelming, relaxing or boring. -->
+<!-- After some time, users implicitly learn what media help them balance their mood and affect according to their own situational needs [@zillmannMoodManagementCommunication1988]. -->
+<!-- Those media that eventually become part of one's media repertoire are hence, on average, beneficial for users to regulate their mood. -->
+<!-- Using experience sampling of well-being and logs of social media use, a study with 82 participants from Italy found that after episodes of social media use, levels of positive affect increased significantly [@marcianoDynamicsAdolescentsSmartphone2022].
+-->
 
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
-In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
+<!-- While mood management theory considers media use mainly driven by implicit learning experiences, uses and gratifications theory upholds that the process is more explicit and rational [@katzUsesGratificationsResearch1973].
+-->
+<!-- Users select those media that they expect to have a desired effect, for example on mood, knowledge, or entertainment.  -->
+<!-- If those beneficial media effects do not exist or if they are not expected, people will spend their time elsewhere. -->
+<!-- And social media offer several beneficial effects, explaining why they are used that much. -->
+<!-- They help find relevant information, maintain and foster relationships, express one's personality, and entertain oneself [@pelletierOneSizeDoesn2020]. -->
+<!-- In conclusion, because people spend so much time on social media consuming COVID-19 related content, according to both mood management theory and uses and gratifications theory this is indirect proof that average effects on well-being are likely not particularly negative.
+-->
 
-First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news can be depressing, perhaps even changing some general perspectives on life.
-<!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.
+<!-- But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017]. -->
+<!-- Precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies. -->
+<!-- In other words, social media might not problematic because they are inherently bad, but rather because they are too good. -->
+<!-- As with many other things, there can be too much of a good thing. -->
+<!-- It is therefore often asked whether social media can become addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021].  -->
+<!-- However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020].
+-->
+<!-- In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].
+-->
+
+<!-- Because media effects can differ across users, situations, and content [@livingstoneEuropeanResearchChildren2018; @valkenburgDifferentialSusceptibilityMedia2013], I now briefly focus on the effects of COVID-19 related social media use specifically.
+-->
+<!-- First, one could assume a _direct_ negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating. -->
+<!-- Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide. -->
+<!-- If one learns about such events, the initial reaction might be shock, fear, or dismay. -->
+<!-- Consuming such news can be depressing [@dornemannHowGoodBad2021], perhaps even changing some general perspectives on life. -->
+<!-- <!-- So, just like being hit by a hammer hurts and we don't need any "mediating mechanism", this could be the case here as well. -->
+<!-- That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects [@dornemannHowGoodBad2021].
+-->
+<!-- A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness even increased [@guazziniSecondWaveAnalysis2022]. -->
+
+<!-- In a study with 735 participants from Finland, levels of loneliness did not decrease during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
+-->
 <!-- However, in light of a worldwide pandemic with millions of deaths, the negative direct effect seems more plausible. -->
 
-There could also be _indirect_ effects.
-When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
-For example, during the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
-In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020].
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
-To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
+<!-- Second, there could also be _indirect_ effects. -->
+<!-- Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
+-->
+<!-- To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising. -->
+<!-- Another potentially negative mechanism at play are problematic social comparison processes.  -->
+<!-- Therefore, if a society collectively overuses social media during a pandemic, there is potential for negative effects. -->
 <!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->
 
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-<!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
+<!-- On the other hand, one can make the case that using social media for COVID-19 related reasons might even be beneficial, especially in times of a pandemic. -->
+<!-- Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time. -->
+<!-- In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea. -->
+<!-- Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments. -->
+<!-- <!-- On the other hand, there is of course also much disinformation, and "bingeing" on COVID-19 fake news might also pose risks for impaired well-being. -->
 
-Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.
-Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
-Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
+<!-- To summarize, it seems that from a theoretical perspective it is most likely that the average effects of social media use on well-being are negligible.  -->
+<!-- Building on established theories from Communication and current empirical findings, we would not assume that effects are either profoundly negative or strongly positive. -->
+
+<!-- ## Empirical Studies on Social Media Effects -->
+
+There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
+Results are mixed.
+When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media [@kleinDarklySoothingCompulsion2021].
+<!-- During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021], for example to cope with the required physical distancing measures. -->
+In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media [@yuePassiveSocialMedia2022]. 
+Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.
+People who used social media as a primary source of information reported on average "significantly more unspecific anxiety and depression [...] and significantly more specific COVID-19 related anxiety symptoms" [@bendauAssociationsCOVID19Related2021, p.
+288].
+@edenMediaCopingCOVID192020 analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being.
+<!-- In a sample of 312 respondents collected via Amazon Mechanical Turk, people who used media to attain information were more lonely and less satisfied with their lives [@choiMediatedCommunicationMatters2021].
+-->
+In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress [@stainbackCOVID1924News2020].
+A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health [@sewallObjectivelyMeasuredDigital2021; for a similar study with comparable results, see @bradleyStressMoodSmartphone2021].
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased [@guazziniSecondWaveAnalysis2022].
+In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness [@latikkaLonelinessPsychologicalDistress2022].
+Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being [see also @dornemannHowGoodBad2021; @liuRelationOfficialWhatsAppdistributed2020; @riehmAssociationsMediaExposure2020].
+<!-- However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below). -->
+<!-- We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education. -->
+
+<!-- The question of whether and how social media use affects well-being _in general_, on the other hand, is well-researched. -->
+<!-- This also holds true for the different types of communication such as active or passive use. -->
+<!-- What determines whether or not an effect should be considered small or trivial? -->
+<!-- As a starting point, we could refer to standardized effect sizes. -->
+<!-- According to @cohenPowerPrimer1992, small effect sizes start at _r_ = .10. -->
+<!-- Finally, several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
+-->
+<!-- Although for roughly one quarter of all users the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral [@beyensSocialMediaUse2021].
+-->
+<!-- This finding is aligned with the Differential Susceptibility to Media Effects Model:  -->
+<!-- Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].
+-->
+
+<!-- Together, the strongest argument to me is that _in general_ the effects of social media on well-being are, on average, small at best.  -->
+<!-- Because this study only looks at _one part_ of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further. -->
+<!-- Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear. -->
 <!-- At least, there does not seem to be a case where we should expect a clear deviation to the negative, despite anecdotes of doomscrolling. -->
-Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+
+In conclusion, these mixed empirical results, together with the observation that social media effects on well-being are very small in general, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
+
+<!-- A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small, potentially too small to matter [@meierComputermediatedCommunicationSocial2020a].
+-->
+<!-- And indeed, several if not most of the current meta-analyses find small to trivial effects [@huangTimeSpentSocial2017; @meierComputermediatedCommunicationSocial2020a; @fergusonThisMetaanalysisScreen2021].
+-->
 
@@
 
-# Current Study
+<!-- In this study, this general hypothesis will be analyzed specifically for the three communication types of (a) time spent reading, (b) liking and sharing, and (c) actively posting COVID-19 related content.
+-->
+<!-- In addition, I will analyze how well-being is influenced by spending time on five prominent social media apps, including (a) Facebook, (b) Instagram, (c) Twitter, (d) WhatsApp, and (e) YouTube.
+-->
+
+# Method
+## Preregistration
+
+The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, _p_-value) were preregistered on the Open Science Framework, accessible here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
+Because in this study I analyze data from an already existing large-scale data set, the preregistration was done prior to accessing the data.
+The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
+In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project) (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+
+## Sample
+
+The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021], which is a large-scale standalone panel study. 
+The data are hosted on AUSSDA, are publicly available here (https://doi.org/10.11587/28KQNS), and consist of 32 waves.
+The study was conducted between March 2020 and June 2022, and data collection is now finished.
+Between March 2020 and July 2020, the intervals between waves were weekly and afterward monthly.
+Each wave consists of at least 1,500 respondents.
+The sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, with overall `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations.
+Panel mortality was compensated through a continuous re-acquisition of new participants. 
+Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
+Panel members were incentivized with 180 credit points for each wave of the study.
+
+Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
+In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
+All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets). 
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements. 
+The participants provided their written informed consent to participate in this study.
+The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
+
+<!-- # Current Study -->
 ## Smallest Effect Size of Interest
+<!-- ## Inference Criteria -->
 
-Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis necessitates defining what's considered a "trivial effect size" and what's not.
-Above I already referred to standardized effect sizes.
-However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009].
-Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic: 
-The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
-Instead, it should depend on absolute criteria.
-
-What could be a minimally interesting, nontrivial effect?
-Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
+Testing the hypothesis necessitates defining what is considered a "trivial effect size". 
+<!-- Effectively testing this hypothesis  -->
+<!-- Above I already referred to standardized effect sizes. -->
+<!-- However, standardized effect sizes should only be a first step toward evaluating an effect's relevance [@baguleyStandardizedSimpleEffect2009]. -->
+<!-- Standardized effect sizes are determined by a sample's variance,[^cohensd] which is problematic:  -->
+<!-- The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected. -->
+<!-- Instead, it should depend on absolute criteria. -->
+<!-- What could be a minimally interesting, nontrivial effect? -->
+Being a normative question, finding a clear, single, or unanimous answer is impossible.
 <!-- In the end, it is a personal question. -->
-However, it is still necessary and helpful to try provide such a plausible benchmark. 
-I therefore suggest the following SESOI for this research question:
-
-[^cohensd]: Consider the effect size Cohen's _d_: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation [@cohenPowerPrimer1992].
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.
+However, it is still necessary and helpful to work toward a plausible benchmark. 
+I suggest the following SESOI for this research question:
 
@@
 Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-But what's a noticeable change in well-being?
-According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish _seven_ levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction. 
-(For more information, see Methods section "Inference Criteria.")
-
-## Causality
-
-The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs can help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements [@rohrerTheseAreNot2021].
-In addition, we for example also need to control for confounding third variables, and importantly also for _varying_ third variables.
-
-To illustrate, consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life.
-Eventually, use and life satisfaction return to prior levels.
-If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
-
-Thus, only when controlling for _all_ relevant confounders, we can correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
-Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
-In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
-Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider]
-However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
-
-[^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018].
-
-When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being. 
-Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect only social media use or well-being. 
-However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
-
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
-
-Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
-If a specific person changes their media diet, we need to measure how this behavior affects _their own_ well-being.
-Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I hence differentiate between-person relations from within-person effects. 
-And as explicated above, to test the hypothesis I thus consider only the within-person effects.
-
-Finally, one precondition of causality is temporal order. 
-The cause needs to precede the effect.
-Finding the right interval between cause and effect is crucial.
-For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-<!-- Finding the right interval is difficult. -->
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
-Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
-
-In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there _simultaneous_ changes in their well-being?
-When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-<!-- Similar approaches were already implemented by several other studies [@johannesNoEffectDifferent2021; @scharkowHowSocialNetwork2020] and they are considered a best practice to analyze causality [@bellFixedRandomEffects2019].
+What is a noticeable change in well-being?
+According to @normanInterpretationChangesHealthrelated2003, people can reliably distinguish seven levels of satisfaction with health.
+So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction. 
+<!-- Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted. -->
+<!-- The sample is large, and it is hence well-equipped to reliably detect small effects. -->
+<!-- <!-- , which is why no exact post hoc power analysis were conducted. -->
+<!-- In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
 -->
-
-# Method
-
-In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
-
-## Preregistration
-
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework. 
-The (anonymous) preregistration can be accessed here: https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3.
-Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online [@kittelAustrianCoronaPanel2020].
-In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html) (https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
-
-## Sample
-
-The data come from the Austrian Corona Panel Project [@kittelAustrianCoronaPanel2021].
-The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing the first 24 waves were available for download. 
-Each wave consists of at least 1,500 respondents.
-The overall sample size was _N_ = `r d_raw %>% nrow() %>% prettyNum(big.mark = ",")`, and `r d_long %>% nrow() %>% prettyNum(big.mark = ",")` observations were collected.
-Panel mortality was compensated through a continuous acquisition of new participants. 
-All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.
-
-The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level. 
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age. 
-The study received IRB approval from the University of Vienna.
-The average age was `r mean(2021 - d_wide$year_birth, na.rm = T) %>% round(0)` years, `r mean(d_wide$male, na.rm = T) %>% round(., 2) * 100` percent were male, `r (d_wide %>% filter(.$edu_fac %in% c("State college", "Bachelor", "Master", "PhD"))) %>% nrow() %>% divide_by(sum(table(d_wide$edu_fac))) %>% multiply_by(100) %>% round(0)` percent had a University degree, and `r (table(d_wide$employment_fac)["Unemployed"] / sum(table(d_wide$employment_fac))) %>% unname() %>% multiply_by(100) %>% round(., digits = 0)` percent were currently unemployed.
-
-## Inference Criteria
-
-Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped reliably to detect also small effects.
-<!-- , which is why no exact post hoc power analysis were conducted. -->
-In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
-To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
-On the basis of the SESOI, I then defined a null region.
-In what follows, I explain how I determined the SESOI and the null region.
+<!-- To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014. -->
+<!-- On the basis of the SESOI, I defined a null region. -->
+<!-- In what follows, I explain how I determined the SESOI and the null region. -->
 
 In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
+If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = `r round(11/7, 2)` unit change on an 11-point scale.
 Hence, a four-point change in media use (e.g., a complete stop) should result in a `r round(11/7, 2)`-point change in life satisfaction.
 In a statistical regression analysis, _b_ estimates the change in the dependent variable if the independent variable increases by one point.
-We would therefore expect a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
-For affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
-Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
+For life satisfaction, we would therefore define a SESOI of _b_ = `r round(11/7, 2)` / 4 = `r round(11/7*.25, 2)`.
+For positive or negative affect, which was measured on a 5-point scale, our SESOI would be _b_ = `r round(5 / 7, 2)` / 4 = `r round(5 / 7 * .25, 2)`.
+Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
 Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.[^rounding] 
@@
 <!-- As explained above, for well-being the null region was between _b_ = -.30 and _b_ = .30 (i.e., _b_ = |.30|). -->
-Let's briefly illustrate what this means in practice.
-If the 95% confidence interval falls completely within the null-region (e.g., _b_ = .20, [95% CI: .15, .25]), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., _b_ = .30, [95% CI: .25, .35]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., _b_ = .40, [95% CI: .35, .45]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+The hypothesis is analyzed using the interval testing approach as proposed by @dienesUsingBayesGet2014.
+To illustrate, if the 95% confidence interval falls completely within the null-region (e.g., _b_ = -.02, [95% CI: -.12, .08]), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., _b_ = -.22, [95% CI: -.27, -.17]), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., _b_ = -.40, [95% CI: -.45, -.35]), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+For an illustration, see Figure \@ref(fig:sesoi).
+
+```{r sesoi, warning=F, message=F, fig.cap="Using confidence intervals to test a null region.
+In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30", fig.width=6, fig.height=2, warning=F, message=F, prompt=F}
+# make table with data
+d_plot <- tribble(
+  ~name, ~type, ~Effect, ~ll, ~ul,
+  "1. Accept trivial effect", "Decision Rule", -.02, -.12, .08, 
+  "2. Reject trivial effect", "Decision Rule", -.40, -.45, -.35,
+  "3. Reject positive effect", "Decision Rule", -.22, -.27, -.17,
+  "4. Suspend judgement", "Decision Rule", -.1, -.32, .08,
+) %>% 
+  mutate(
+    name = factor(name, levels = name),
+    name = fct_rev(name)
+         )
+
+# design plot
+plot <- ggplot(d_plot, aes(y = name, x = Effect)) +
+  geom_vline(xintercept = 0, color = "darkgrey") + 
+  geom_vline(xintercept = -.3, color = "darkgrey", linetype = "dashed") +
+  geom_vline(xintercept = .3, color = "darkgrey", linetype = "dashed") +
+  geom_point(aes(Effect)) +
+  geom_errorbarh(aes(xmin = ll, xmax = ul), height = .5) +
+  theme(axis.title.y = element_blank()) +
+  labs(x = "Effect Size (b)",
+       caption = "Smallest effect size of interest: b = |.30|
+                  Null region: b = -.30, .30")
+
+# ggsave("figures/figure_intervals.png", height = 3.5)
+plot
+```
 
@@
 ## Data Analysis
+### Causality
 
-The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)[@bellFixedRandomEffects2019]. 
-Three models were run, one for each dependent variable.
+<!-- The hypothesis explicitly states a causal effect. -->
+<!-- In non-experimental studies, longitudinal designs can help investigate causality. -->
+When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects [@hamakerWhyResearchersShould2014]; to (b) control for confounders [@rohrerTheseAreNot2021]; and to (c) test a plausible interval between measures [@dormannOptimalTimeLags2015].
+<!-- To illustrate, consider the following example. -->
+<!-- Imagine that a person suddenly starts using social media much more than usual, and then after some time  becomes less satisfied with their life. -->
+<!-- Eventually, use and life satisfaction return to prior levels. -->
+<!-- If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction. -->
+<!-- However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs. -->
+<!-- Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down. -->
+<!-- Thus, only when controlling for _all_ relevant confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018]. -->
+<!-- Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible. -->
+To elaborate, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective [@hamakerWhyResearchersShould2014].
+If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
+Between-person comparisons from longitudinal data cannot provide such insights [@lucasWhyCrossLaggedPanel2022].
+To test the hypothesis, I thus consider only the within-person effects.
+
+Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect of social media use on well-being [@rohrerThinkingClearlyCorrelations2018].
+Because we are adopting a within-person perspective, we need to implement _time-varying_ confounders [@rohrerTheseAreNot2021]. 
+And because we are determining the _overall_ causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
+<!-- Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider] -->
+<!-- However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly. -->
+<!-- The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
+-->
+<!-- [^collider]: In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further [@rohrerThinkingClearlyCorrelations2018]. -->
+<!-- We can also control for variables that affect only social media use or well-being.  -->
+<!-- However, in doing so not much is gained, because the effects of social media use would remain virtually the same [@klinePrinciplesPracticeStructural2016; but see @mcelreathYesterdayClass2021].
+-->
+In this study, I hence preregistered to control for the following variables, which either have already been shown or are very likely to affect both social media use and well-being, and which also are not mediators:
+<!-- (I'll additionally include variables that likely affect only well-being, also to obtain a comparison benchmark for social media effects): -->
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+<!-- [^explore] -->
+<!-- I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent. -->
+
+<!-- [^explore]: The data-set includes many other variables that one could also potentially control for, and I invite interested readers to download the and explore potential interesting relationships.
+-->
+
+<!-- Next to including covariates, it is now increasingly understood that  -->
+Finally, one precondition of causality is temporal order and finding a plausible interval [@dormannOptimalTimeLags2015].
+<!-- The cause needs to precede the effect. -->
+<!-- Finding the right interval between cause and effect is crucial. -->
+<!-- For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
+-->
+<!-- Finding the right interval is difficult. -->
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are required.
+In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction [@dienlinImpactDigitalTechnology2020].
+Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].
+Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].
+<!-- Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
+-->
+In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
+For life satisfaction, I implemented a longer interval.
+If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
+These relations will be controlled for varying confounders, which fosters a causal interpretation.
+<!-- Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].
+-->
+
+### Statistical model
+
+The hypothesis was analyzed using random effect within-between models [REWB, @bellFixedRandomEffects2019]. 
+Altogether three models were run, one for each dependent variable.
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
@@
 They included social media communication types and channels, separated into within and between-person factors, as well as stable and varying covariates.
-All predictors were included simultaneously and in each of the three models.
+All predictors were included simultaneously in each of the three models.
 
@@
 Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid overfitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
-Finally, REWB-models cannot model latent variables.
-To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-Respondents who answered less than 50% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
+<!-- To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016]. -->
+<!-- Finally, REWB-models cannot model latent variables. -->
+<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
+Mean scores were used for positive and negative affect.
+<!-- Respondents who answered less than 50% of all questions were removed. -->
+Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.
+All variables were imputed except the social media use measures, as they were not collected on each wave.
+All variables included in the analyses presented here were used to impute missing data.
+<!-- For the main analyses, results were pooled across all five data-sets. -->
 
-For more information on the analyses, a complete documentation of the models and results, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
+### Robustness-checks
+
+To find out whether the inferences were robust across plausible but subordinate analyses, I conducted additional exploratory analyses. 
+I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
+For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+<!-- The additional analyses are reported on the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html). -->
+<!-- For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were negligible.
+-->
 
@@
 
-In what follows, I list all the variables that I analyzed.
-For the variables' means, range, and variance, see Table \@ref(tab:descriptives).
-For a complete list of all items and item characteristics, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
+<!-- In what follows, I list all the variables that I analyzed. -->
+For the variables' means, range, and variance, see Table \@ref(tab:tab-descriptives).
+For a complete list of all items and item characteristics, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 
-Life satisfaction was measured with the item "Taken everything together, how satisfied are you currently with your life?".
-The response options ranged from 0 (_extremely unsatisfied_) to 10 (_extremely satisfied_).
+Life satisfaction was measured with the item "All things considered, how satisfied are you with your life as a whole nowadays?", which comes from the European Social Survey [@europeansocialsurveyESS9Edition20182021].
+The response options ranged from 0 (_extremely dissatisfied_) to 10 (_extremely satisfied_).
 
-To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
+To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy [@worldhealthorganizationWellbeingMeasuresPrimary1998].
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
@@
 
-For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
+For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad [@worldhealthorganizationWellbeingMeasuresPrimary1998].
 The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
@@
 COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
-The general introductory question was "How often during the last week have you engaged in the following activities on social media?"
+The items come from @wagnerAUTNESOnlinePanel2018 and were adapted for the context of this study.
+The general introductory question was "How often during the last week have you engaged in the following activities on social media?".
 The three items were "Reading the posts of others with content on the Coronavirus", "When seeing posts on the Coronavirus, I clicked 'like', 'share' or 'retweet'", "I myself wrote posts on the Coronavirus on social media."
@@
 
-COVID-19 related social media use focused on channels was measured with five variables.
+COVID-19 related social media use focused on channels was measured with five variables from @wagnerAUTNESOnlinePanel2018, adapted for this study. 
 The general introductory question was "How often in the last week have you followed information related to the Corona-crisis on the following social media?"
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, and 23. 
-Freshly recruited respondents always answered all questions on social media use.
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28.
+Freshly recruited respondents always answered all questions on COVID 19-related social media use.
+For the analyses I hence used all 32 waves.
 
-```{r descriptives, cache=F}
+```{r tab-descriptives, cache=F, warning=F, message=F, prompt=F}
 apa_table(tab_desc_dat,
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables: 
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement additional variables as varying covariates.
-However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
+gender (female, male, diverse), age, education (ten options), Austria country of birth (yes/no), Austria parents' country of birth (no parent, one parent, both parents), and household size.
+<!-- I originally planned to implement additional variables as varying covariates. -->
+I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and three more psychological measures including locus of control, disposition to take risks, and satisfaction with democracy.
 
 # Results
+## Descriptive Analyses
 
-First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
-COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
+Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were surprisingly stable during data collection (see Figure \@ref(fig:fig-descriptives)).
+COVID-19 related social media use, however, showed changes. 
+Reading, sharing and liking COVID-19 related content decreased substantially (almost one scale point from 3 to 2). 
+Posting about COVID-19 related content stayed the same.
+Using Facebook, WhatsApp, and YouTube for COVID-19 related content decreased. 
+Instagram stayed the same.
+Twitter use slightly increased.
+The general initial decrease could be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic's onset.
+After an initial uptick, COVID-19 related social media use might have already been declining at the time.
 
-```{r fig-desc, fig.cap="Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_descriptives.pdf")
+```{r fig-descriptives, fig.cap="Well-being and media use across the 32 waves.
+Note.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=F}
+# fig_desc
+knitr::include_graphics("figures/fig_descriptives.png")
+```
+
+Using the average values across all waves, which provides a stable picture of the general relations, I next looked at the correlations between social media use and well-being (see Figure \@ref(fig:fig-correlations)).
+Several interesting patterns emerged.
+In general, people who spend more time engaging with COVID-19 related content on social media reported reduced well-being.
+Users who spend more time reading, liking and sharing, and posting COVID-19 related content were less satisfied with their lives.
+They also showed slightly less positive affect.
+This overall negative picture was even more pronounced for negative affect.
+People who engaged more with COVID-19 related content, including all types and channels of communication, reported substantially higher levels of negative affect.
+For example, people who were more likely to post COVID-19 content had much higher levels of negative affect (_r_ = `r tab_cor[tab_cor$term == "Posting", "Affect.negative"]`).
+Note that these results represent between-person correlations, not causal within-person effects.
+
+```{r fig-correlations, fig.cap="Descriptives of the main variables, capturing well-being and social media use with their average values across all waves.
+Upper triangle: correlation coefficients; diagonal: density plots; lower triangle: scatter plots.", out.width = "\\textwidth", cache=F, warning=F, message=F, prompt=F}
+# fig_cor
+knitr::include_graphics("figures/fig_cor.png")
 ```
@@
 ## Preregistered Analyses
+### Social media communication types
 
-The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication _types_---that is, reading vs.
-sharing vs.
-posting---all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:res-activity)).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (_b_ = `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_activity %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within") %>% select(conf.high) %>% round(2)`]).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
-As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
+Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \@ref(fig:fig-within)).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (_b_ = `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Liking & Sharing" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+As a result, the hypothesis of trivial effects was supported for all COVID-19 related types of social media communication.
 
-```{r res-activity, fig.cap="The effects of various types of social media use on three indicators of well-being.
+However, three effects stood out, as they were significantly different from zero.
+Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (_b_ = `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Positive affect" & iv == "Reading" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Posting" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+
+### Social media communication channels
+
+Regarding the COVID-19 related use of social media channels (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \@ref(fig:fig-within)).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in life satisfaction (_b_ `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% my_round("b")` [95% CI `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% round(2)`, `r data_tab_within %>% filter(dv == "Life satisfaction" & iv == "Facebook" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% round(2)`]).
+In sum, the hypothesis of trivial effects was supported also for the COVID-19 related use of important social media channels.
+
+That said, one effect differed substantially from zero.
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (_b_ = `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(estimate) %>% round(2)` [95% CI `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.low) %>% my_round("b")`, `r data_tab_within %>% filter(dv == "Negative affect" & iv == "Twitter" & Variance == "within" & Analysis == "Publication") %>% select(conf.high) %>% my_round("b")`]).
+However, the effect was still completely inside of the null region, hence likely not large enough to be considered meaningful.
+
+For an overview of all within-person effects, see Table \@ref(tab:tab-within) and Figure \@ref(fig:fig-within).
+
+```{r fig-within, fig.cap="Within-person effects of COVID-19 related social media use on well-being.
+Note.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_activity.pdf")
+Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=6, fig.height=4}
+fig_results_within
+# knitr::include_graphics("figures/fig_results_within.png")
 ```
 
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (_b_ `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_activity %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between") %>% select(conf.low) %>% my_round("b")`]).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
-
-Note that when comparing the results with and without control variables, the results differed.
-For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
-
-```{r res-channels, fig.cap="The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_channel.pdf")
+```{r tab-within, cache=F, warning=F, message=F, prompt=F}
+apa_table(tab_within %>% select(-Outcome),
+          align = c("l", "r", "r", "r", "r", "r"),
+          stub_indents = list(
+            "Life satisfaction" = c(1:8),
+            "Positive affect" = c(9:16),
+            "Negative affect" = c(17:24)),
+          caption = "Overview of all within-person effects.",
+          col_spanners = list(`Confidence interval` = c(3, 4)),
+          font_size = "footnotesize"
+          )
 ```
-
-Regarding the COVID-19 related use of social media _channels_, the results were comparable (see Figure \@ref(fig:res-channels)).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Facebook" & type == "within") %>% select(conf.high) %>% round(2)`]).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly _higher_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(estimate) %>% my_round("b")` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Instagram" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly _lower_ levels of life satisfaction than usual (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "Twitter" & type == "within") %>% select(conf.high) %>% my_round("b")`]).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant. 
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally _higher_ levels of life satisfaction (_b_ = `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.low) %>% my_round("b_txt")`, `r dat_fig_results_channel %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between") %>% select(conf.high) %>% my_round("b")`]).
-However, please note that this effect again was not large enough to be considered practically relevant.
-
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \@ref(fig:res-channels)).
 
@@
 
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
-Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here. 
-Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at _r_ = |.10|.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \@ref(fig:res-control).
+<!-- In what follows, I briefly report some exploratory analyses that weren't preregistered. -->
 
-```{r res-control, fig.cap="Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.", out.width = "\\textwidth", cache=F}
-knitr::include_graphics("figures/fig_results_control_std.pdf")
+<!-- ### Between-person relations -->
+
+<!-- For between-person relations, no a-priori hypotheses were formulated. -->
+<!-- Results showed that no relation crossed or was completely outside of the SESOI. -->
+<!-- Four relations were statistically significant. -->
+<!-- Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`]).
+-->
+<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->
+
+<!-- Note that when comparing the results with and without control variables, the results differed. -->
+<!-- For example, on the between-person level, one effect stopped being significant if controlled for additional variables. -->
+<!-- Initially, actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction. -->
+<!-- However, when controlling for potential confounders, the effect became virtually zero. -->
+
+<!-- For an overview of all between-person relations, see Figure \@ref(fig:fig-between). -->
+
+<!-- ### Covariates -->
+
+To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
+Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+Here, we can build on Cohen's convention that small effects begin at _r_ = |.10|.
+
+The results showed that several effects crossed or fell completely outside of the SESOI. 
+They can hence be considered meaningful.
+For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]), and negative affect ($\beta$ `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Health") %>% select(conf.high) %>% my_round("std")`]).
+Spending more time outside to exercise meaningfully increased positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Sports") %>% select(conf.high) %>% my_round("std")`]).
+Being more satisfied with the current democratic system meaningfully increased life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Satisfaction democracy") %>% select(conf.high) %>% my_round("std")`]).
+The strongest aspect affecting well-being was internal locus of control.
+If people felt more in control of their lives, this strongly increased both life satisfaction ($\beta$ `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Life satisfaction" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]) and
+positive affect ($\beta$ `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Positive affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]),
+while decreasing negative affect ($\beta$ `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(estimate) %>% my_round("std_txt")` [95% CI `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.low) %>% my_round("std")`, `r data_tab_comp %>% filter(dv == "Negative affect" & iv == "Internal locus of control") %>% select(conf.high) %>% my_round("std")`]).
+
+For an overview, see Figure \@ref(fig:fig-control).
+
+```{r fig-control, fig.cap = "Results of main variables together with covariates to provide context.
+All variables standardized.
+SESOI: beta = |.10|", out.width = "\\textwidth", cache=F, warning=F, message=F, fig.width=7, fig.height=8}
+fig_results_control_std
+# knitr::include_graphics("figures/fig_results_control_std.png")
 ```
-
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
-The results were virtually the same.
-For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of $\beta$ = |.10|.
-The additional analyses are reported in the [companion website](https://tdienlin.github.io/Austrian_Corona_Panel/analyses_additional.html).
 
@@
 
-In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
+Based on a panel study with 32 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
+Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being. 
+The within-person relations, which are more informative regarding causal effects, showed a different pattern, however.
+If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
+Several statistically significant effects were found, but these were very small.
+For example, people who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
+Posting more content about COVID-19 than usual slightly increased negative affect.
+People who read more COVID-19 related posts than usual reported slightly decreased positive affect.
+Using Twitter for COVID-19 related content slightly increased negative affect. 
+<!-- was related to Reading COVID-19 related content on social media slightly reduced positive affect. -->
+Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
+According to the preregistered procedure, they should hence be considered irrelevant. 
+Other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being. 
+In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being. 
+The results do not support the popular fears that "doomscrolling" or overusing social media during times of crises constitutes a relevant risk for well-being.
 
-The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that "doomscrolling" or overusing social media during times of crises don't seem to be justified.
+<!-- People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being. -->
 
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by @bendauAssociationsCOVID19Related2021, which showed negative relations between social media and well-being.
-However, note that @bendauAssociationsCOVID19Related2021 analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle. 
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter [@meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], which echoes the results obtained here.
+<!-- The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being. -->
+<!-- Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is rather determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
+-->
 
-If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-That said, all these effects were still very small and arguably too small to matter. 
-But future research might elaborate on these specific relations to probe their stability and relevance.
+These specific observations notwithstanding, several general trends can be observed.
+First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
+Although very small, four statistically significant negative results of COVID-19 related social media use on well-being were found.
+Not a single positive effect emerged.
+Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
+For example, it was controlled for whether or not participants contracted a COVID-19 infection during a specific wave.
+Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
 
-Finally, another interesting observation is that life satisfaction was remarkably stable.
-Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
-This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019].
+Second, significant outcomes emerged for positive or negative affect, but not for life satisfaction. 
+Life satisfaction was stable and not affected by any type or channel of social media communication.
+The more fluctuating positive and negative affect, however, were affected.
+Liking, sharing, and posting COVID-19 related content, and spending time on Twitter to browse COVID-19 related content, all negatively influenced affect. 
+This is aligned with prior findings which showed that social media can trigger negative affect but does not reduce life satisfaction [@huangTimeSpentSocial2017].
+<!-- Writing posts implies more substantial engagement, and perhaps also negative exchanges with other users, which could explain the results. -->
+<!-- For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual. -->
+<!-- Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect. -->
+<!-- As a potential explanation, when writing posts and comments on social media people explicitly and more deeply engage with COVID-19 related content. -->
+Conversations about COVID-19 on social media are often extreme, negative, or aggressive [@fanStigmatizationSocialMedia2020].
+More deeply engaging with this type of content could negatively affect active authors.
+The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter than usual increased negative affect.
+Communication on Twitter is often found to be more negative and impolite compared to other SNSs [@halpernSocialMediaCatalyst2013].
+Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
+
+Third, the results show that it makes sense to analyze different communication types and communication channels.
+Reading reduced positive affect, while liking, sharing and commenting increased negative affect.
+Whereas it was often stated that passive use is bad and active use good [@verduynPassiveFacebookUsage2015], this pattern thus did not emerge here.
+Instead, all three types were negative.
+The results hence support the findings from @valkenburgAssociationsActivePassive2022, who also could not confirm the claim that active use is good and that passive use is bad.
+Focusing on communication channels, Twitter seems to be more negative, as has often been observed [@halpernSocialMediaCatalyst2013].
+But, again, all of these effects are very small.
+Future research might elaborate on these specific relations to probe their stability and relevance.
+
+Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
+The findings support the Different Susceptibility to Media Effects Theory, which states that effects are generally small and depend on the type of communication.
+The results are well-aligned with mood management theory [@zillmannMoodManagementCommunication1988] or the uses and gratifications approach [@katzUsesGratificationsResearch1973], whose premises preclude particularly negative effects of routine and widespread media consumption.
+If the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
+Likewise, recent empirical studies and meta-analyses reported rather small negative effects, too.
+Several studies found that the effects of various types of social media use on well-being are small, often too small to matter [@bendauAssociationsCOVID19Related2021; @meierComputermediatedCommunicationSocial2020a; @orbenTeenagersScreensSocial2020; @fergusonThisMetaanalysisScreen2021], echoing the results obtained here.
+
+In light of the overall very small effects, engaging in COVID 19-related social media use should not be a major concern for one's well-being.
+That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use. 
+On average, it is likely more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter consuming COVID-19 related content.
+The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.
+
+<!-- First, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction. -->
+<!-- Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram [@waterlooNormsOnlineExpressions2018], which might help explain the results.
+-->
+<!-- Second, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
+-->
+<!-- On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent. -->
+
+<!-- Finally, another interesting observation is that life satisfaction was remarkably stable. -->
+<!-- Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly. -->
+<!-- This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes [@brownEasyHappinessPie2019]. -->
 
@@
 
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
-Together, this allows for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
-However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
+<!-- The current study analyzed whether changes in media use were related to changes in well-being. -->
+Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
+However, three major challenges remain. 
+First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals. 
+Next, it could be that there was not sufficient variance in media use and well-being. 
+Without sufficient variability, statistical models cannot detect existing effects.
+Third, it is necessary to control for _all_ relevant confounding third variables. 
+Although this study included are large list of confounders, it could still be that crucial variables were missed  [@zillmannMoodManagementCommunication1988].
+<!-- While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality. -->
+<!-- Causality necessitates temporal order, and the cause needs to precede the effect. -->
+<!-- The challenge is that regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds. -->
+<!-- In many cases only experience sampling studies asking users at the very moment can produce such knowledge. -->
+<!-- However, even then we don't know for certain if we actually measured the right interval. -->
+<!-- Effects depend on the intensity of use or the length of the interval. -->
+To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
+The results reported here are hence contingent on the design I used.
+To document how effects unfold, future research needs to employ different study designs probing different intervals.
+More thought needs to be invested in what factors to control for, and I hope this study provides a first step into this direction.
 
-Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
+Although I had already reduced the predefined SESOIs to be less conservative, one could argue they were still too large.
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it really realistic to expect that extremely changing only _one_ of these aspects should manifest in a detectable change in well-being?
+Is it realistic to expect that changing only _one_ of these aspects should already manifest in a detectable change in well-being?
 Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g.
 regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for well-being and _b_ = |.075| for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
-Again, with this study I hope to provide some first input and guidelines.
+Practically, this would imply a SESOI half the size defined here, namely _b_ = |.15| for life satisfaction and _b_ = |.075| for affect.
+In the case of this study, however, even halving the SESOI would not make a difference.
+All but one effect would still be completely in the null region, and no effect would fall completely outside of the null region.
+I encourage future research to elaborate on what effect sizes are considered meaningful and what not.
+<!-- Again, with this study I hope to provide some first input and guidelines. -->
 
 Both media use and well-being were measured using self-reports.
-Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures.
+Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
+However, for social media use objective measures are clearly preferable, as people often cannot reliably estimate their use [@scharkowAccuracySelfreportedInternet2016].
+At the same time, most objective measures cannot capture the content or the motivation of use.
+Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures.
+In many cases they can still be informative [@verbeijSelfreportedMeasuresSocial2021].
 
-Because the data were collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
-
+Being collected in a single country, the generalizability of the results is limited.
+The results apply primarily to the more Western sphere.
+They might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
+<!-- That said, being a comparatively large study largely representative of one country, with several waves collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+-->
 <!-- Social media use was measured with an ordinal variable, however in the analyses it was treated as a numerical one. -->
@@
 
-In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+In this study, COVID-19 related social media use did not meaningfully affect well-being. 
+Very small negative effects were found for writing COVID-19 related posts, sharing COVID-19 related content, and spending more time than usual on Twitter.
+Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 
@@
 
-\newpage
-
 # Competing Interests
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://tdienlin.github.io/Austrian_Corona_Panel/index.html).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
 
@@
 The data can only be used for scientific purposes.
+
+# Acknowledgements
+
+I would like to thank BLINDED for providing valuable feedback on this manuscript.

### 2023-01-30 manuscript.tex
- major updates to preprint

--- 2021-12-01 manuscript.tex
@@
 \documentclass[
-  english,
-  man,mask,floatsintext]{apa6}
+  man,floatsintext]{apa7}
 \usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={No meaningful effects of COVID-19 related social media use on well-being},
-  pdflang={en-EN},
-  pdfkeywords={COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 % \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+%   {\begin{landscape}\centering\begin{threeparttable}}
+%   {\end{threeparttable}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
 
@@
 \makeatother
-\keywords{COVID-19, Coronavirus, well-being, affect, life satisfaction, social media use, news use, communication, random effects within between model, panel study, longitudinal.}
+
+\usepackage{xpatch}
+\makeatletter
+\xapptocmd\appendix
+  {\xapptocmd\section
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {}{\InnerPatchFailed}%
+  }
+{}{\PatchFailed}
+\keywords{COVID-19, well-being, social media, news use, panel study.}
 \usepackage{lineno}
@@
 \usepackage{csquotes}
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
+\usepackage{endnotes}
 \setlength{\parskip}{0em}
@@
 \note{\clearpage}
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
-\ifluatex
+
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={No meaningful effects of COVID-19 related social media use on well-being},
+  pdfauthor={Tobias Dienlin1},
+  pdflang={en-EN},
+  pdfkeywords={COVID-19, well-being, social media, news use, panel study.},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 
-\abstract{
+\note{
+
+Preprint submitted for publication. Please cite carefully.
+
+}
+
+\abstract{%
 In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
 To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling.' In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
+However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling'.
+In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
 To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
@@
 
-During the COVID-19 pandemic,
-numerous events unfolded in quick succession and several open questions emerged.
+During the COVID-19 pandemic it was critical to stay informed regarding the latest developments.
 How dangerous is the virus?
-Is it spreading in my region?
-How is it transmitted, and how can I protect myself?
-Because for many it was (and still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
-Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
-In this extraordinary situation, many people used media excessively to attain information, and especially social media were at an all time high (Statista, 2021).
-
-Some people actually couldn't stop using social media to learn about COVID-19 related news.
-A new phenomenon emerged, termed ``doomscrolling'':
-Users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or looking after their children (Klein, 2021).
-As doomscrolling increased it became doubtful whether such social media is helpful, or whether it created an additional burden on mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
-These concerns seem justified:
-A study with 6,233 people from Germany conducted during the pandemic found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression'' (Bendau et al., 2021, p.
-283).
-
-As a result, with this study I want to build on this research and investigate whether or not COVID-19 related social media use affected well-being during the pandemic.
-To this end, I analyzed a large-scale panel study from the Austrian Corona Panel Project (Kittel et al., 2020).
-The panel consists of 24 waves and has an overall sample size of 3,018 participants.
-The panel study collected a large number of psychological and demographic variables.
-I explicitly aimed to investigate the causal effects of COVID-19 related social media use on well-being.
-
-\hypertarget{defining-well-being-and-media-use}{%
-\subsection{Defining Well-being and Media Use}\label{defining-well-being-and-media-use}}
-
-The underlying theories that guided the selection of variables for my analysis are the two-continua model of mental health (Greenspoon \& Saklofske, 2001) and the hierarchical taxonomy of computer-mediated communication (Meier \& Reinecke, 2020).
-According to the two-continua model, mental health consists of (a) psychopathology and (b) well-being.
-Well-being can be differentiated into subjective and psychological well-being (Diener, Lucas, \& Oishi, 2018).
-Whereas subjective well-being emphasizes hedonic aspects such as happiness and joy, psychological well-being addresses eudaimonic aspects such as fulfillment and meaning.
-Subjective well-being is primarily about achieving positive affect and avoiding negative affect.
-One of the most prominent indicators of well-being is life satisfaction.
-In my view, because it represents a general appraisal of one's life, life satisfaction is best thought of as a meta concept that combines psychological and subjective well-being.
-Notably, life satisfaction is stable and fluctuates only little, whereas it's the exact opposite for affect (Dienlin \& Johannes, 2020).
-To capture well-being in this study I thus build on life satisfaction, positive affect, and negative affect.
-Together, this should provide an encompassing perspective on potential media effects.
-
-The hierarchical taxonomy of computer-mediated communication differentiates six levels of how people engage with digital technology.
-First, the device (e.g., smartphone); second, the type of application (e.g., social networking site); third, the branded application (e.g., Twitter); fourth, the feature (e.g., status post); fifth, the interaction (e.g., one-to-many); and sixth, the message (e.g., content) (Meier \& Reinecke, 2020).
-Whereas the first four levels focus on the communication \emph{channel}, the last two address the communication \emph{type}.
-To measure social media use for the consumption of COVID-19 related news and topics, I here employ both the channel and the type of communication perspective, which together provides a nuanced understanding of communication.
-
-First, I investigate how different types of communication affect well-being.
-Specifically, I differentiate between active and passive use.
-I distinguish (a) \emph{reading} (passive), (b) \emph{posting} (active), and (c) \emph{liking and sharing} COVID-19 related posts (both active and passive).
-Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
-Branded apps are separate entities with potentially divergent effects.
-Twitter might have a different effect as compared to WhatsApp because of their respective affordances.
-For example, Waterloo, Baumgartner, Peter, and Valkenburg (2018) found that it's more adequate to express negative emotions on WhatsApp than on Twitter or on Instagram.
-The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube.
-Worth noting, this study is not about \emph{general} social media use during times of COVID, but on social media use \emph{focused} on COVID-19 related content.
-Examples of such media use include posting thoughts about the pandemic or retweeting COVID-19 related news.
-
-\hypertarget{effects-of-social-media-on-well-being}{%
-\subsection{Effects of Social Media on Well-Being}\label{effects-of-social-media-on-well-being}}
-
-So far, there is only little empirical research on COVID-19 related social media use on well-being.
-In their study on the relations between media use and mental health during the pandemic, Bendau et al.
-(2021) found that people who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}{]} and significantly more specific COVID-19 related anxiety symptoms'' (p.~288).
-Eden, Johnson, Reinecke, and Grady (2020) analyzed the media use of 425 US college students during the first wave of the pandemic, finding both positive and negative relations with well-being.
-In a sample of 312 respondents collected via Amazon Mechanical Turk, Choi and Choung (2021) reported that people who used media to attain information were more lonely and less satisfied with their lives.
-Stainback, Hearne, and Trieu (2020) analyzed a large-scale study with 11,537 respondents from the US and found that increased COVID-19 media consumption was related to more psychological distress.
-Together, the literature emphasizes potentially negative effects of social media as news use (see also Liu \& Tong, 2020; Riehm et al., 2020).
-However, note that all of these findings represent between-person relations stemming from cross-sectional data (see below).
-We therefore don't know whether the differences in mental health and well-being are due to social media use or due to other third variables, such as age, health, employment, or education.
-
-The question of whether and how social media use affects well-being \emph{in general}, on the other hand, is well-researched.
-This also holds true for the different types of communication such as active or passive use.
-A meta review (i.e., an analysis of meta-analyses) found that the relation between social media use and well-being is likely in the negative spectrum but very small (Meier \& Reinecke, 2020)---potentially too small to matter.
-What determines whether or not an effect is considered small or trivial?
-As a starting point, we could refer to standardized effect sizes.
-According to Cohen (1992), small effect sizes start at \emph{r} = .10.
-And indeed, several if not most of the current meta-analyses find effect sizes below that threshold (Ferguson et al., 2021; Huang, 2017; Meier \& Reinecke, 2020).
-
-These overviews are well aligned with several individual studies employing advanced methods (Keresteš \& Štulhofer, 2020; Orben, Dienlin, \& Przybylski, 2019; Przybylski, Nguyen, Law, \& Weinstein, 2021; Schemer, Masur, Geiß, Müller, \& Schäfer, 2021).
-For example, Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021) reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
-In conclusion, most effects are likely somewhere between trivial and small.
-I therefore expect that also in the case of COVID-19 related social media use effects will be trivial to small.
-
-From a theoretical perspective, how could we explain whether COVID-19 related social media use might affect well-being?
-In what follows, I outline potential arguments as to why the effect might be positive or negative, direct or indirect, or nonexistent.
-In advance, there doesn't seem to be a clear winner, and it's likely that both positive and negative effects cancel each other out.
-
-First, one could assume a \emph{direct} negative effect on well-being, and especially on positive or negative affect, which are more volatile and fluctuating.
-Dangers, inequalities, corruption---these were the headlines during the pandemic across many countries worldwide.
-If one learns about such events, the initial reaction might be shock, fear, or dismay.
-Repeatedly consuming such news can be depressing, perhaps even changing some general perspectives on life.
-That said, because not all news was negative, and because many people showed solidarity and compassion, there was also positive and uplifting content, potentially compensating for the negative effects.
-
-There could also be \emph{indirect} effects.
-When doomscrolling, users are captivated to such an extent that they cannot stop using social media.
-For example, during the pandemic social media use was at an all-time high in the US (Statista, 2021).
-In general, as has been expressed by many before, it is most likely that moderate social media use is not detrimental (Orben, 2020).
-Overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users (Przybylski \& Weinstein, 2017).
-To explain, overuse likely impairs well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
-So if a society collectively overuses social media when doomscrolling, there is potential for negative effects.
-
-On the other hand, one can make the case that overuse might also be beneficial, especially in times of a pandemic---even if the use is mainly COVID-19 related.
-Exchanging COVID-19 related messages with friends via WhatsApp might replace the in-person contact one would have otherwise, but which is literally impossible at the time.
-In situations where meaningful and functional activities are prohibited, using social media to exchange about COVID-19 related topics might not be the worst idea.
-Besides, given that nowadays a large number of experts, scientists, and politicians converse directly on social media, one can get first-hand high quality information on current developments.
-
-Together, the strongest argument to me is that \emph{in general} the effects of social media on well-being are, on average, small at best.
-Because this study only looks at \emph{one part} of social media use---namely, COVID-19 related interactions---it is very focused, diminishing the overall potential of the effects even further.
-Whether or not using social media for COVID-19 related aspects is detrimental during a pandemic is also not entirely clear.
-Therefore, I expect that COVID-19 related communication on social media doesn't affect well-being in a meaningful or relevant way.
+In what region is it spreading?
+How is it transmitted? What are the current safety regulations?
+To obtain relevant information, many people heavily relied on social media, which were at an all time high (Statista, 2021).
+
+Some actually could not stop using social media to learn about COVID-19 related news.
+A new phenomenon termed ``doomscrolling'' emerged.
+Many users were glued to their screens and found it hard to pursue other relevant activities such as working, taking a break, or even looking after their children (Klein, 2021).
+In the media it was hence increasingly asked whether using social media for COVID-19 related reasons would, next to all other stressors, create an additional burden on mental health (Sandstrom et al., 2021).
+Although research has begun addressing this question (Bendau et al., 2021; Bradley \& Howard, 2021; Choi \& Choung, 2021; Dörnemann et al., 2021; Eden et al., 2020; Guazzini et al., 2022; Latikka et al., 2022; Liu \& Tong, 2020; Riehm et al., 2020; Sewall et al., 2021; Stainback et al., 2020; Yue et al., 2022), it still largely unknown if COVID-19 related social media use during the pandemic has had a meaningful impact on well-being.
+By analyzing large-scale panel data from Austria and by focusing on intra-individual causal effects, this study aims to answer to this question.
+
+\hypertarget{understanding-well-being-and-media-use}{%
+\subsection{Understanding Well-being and Media Use}\label{understanding-well-being-and-media-use}}
+
+This study investigates how different \emph{facets} of well-being are affected by different \emph{types} and different \emph{channels} of communication (Meier \& Reinecke, 2020).
+Building on the typology of subjective well-being (Diener et al., 2018), three different well-being facets are analyzed: life satisfaction, positive affect, and negative affect.
+Because effects of social media depend on how they are used (Verduyn et al., 2015), I further distinguish three types of use and five popular channels.
+The types of use include reading, posting, and liking and sharing COVID-19 related posts.
+In doing so, this study analyzes social media use focused on COVID-19 related content, which includes posting thoughts about the pandemic, reading posts and comments, or retweeting COVID-19 related news.
+The five channels to be investigated are Facebook, Twitter, Instagram, WhatsApp, and YouTube, which at the time ranked among the most popular social media services in Austria.
+
+How could the various types and channels of COVID-19 related social media use affect well-being?
+According to the set-point theory, well-being is surprisingly stable (Lykken, 1999).
+Although specific events such as marriage or salary can have significant impacts on well-being, in most cases effects are only short-term, with well-being after some time routinely returning to prior levels (Sheldon \& Lucas, 2014).
+Only very specific factors such as unemployment, disability, or death can cause long-term changes in well-being (Lucas, 2007).
+Can media use be such a factor?
+Current literature overviews suggest no:
+Although social media use on average decreases well-being, the effects are small (Meier \& Reinecke, 2020).
+In addition, the effects of media use differ across individuals and types of content (Valkenburg \& Peter, 2013).
+Whereas for some users social media are more beneficial, for others they are more harmful.
+For example, in one study it was estimated that roughly one quarter of all users experienced negative effects, another quarter positive effects, while for the rest the effects were neutral (Beyens et al., 2021).
+This finding is aligned with the Differential Susceptibility to Media Effects Model:
+Although there is substantial \emph{variation} of media effects for individual users, there are both negative and positive aspects, leading to small effects on average (Valkenburg \& Peter, 2013)---as confirmed by several literature overviews (Dienlin \& Johannes, 2020; Huang, 2017; Meier \& Reinecke, 2020; Orben, 2020).
+
+Why are the effects of social media use on well-being small on average?
+Using Social media is ambivalent.
+It can impair well-being when causing embarrassment, stress, or disinformation, and it can improve well-being when providing connectedness, information, or entertainment (Büchi, 2021).
+Although social media are often associated with negative outcomes, there are a large number of positive effects, which either offset or at least significantly reduce negative outcomes.
+Positive outcomes encompass finding relevant information; maintaining and fostering relationships; expressing one's personality; or entertaining oneself (e.g., Pelletier et al., 2020).
+Negative effects include distraction; displacement of other meaningful activities; consumption of shallow, misleading or challenging content; or negative social comparisons (e.g., Meier \& Krause, 2022).
+People are much more likely to use media that offer plentiful benefits (Katz et al., 1973).
+Users implicitly learn what media help them regulate their mood and thereby well-being (Zillmann, 1988).
+Given the ubiquitous use of social media, the many benefits they offer, plus the general and implicit competence to use media that foster mood management, it is hence unlikely that the effects of social media use on well-being are profusely negative.
+
+There is still little empirical research on how well-being is affected by social media use that is focused on COVID-19 specifically.
+Results are mixed.
+When browsing social media for COVID-19 related news, many users reported being captivated to such an extent they could not stop using social media (Klein, 2021).
+In a study with 1,131 residents from Wuhan in China, people who spent more time in quarantine also spent more time on social media (Yue et al., 2022).
+Those who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.
+People who used social media as a primary source of information reported on average ``significantly more unspecific anxiety and depression {[}\ldots{]} and significantly more specific COVID-19 related anxiety symptoms'' (Bendau et al., 2021, p.
+288).
+Eden et al. (2020) analyzed how 425 US college students used media during the first wave of the pandemic, finding both positive and negative relations with well-being.
+In a large-scale study with 11,537 respondents from the US, increased COVID-19-related media consumption was related to more psychological distress (Stainback et al., 2020).
+A four-wave panel study with 384 young adults from the U.S.
+analyzed the effects of general digital technology use during the pandemic on mental health, finding that digital technology did not have significant effects on mental health (Sewall et al., 2021; for a similar study with comparable results, see Bradley \& Howard, 2021).
+A study with 2.057 respondents from Italy reported that during the pandemic virtual community and social connectedness increased (Guazzini et al., 2022).
+In a study with 735 participants from Finland, levels of loneliness were stable during the pandemic, and people who engaged more on social media experienced less loneliness (Latikka et al., 2022).
+Together, the literature is mixed, showing both positive and negative affects of social media use focused on or during COVID-19 on well-being (see also Dörnemann et al., 2021; Liu \& Tong, 2020; Riehm et al., 2020).
+
+In conclusion, these mixed empirical results, together with the observation that social media effects on well-being are very small in general, I expect that COVID-19 related communication on social media does not affect well-being in a meaningful or relevant way.
 
@@
 
-\hypertarget{current-study}{%
-\section{Current Study}\label{current-study}}
+\hypertarget{method}{%
+\section{Method}\label{method}}
+
+\hypertarget{preregistration}{%
+\subsection{Preregistration}\label{preregistration}}
+
+The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, \emph{p}-value) were preregistered on the Open Science Framework, accessible here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
+Because in this study I analyze data from an already existing large-scale data set, the preregistration was done prior to accessing the data.
+The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
+In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
+
+\hypertarget{sample}{%
+\subsection{Sample}\label{sample}}
+
+The data come from the Austrian Corona Panel Project (Kittel et al., 2021), which is a large-scale standalone panel study.
+The data are hosted on AUSSDA, are publicly available here (\url{https://doi.org/10.11587/28KQNS}), and consist of 32 waves.
+The study was conducted between March 2020 and June 2022, and data collection is now finished.
+Between March 2020 and July 2020, the intervals between waves were weekly and afterward monthly.
+Each wave consists of at least 1,500 respondents.
+The sample size was \emph{N} = 3,485, with overall 111,520 observations.
+Panel mortality was compensated through a continuous re-acquisition of new participants.
+Participants were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
+Panel members were incentivized with 180 credit points for each wave of the study.
+
+Achieved via quota sampling, the sample matched the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
+In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
+All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
+Ethical review and approval was not required for the study in accordance with the local legislation and institutional requirements.
+The participants provided their written informed consent to participate in this study.
+The average age was 41 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
 
@@
 
-Testing this hypothesis, however, is not trivial.
-First, in contrast to most hypotheses typically posited in the social sciences it implicitly contains an effect size, a so-called smallest effect size of interest (SESOI).
-Effectively testing this hypothesis necessitates defining what's considered a ``trivial effect size'' and what's not.
-Above I already referred to standardized effect sizes.
-However, standardized effect sizes should only be a first step toward evaluating an effect's relevance (Baguley, 2009).
-Standardized effect sizes are determined by a sample's variance,\footnote{Consider the effect size Cohen's \emph{d}: The mean's of the two groups that are to be compared are subtracted from one another and then divided by the sample's standard deviation (Cohen, 1992).
-Hence, if there is more deviation/variance in a sample, the effect size decreases, even if the difference of the group's means stays the same.} which is problematic:
-The question of whether or not social media use affects a particular person in a relevant way should not depend on the variance in the sample in which that person's data were collected.
-Instead, it should depend on absolute criteria.
-
-What could be a minimally interesting, nontrivial effect?
-Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer.
-However, it is still necessary and helpful to try provide such a plausible benchmark.
-I therefore suggest the following SESOI for this research question:
+Testing the hypothesis necessitates defining what is considered a ``trivial effect size''.
+Being a normative question, finding a clear, single, or unanimous answer is impossible.
+However, it is still necessary and helpful to work toward a plausible benchmark.
+I suggest the following SESOI for this research question:
 
@@
 Thus, a change of four units in social media use (e.g., a complete stop) should correspond to a noticeable change in well-being.
-But what's a noticeable change in well-being?
-According to Norman, Sloan, and Wyrwich (2003), people can reliably distinguish \emph{seven} levels of satisfaction with health.
-So if satisfaction is measured on a 7-point scale, we would state that a four unit change in social media use should result in a one unit change in life satisfaction.
-(For more information, see Methods section ``Inference Criteria.'')
-
-\hypertarget{causality}{%
-\subsection{Causality}\label{causality}}
-
-The hypothesis explicitly states a causal effect.
-In non-experimental studies, longitudinal designs can help investigate causality.
-Using longitudinal designs alone, however, is not sufficient for establishing correct causal statements (Rohrer \& Murayama, 2021).
-In addition, we for example also need to control for confounding third variables, and importantly also for \emph{varying} third variables.
-
-To illustrate, consider the following example.
-Imagine that a person suddenly starts using social media much more than usual, and then after some time becomes less satisfied with their life.
-Eventually, use and life satisfaction return to prior levels.
-If this happens to several people at the same time, in a longitudinal study we could then observe a significant effect of social media use on life satisfaction.
-However, it could also be the case that during the study there was a major exogenous event (say, a pandemic), which caused large parts of the working population to loose their jobs.
-Hence, the causal effect reported above was confounded, because in reality it was the pandemic that caused both social media use to rise and life satisfaction to go down.
-
-Thus, only when controlling for \emph{all} relevant confounders, we can correctly estimate causality without bias (Rohrer, 2018).
-Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
-In addition, when determining the overall causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
-Complicating matters further, it is often unclear if a variable is a mediator or a confounder.\footnote{In addition, there also exist colliders, which I don't discuss here and which complicate the issue even further (Rohrer, 2018).}
-However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
-The aim should therefore be to collect as many varying and nonvarying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.
-
-When searching for suitable candidates for confounders, we should look for variables that affect both media use and well-being.
-Controlling for these factors isolates the actual effect of social media use on well-being.
-We can also control for variables that affect only social media use or well-being.
-However, in doing so not much is gained or lost, because the effects of social media use would remain virtually the same (Kline, 2016; but see McElreath, 2021).
-
-In this study, I hence plan to control for the following variables, which either have already been shown to affect both social media use and well-being or which are likely to do so, and which also aren't mediators:
-gender, age, education, Austria country of birth, Austria country of birth of parents, text-based news consumption, video-based news consumption, residency Vienna, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
-I will not control for variables such as trust in institutions or trust in media, because these variables might be influenced by social media use to a meaningful extent.
-
-Next to including covariates, it's now increasingly understood that causal effects should be analyzed from an internal, within-person perspective (Hamaker, 2014).
-If a specific person changes their media diet, we need to measure how this behavior affects \emph{their own} well-being.
-Between-person comparisons from cross-sectional data, where participants are interviewed only once, cannot provide such insights.
-In this study, I hence differentiate between-person relations from within-person effects.
-And as explicated above, to test the hypothesis I thus consider only the within-person effects.
-
-Finally, one precondition of causality is temporal order.
-The cause needs to precede the effect.
-Finding the right interval between cause and effect is crucial.
-For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
-If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
-In the case of well-being, we need shorter intervals for affect and longer ones for life satisfaction.
-Still, choosing the right interval is challenging, because especially short intervals are hard to implement in practice and often require advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) (Schnauber-Stockmann \& Karnowski, 2020).
-
-In this study, I therefore adopt an intermediate perspective.
-I analyze if when a person changes their social media diet, are there \emph{simultaneous} changes in their well-being?
-When additionally controlling for both stable and varying confounders, we can then be more sure that the effect is indeed causal.
-
-\hypertarget{method}{%
-\section{Method}\label{method}}
-
-In this section I describe the preregistration and how I determined the sample size, data exclusions, the analyses, and all measures in the study.
-
-\hypertarget{preregistration}{%
-\subsection{Preregistration}\label{preregistration}}
-
-The hypotheses, the sample, the measures, the analyses, and the inference criteria (SESOI, p-value) were preregistered on the Open Science Framework.
-The (anonymous) preregistration can be accessed here: \url{https://osf.io/87b24/?view_only=b2289b6fec214fa88ee75a18d45c18f3}.
-Because in this study I analyzed data from an already existing large-scale data set, all of these steps were done prior to accessing the data.
-The preregistration was designed on the basis of the panel documentation online (Kittel et al., 2020).
-In some cases I couldn't execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when inspecting the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website} (\url{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}).
-
-\hypertarget{sample}{%
-\subsection{Sample}\label{sample}}
-
-The data come from the Austrian Corona Panel Project (Kittel et al., 2021).
-The study was conducted between March 2020 and October 2021.
-It contains 26 waves, and at the time of writing the first 24 waves were available for download.
-Each wave consists of at least 1,500 respondents.
-The overall sample size was \emph{N} = 3,018, and 72,432 observations were collected.
-Panel mortality was compensated through a continuous acquisition of new participants.
-All respondents needed to have access to the internet (via computer or mobile devices such as smartphones or tablets).
-They were sampled from a pre-existing online access panel provided by the company Marketagent, Austria.
-Respondents were asked and incentivized with 180 credit points to participate in each wave of the panel.
-
-The sample was representative of the Austrian population in terms of age, gender, region/state, municipality size, and educational level.
-In order to participate in the study, the respondents needed to be Austrian residents and had to be at least 14 years of age.
-The study received IRB approval from the University of Vienna.
-The average age was 42 years, 49 percent were male, 14 percent had a University degree, and 5 percent were currently unemployed.
-
-\hypertarget{inference-criteria}{%
-\subsection{Inference Criteria}\label{inference-criteria}}
-
-Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
-The sample is large, and it is hence well-equipped reliably to detect also small effects.
-In addition, because such large samples easily generate significant \emph{p}-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
-To this end, I adopted the interval testing approach as proposed by Dienes (2014).
-On the basis of the SESOI, I then defined a null region.
-In what follows, I explain how I determined the SESOI and the null region.
+What is a noticeable change in well-being?
+According to Norman et al. (2003), people can reliably distinguish seven levels of satisfaction with health.
+So if satisfaction is measured on a 7-point scale, a four unit change in social media use should result in a one unit change in life satisfaction.
 
 In this study, life satisfaction was measured on an 11-point scale.
-If people can reliably differentiate 7 levels as mentioned above, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
+If people can reliably differentiate 7 levels, this corresponds to 11 / 7 = 1.57 unit change on an 11-point scale.
 Hence, a four-point change in media use (e.g., a complete stop) should result in a 1.57-point change in life satisfaction.
 In a statistical regression analysis, \emph{b} estimates the change in the dependent variable if the independent variable increases by one point.
-We would therefore expect a SESOI of \emph{b} = 1.57 / 4 = 0.39.
-For affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
-Because we're agnostic as to whether the effects are positive or negative, the null region includes negative and positive effects.
-Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).}
+For life satisfaction, we would therefore define a SESOI of \emph{b} = 1.57 / 4 = 0.39.
+For positive or negative affect, which was measured on a 5-point scale, our SESOI would be \emph{b} = 0.71 / 4 = 0.18.
+Because we are agnostic as to whether the effects are positive or negative, the null region includes both negative and positive effects.
+Finally, in order not to exaggerate precision and to be less conservative, these numbers are reduced to nearby thresholds.\footnote{Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens et al., 2021; Funder \& Ozer, 2019).}
 Together, this leads to a null region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.
 
-Let's briefly illustrate what this means in practice.
-If the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = .20, {[}95\% CI: .15, .25{]}), the hypothesis that the effect is trivial is supported.
-If the confidence interval and the null region overlap (e.g., \emph{b} = .30, {[}95\% CI: .25, .35{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
-If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = .40, {[}95\% CI: .35, .45{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+The hypothesis is analyzed using the interval testing approach as proposed by Dienes (2014).
+To illustrate, if the 95\% confidence interval falls completely within the null-region (e.g., \emph{b} = -.02, {[}95\% CI: -.12, .08{]}), the hypothesis that the effect is trivial is supported.
+If the confidence interval and the null region overlap (e.g., \emph{b} = -.22, {[}95\% CI: -.27, -.17{]}), the hypothesis is not supported and the results are considered inconclusive, while a meaningful negative effect is rejected.
+If the confidence interval falls completely outside of the null-region (e.g., \emph{b} = -.40, {[}95\% CI: -.45, -.35{]}), the hypothesis is rejected and the existence of a meaningful positive effect is supported.
+For an illustration, see Figure \ref{fig:sesoi}.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/sesoi-1.pdf}
+\caption{\label{fig:sesoi}Using confidence intervals to test a null region. In this study, a trivial effect of social media use on life satisfaction is defined as ranging from b = -.30 to b = .30}
+\end{figure}
 
@@
 
-The hypothesis was analyzed using mixed effects models, namely random effect within-between models (REWB)(Bell, Fairbrother, \& Jones, 2019).
-Three models were run, one for each dependent variable.
+\hypertarget{causality}{%
+\subsubsection{Causality}\label{causality}}
+
+When using longitudinal designs to analyze causality, it is important to (a) focus on within-person effects (Hamaker, 2014); to (b) control for confounders (Rohrer \& Murayama, 2021); and to (c) test a plausible interval between measures (Dormann \& Griffin, 2015).
+To elaborate, in non-experimental designs it makes much sense to analyze causal effects from an internal, within-person perspective (Hamaker, 2014).
+If a specific person changes their media diet, we need to measure how this behavior affects their well-being.
+Between-person comparisons from longitudinal data cannot provide such insights (Lucas, 2022).
+To test the hypothesis, I thus consider only the within-person effects.
+
+Second, to identify confounders we should control for variables that affect both media use and well-being, which helps isolate the actual effect of social media use on well-being (Rohrer, 2018).
+Because we are adopting a within-person perspective, we need to implement \emph{time-varying} confounders (Rohrer \& Murayama, 2021).
+And because we are determining the \emph{overall} causal effect, we need to make sure \emph{not} to control for mediating variables (Rohrer, 2018), for doing so would bias our assessment of the causal effect.
+In this study, I hence preregistered to control for the following variables, which either have already been shown or are very likely to affect both social media use and well-being, and which also are not mediators:
+gender, age, education, Austria country of birth, Austria country of birth of parents, residency Vienna, text-based news consumption, video-based news consumption, household size, health, living space, access to garden, access to balcony, employment, work hours per week, being in home-office, household income, outdoor activities, satisfaction with democracy, disposition to take risks, and locus of control.
+
+Finally, one precondition of causality is temporal order and finding a plausible interval (Dormann \& Griffin, 2015).
+If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals are required.
+In the case of well-being, we need shorter intervals for the more fluctuating positive and negative affect, and longer ones for the more stable life satisfaction (Dienlin \& Johannes, 2020).
+Using social media can have instant effects on mood (Marciano et al., 2022).
+Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then in turn affect life satisfaction (Dienlin et al., 2017).
+In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
+In other words, if people during the last week engaged in more COVID-19 related social media use than usual, did they feel better or worse during that week than usual?
+For life satisfaction, I implemented a longer interval.
+If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
+I hence analyze if when a person changes their social media diet, are there (a) \emph{simultaneous} changes in their affect and (b) \emph{subsequent} changes in their life satisfaction?
+These relations will be controlled for varying confounders, which fosters a causal interpretation.
+
+\hypertarget{statistical-model}{%
+\subsubsection{Statistical model}\label{statistical-model}}
+
+The hypothesis was analyzed using random effect within-between models (REWB, Bell et al., 2019).
+Altogether three models were run, one for each dependent variable.
 The data were hierarchical, and responses were separately nested in participants and waves (i.e., participants and waves were implemented as random effects).
@@
 Because Mardia's test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
-To avoid overfitting, I tested the scales on more liberal fit criteria (CFI \textgreater{} .90, TLI \textgreater{} .90, RMSEA \textless. .10, SRMR \textless{} .10) (Kline, 2016).
-Finally, REWB-models cannot model latent variables.
-To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect.
-Respondents who answered less than 50\% of all questions were removed.
-The remaining missing responses were imputed using predictive mean matching.
-
-For more information on the analyses, a complete documentation of the models and results, see \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website}.
+Mean scores were used for positive and negative affect.
+Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.
+All variables were imputed except the social media use measures, as they were not collected on each wave.
+All variables included in the analyses presented here were used to impute missing data.
+
+\hypertarget{robustness-checks}{%
+\subsubsection{Robustness-checks}\label{robustness-checks}}
+
+To find out whether the inferences were robust across plausible but subordinate analyses, I conducted additional exploratory analyses.
+I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
+For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
@@
 
-In what follows, I list all the variables that I analyzed.
-For the variables' means, range, and variance, see Table \ref{tab:descriptives}.
-For a complete list of all items and item characteristics, see \href{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}{companion website}.
+For the variables' means, range, and variance, see Table \ref{tab:tab-descriptives}.
+For a complete list of all items and item characteristics, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
 
@@
 
-Life satisfaction was measured with the item ``Taken everything together, how satisfied are you currently with your life?''
-The response options ranged from 0 (\emph{extremely unsatisfied}) to 10 (\emph{extremely satisfied}).
-
-To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
+Life satisfaction was measured with the item ``All things considered, how satisfied are you with your life as a whole nowadays?'', which comes from the European Social Survey (European Social Survey, 2021).
+The response options ranged from 0 (\emph{extremely dissatisfied}) to 10 (\emph{extremely satisfied}).
+
+To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy (World Health Organization, 1998).
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \textit{p} = .032, cfi = 1.00, rmsea = .02, 90\% CI {[}.01, .03{]}, srmr = .01.
+The scale showed good factorial fit, \(\chi^2\)(62) = 67.03, \emph{p} = .309, CFI = 1.00, RMSEA \textless{} .01, 90\% CI {[}\textless{} .01, .02{]}, SRMR = .01.
 Reliability was high, \(\omega\) = .85.
 
-For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
+For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad (World Health Organization, 1998).
 The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
-The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \textit{p} \textless{} .001, cfi = .97, rmsea = .08, 90\% CI {[}.07, .08{]}, srmr = .03.
-Reliability was high, \(\omega\) = .89.
+The scale showed good factorial fit, \(\chi^2\)(443) = 3810.53, \emph{p} \textless{} .001, CFI = .98, RMSEA = .07, 90\% CI {[}.07, .08{]}, SRMR = .03.
+Reliability was high, \(\omega\) = .91.
 
@@
 COVID-19 related social media use focused on communication types was measured with the three dimensions of (a) reading, (b) liking and sharing, and (c) posting.
-The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''
-The three items were ``Reading the posts of others with content on the Coronavirus,'' ``When seeing posts on the Coronavirus, I clicked `like,' `share' or `retweet','' ``I myself wrote posts on the Coronavirus on social media.''
+The items come from Wagner et al. (2018) and were adapted for the context of this study.
+The general introductory question was ``How often during the last week have you engaged in the following activities on social media?''.
+The three items were ``Reading the posts of others with content on the Coronavirus'', ``When seeing posts on the Coronavirus, I clicked `like', `share' or `retweet'\,'', ``I myself wrote posts on the Coronavirus on social media.''
 Answer options were 1 (\emph{several times per day}), 2 (\emph{daily}), 3 (\emph{several times per week}), 4 (\emph{weekly}), 5 (\emph{never}).
@@
 
-COVID-19 related social media use focused on channels was measured with five variables.
+COVID-19 related social media use focused on channels was measured with five variables from Wagner et al. (2018), adapted for this study.
 The general introductory question was ``How often in the last week have you followed information related to the Corona-crisis on the following social media?''
@@
 
-Social media use was measured for all participants on waves 1, 2, 8, 17, and 23.
-Freshly recruited respondents always answered all questions on social media use.
+Social media use was measured for all participants on waves 1, 2, 8, 17, 23, and 28.
+Freshly recruited respondents always answered all questions on COVID 19-related social media use.
+For the analyses I hence used all 32 waves.
 
@@
 
-\caption{\label{tab:descriptives}Descriptives of the main variables.}
+\caption{\label{tab:tab-descriptives}Descriptives of the main variables.}
 
@@
 Well-being &  &  &  & \\
-\ \ \ Life satisfaction & 1.68 & 6.38 & 6.81 & 6.59\\
-\ \ \ Positive affect & 0.57 & 3.05 & 3.28 & 3.15\\
-\ \ \ Negative affect & 0.39 & 1.66 & 1.81 & 1.73\\
+\ \ \ Life satisfaction & 2.22 & 6.37 & 6.59 & 6.51\\
+\ \ \ Positive affect & 0.94 & 3.11 & 3.22 & 3.16\\
+\ \ \ Negative affect & 0.76 & 1.74 & 1.87 & 1.79\\
 Social media use &  &  &  & \\
-\ \ \ Read & 1.03 & 2.10 & 2.92 & 2.43\\
-\ \ \ Like \& share & 0.87 & 1.61 & 1.92 & 1.78\\
-\ \ \ Posting & 0.63 & 1.33 & 1.46 & 1.39\\
+\ \ \ Read & 1.39 & 2.02 & 2.87 & 2.37\\
+\ \ \ Like \& share & 1.19 & 1.56 & 1.95 & 1.74\\
+\ \ \ Posting & 0.89 & 1.37 & 1.42 & 1.40\\
 Social media channel &  &  &  & \\
-\ \ \ Facebook & 0.96 & 2.34 & 2.68 & 2.45\\
-\ \ \ Twitter & 0.52 & 1.16 & 1.72 & 1.37\\
-\ \ \ Instagram & 0.83 & 1.84 & 2.66 & 2.09\\
-\ \ \ WhatsApp & 1.23 & 2.28 & 2.62 & 2.45\\
-\ \ \ YouTube & 0.88 & 1.77 & 2.30 & 2.00\\
+\ \ \ Facebook & 1.60 & 2.21 & 2.74 & 2.43\\
+\ \ \ Twitter & 0.95 & 1.32 & 1.45 & 1.38\\
+\ \ \ Instagram & 1.33 & 2.04 & 2.05 & 2.05\\
+\ \ \ WhatsApp & 1.67 & 2.32 & 2.65 & 2.46\\
+\ \ \ YouTube & 1.28 & 1.90 & 2.02 & 1.96\\
 \bottomrule
@@
 The effects of COVID-19 related social media use were controlled for the following stable variables:
-(a) gender (female, male, diverse), (b) age, (c) education (ten options), (d) Austria country of birth (yes/no), (e) Austria parents' country of birth (no parent, one parent, both parents).
-I originally planned to implement additional variables as varying covariates.
-However, because they were not measured often enough or not at the time when social media use was measured, I implemented them as stable variables using their average values across all waves.
-This includes (a) text-based media news consumption (five degrees), (b) video-based media news consumption (five degrees), (c) residency is Vienna (yes/no), (d) self-reported physical health (five degrees), (e) living space (eleven options), (f) access to balcony (yes/no), (g) access to garden (yes/no), (h) employment (nine options), (i) disposition to take risks (eleven degrees), and (j) locus of control (five degrees).
-I also controlled for the following varying covariates: (a) five items measuring outdoor activities such as sport or meeting friends (five degrees), and (b) satisfaction with democracy (five degrees).
-Because it lead to too much attrition in the sample, I did not control for (a) household size, (b) work hours per week, (c) home office, (d) household income.
+gender (female, male, diverse), age, education (ten options), Austria country of birth (yes/no), Austria parents' country of birth (no parent, one parent, both parents), and household size.
+I also controlled for the following varying covariates: five items on current living conditions, including self-reported physical health, whether participants contracted COVID-19 since the last wave, current household income, working in home-office, and overall work hours; nine items measuring use of specific national text-based and video-based news outlets; five items measuring outdoor activities such as exercise or meeting friends; and three more psychological measures including locus of control, disposition to take risks, and satisfaction with democracy.
 
@@
 
-First, when looking at the variables from a descriptive perspective, we see that all well-being measures did not change substantially across the different waves of data collection.
-COVID-19 related media use, however, decreased slightly at the beginning of the study and remained stable after approximately six waves.
-The initial decrease might be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic began.
-It could be that after an initial uptick, COVID-19 related social media use was already declining at the time, returning to more normal levels.
+\hypertarget{descriptive-analyses}{%
+\subsection{Descriptive Analyses}\label{descriptive-analyses}}
+
+Looking at the variables from a descriptive perspective, aligned with set-point theory we can see that the level of all well-being measures were surprisingly stable during data collection (see Figure \ref{fig:fig-descriptives}).
+COVID-19 related social media use, however, showed changes.
+Reading, sharing and liking COVID-19 related content decreased substantially (almost one scale point from 3 to 2).
+Posting about COVID-19 related content stayed the same.
+Using Facebook, WhatsApp, and YouTube for COVID-19 related content decreased.
+Instagram stayed the same.
+Twitter use slightly increased.
+The general initial decrease could be explained by the fact that the collection of data began at the end of March 2020, hence approximately three months after the pandemic's onset.
+After an initial uptick, COVID-19 related social media use might have already been declining at the time.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Development of well-being and media use measures across the pandemic.
-Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-desc}
+\includegraphics[width=\textwidth]{figures/fig_descriptives} \caption{Well-being and media use across the 32 waves.
+Note.
+Values obtained from mixed effect models, with participants and waves as grouping factors and without additional predictors.}\label{fig:fig-descriptives}
+\end{figure}
+
+Using the average values across all waves, which provides a stable picture of the general relations, I next looked at the correlations between social media use and well-being (see Figure \ref{fig:fig-correlations}).
+Several interesting patterns emerged.
+In general, people who spend more time engaging with COVID-19 related content on social media reported reduced well-being.
+Users who spend more time reading, liking and sharing, and posting COVID-19 related content were less satisfied with their lives.
+They also showed slightly less positive affect.
+This overall negative picture was even more pronounced for negative affect.
+People who engaged more with COVID-19 related content, including all types and channels of communication, reported substantially higher levels of negative affect.
+For example, people who were more likely to post COVID-19 content had much higher levels of negative affect (\emph{r} = .56).
+Note that these results represent between-person correlations, not causal within-person effects.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/fig_cor} \caption{Descriptives of the main variables, capturing well-being and social media use with their average values across all waves.
+Upper triangle: correlation coefficients; diagonal: density plots; lower triangle: scatter plots.}\label{fig:fig-correlations}
 \end{figure}
@@
 
-The study's main hypothesis was that the effects of social media use on well-being would be trivial.
-Regarding the effects of different communication \emph{types}---that is, reading vs.~sharing vs.~posting---all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:res-activity}).
-For example, respondents who used social media more frequently than usual to read about COVID-19 related topics did not show a simultaneous change in life satisfaction (\emph{b} = 0.05 {[}95\% CI -0.01, 0.1{]}).
-All confidence intervals included zero; hence, all effects were also statistically non-significant.
-As a result, the hypothesis was supported for all COVID-19 related types of social media communication.
+\hypertarget{social-media-communication-types}{%
+\subsubsection{Social media communication types}\label{social-media-communication-types}}
+
+The study's main hypothesis was that the causal effects of all types and channels of social media use on all facets of well-being would be trivial.
+Regarding the effects of different communication types (i.e., reading, sharing, of posting about COVID-19 related content), all within-person effects fell completely within the a-priori defined null region (see Figure \ref{fig:fig-within}).
+For example, respondents who used social media more frequently than usual to like or share COVID-19 related content did not show a simultaneous change in life satisfaction (\emph{b} = -0.02 {[}95\% CI -0.08, 0.03{]}).
+As a result, the hypothesis of trivial effects was supported for all COVID-19 related types of social media communication.
+
+However, three effects stood out, as they were significantly different from zero.
+Users who read more COVID-19 related content than usual reported slightly reduced levels of positive affect (\emph{b} = -0.03 {[}95\% CI -0.05, -0.01{]}).
+Users who liked and shared more COVID-19 related content than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.02, 0.08{]}).
+Users who wrote more COVID-19 related posts than usual also experienced slightly more negative affect than usual (\emph{b} = 0.05 {[}95\% CI 0.02, 0.08{]}).
+
+\hypertarget{social-media-communication-channels}{%
+\subsubsection{Social media communication channels}\label{social-media-communication-channels}}
+
+Regarding the COVID-19 related use of social media channels (i.e., Facebook, Instagram, WhatsApp, YouTube, and Twitter) the results were comparable (see Figure \ref{fig:fig-within}).
+Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
+For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in life satisfaction (\emph{b} \textless{} 0.01 {[}95\% CI -0.04, 0.04{]}).
+In sum, the hypothesis of trivial effects was supported also for the COVID-19 related use of important social media channels.
+
+That said, one effect differed substantially from zero.
+Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly higher levels of negative affect than usual (\emph{b} = 0.01 {[}95\% CI \textless{} 0.01, 0.03{]}).
+However, the effect was still completely inside of the null region, hence likely not large enough to be considered meaningful.
+
+For an overview of all within-person effects, see Table \ref{tab:tab-within} and Figure \ref{fig:fig-within}.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_activity} \caption{The effects of various types of social media use on three indicators of well-being.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-within-1} \caption{Within-person effects of COVID-19 related social media use on well-being.
+Note.
 The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
 The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-activity}
+Hence, all of the reported effects are not considered meaningful.}\label{fig:fig-within}
 \end{figure}
 
-Regarding between-person relations, about which no hypotheses were formulated, only three effects did not include zero.
-Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (\emph{b} = -0.03 {[}95\% CI \textgreater{} -0.01, -0.06{]}).
-Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported higher levels of negative affect than others (\emph{b} = 0.06 {[}95\% CI 0.09, 0.03{]}).
-Interestingly, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect than others (\emph{b} = 0.05 {[}95\% CI 0.09, \textless{} 0.01{]}).
-However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant.
-
-Note that when comparing the results with and without control variables, the results differed.
-For example, on the between-person level, one effect stopped being significant if controlled for additional variables.
-Actively posting on social media was significantly (though not meaningfully) related to decreased life satisfaction.
-However, when controlling for potential confounders, the effect became virtually zero (see Figure \ref{fig:res-channels}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_channel} \caption{The effects of using various social media applications on three indicators of well-being.
-The black estimates show the effects controlled for a large number of covariates (see text; preregistered); the grey estimates are without control variables (exploratory).
-The SESOI was b = |0.30| for life satisfaction and b = |0.15| for affect.
-Hence, all of the reported effects are not considered meaningful.}\label{fig:res-channels}
-\end{figure}
-
-Regarding the COVID-19 related use of social media \emph{channels}, the results were comparable (see Figure \ref{fig:res-channels}).
-Changes in the frequency of using different social media channels to attain information regarding COVID-19 were unrelated to meaningful changes in well-being.
-For example, respondents who used Facebook more frequently than usual to learn about COVID-19 did not show a simultaneous change in well-being (\emph{b} = -0.05 {[}95\% CI -0.11, 0.01{]}).
-Only two effects differed substantially from zero.
-Respondents who used Instagram more frequently than usual to attain COVID-19 related news reported slightly \emph{higher} levels of life satisfaction than usual (\emph{b} = 0.09 {[}95\% CI 0.02, 0.16{]}).
-Respondents who used Twitter more frequently than usual to attain COVID-19 related news reported slightly \emph{lower} levels of life satisfaction than usual (\emph{b} = -0.12 {[}95\% CI -0.23, -0.02{]}).
-However, both effects were still completely inside of the null region, hence not large enough to be considered meaningful.
-In sum, the hypothesis was supported also for the COVID-19 related use of important social media channels.
-
-In terms of between-person relations---which, again, weren't included in the hypotheses---no relations crossed the null region or fell outside of it.
-Only one relation did not include zero, was hence statistically significant.
-Respondents who across all waves used YouTube more frequently than others for COVID-19 related reasons reported marginally \emph{higher} levels of life satisfaction (\emph{b} = 0.08 {[}95\% CI \textless{} 0.01, 0.16{]}).
-However, please note that this effect again was not large enough to be considered practically relevant.
-
-Again, note that when comparing the results with and without control variables, the results differed.
-Especially on the between-person level, altogether five effects stopped being significant if they were controlled for additional variables.
-For example, using Instagram was significantly (though not meaningfully) related to increased life satisfaction.
-However, when controlling for additional covariates, the effect became virtually zero (see Figure \ref{fig:res-channels}).
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tab-within}Overview of all within-person effects.}
+
+\footnotesize{
+
+\begin{tabular}{lrrrrr}
+\toprule
+ &  & \multicolumn{2}{c}{Confidence interval}  &  &\\
+\cmidrule(r){3-4}
+Predictor & \multicolumn{1}{c}{b} & \multicolumn{1}{c}{Lower} & \multicolumn{1}{c}{Higher} & \multicolumn{1}{c}{beta} & \multicolumn{1}{c}{p}\\
+\midrule
+Life satisfaction &  &  &  &  & \\
+\ \ \ Reading & 0.03 & -0.01 & 0.07 & 0.02 & .091\\
+\ \ \ Liking \& Sharing & -0.02 & -0.08 & 0.03 & -0.01 & .299\\
+\ \ \ Posting & -0.05 & -0.11 & 0.01 & -0.02 & .114\\
+\ \ \ Facebook & 0.00 & -0.04 & 0.04 & 0.00 & .964\\
+\ \ \ Instagram & 0.01 & -0.01 & 0.03 & 0.01 & .226\\
+\ \ \ WhatsApp & 0.00 & -0.03 & 0.02 & 0.00 & .664\\
+\ \ \ YouTube & 0.02 & -0.03 & 0.06 & 0.01 & .402\\
+\ \ \ Twitter & -0.02 & -0.09 & 0.05 & -0.01 & .420\\
+Positive affect &  &  &  &  & \\
+\ \ \ Reading & -0.03 & -0.05 & -0.01 & -0.04 & .011\\
+\ \ \ Liking \& Sharing & 0.01 & -0.02 & 0.04 & 0.01 & .535\\
+\ \ \ Posting & 0.01 & -0.02 & 0.05 & 0.01 & .404\\
+\ \ \ Facebook & -0.01 & -0.04 & 0.02 & -0.01 & .581\\
+\ \ \ Instagram & 0.00 & -0.02 & 0.02 & 0.00 & .704\\
+\ \ \ WhatsApp & 0.01 & -0.02 & 0.04 & 0.01 & .444\\
+\ \ \ YouTube & 0.00 & -0.04 & 0.04 & 0.00 & .904\\
+\ \ \ Twitter & 0.01 & 0.00 & 0.02 & 0.01 & .129\\
+Negative affect &  &  &  &  & \\
+\ \ \ Reading & 0.00 & -0.01 & 0.01 & 0.00 & .971\\
+\ \ \ Liking \& Sharing & 0.02 & 0.00 & 0.03 & 0.02 & .019\\
+\ \ \ Posting & 0.05 & 0.02 & 0.08 & 0.04 & .009\\
+\ \ \ Facebook & -0.01 & -0.02 & 0.01 & -0.01 & .226\\
+\ \ \ Instagram & 0.00 & -0.01 & 0.01 & 0.00 & .371\\
+\ \ \ WhatsApp & 0.00 & -0.01 & 0.02 & 0.01 & .562\\
+\ \ \ YouTube & 0.01 & -0.01 & 0.04 & 0.02 & .250\\
+\ \ \ Twitter & 0.01 & 0.00 & 0.03 & 0.01 & .034\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-In what follows, I briefly report some exploratory analyses that weren't preregistered.
-First, to contextualize the results reported above and to see if the results included any meaningful effects at all, I also looked at the effect sizes of selected (cherry-picked) covariates.
-Because each variable had different response options, we would actually need to define a SESOI for each variable, which for reasons of scope I cannot implement here.
-Therefore, I report the results of the standardized scales,
-which allows for a better comparison across the differently scaled variables.
-For what it's worth, as a rough estimate for the SESOI we can build on the typical convention that small effects start at \emph{r} = \textbar.10\textbar.
-The results showed that several effects fell outside of the SESOI, were hence considered meaningful.
-This includes for example internal locus of control, health, satisfaction with democracy, or exercising.
-For an overview, see Figure \ref{fig:res-control}.
+To contextualize the results reported above and to see if the study included any meaningful effects at all, I also looked at the effect sizes of the covariates.
+Because each variable featured different response options, which would require defining a SESOI for each variable, I hence report the results of the standardized scales, which allows for a better comparison across the differently scaled variables.
+Here, we can build on Cohen's convention that small effects begin at \emph{r} = \textbar.10\textbar.
+
+The results showed that several effects crossed or fell completely outside of the SESOI.
+They can hence be considered meaningful.
+For example, if physical health decreased, this had a meaningful detrimental impact on life satisfaction (\(\beta\) = .18 {[}95\% CI .16, .20{]}), positive affect (\(\beta\) = .17 {[}95\% CI .16, .18{]}), and negative affect (\(\beta\) = -.18 {[}95\% CI -.19, -.17{]}).
+Spending more time outside to exercise meaningfully increased positive affect (\(\beta\) = .12 {[}95\% CI .11, .13{]}).
+Being more satisfied with the current democratic system meaningfully increased life satisfaction (\(\beta\) = .13 {[}95\% CI .12, .14{]}).
+The strongest aspect affecting well-being was internal locus of control.
+If people felt more in control of their lives, this strongly increased both life satisfaction (\(\beta\) = .31 {[}95\% CI .28, .34{]}) and
+positive affect (\(\beta\) = .27 {[}95\% CI .25, .29{]}),
+while decreasing negative affect (\(\beta\) = -.28 {[}95\% CI -.30, -.26{]}).
+
+For an overview, see Figure \ref{fig:fig-control}.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/fig_results_control_std} \caption{Results of selected covariates.
-All variables were standardize except 'Male'  and ‘Employed in public service', because there were measured on a binary scale.}\label{fig:res-control}
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/fig-control-1} \caption{Results of main variables together with covariates to provide context.
+All variables standardized.
+SESOI: beta = |.10|}\label{fig:fig-control}
 \end{figure}
-
-To find out whether my inferences were robust across legitimate (though arguably inferior) alternative analyses, I reran the analyses also using standardized estimates, mean scores instead of factor scores, and with a data set where missing data were not imputed.
-The results were virtually the same.
-For example, all standardized COVID-19 related types of social media use or channels were not significantly larger than a SESOI of \(\beta\) = \textbar.10\textbar.
-The additional analyses are reported in the \href{https://tdienlin.github.io/Austrian_Corona_Panel/analyses_additional.html}{companion website}.
 
@@
 
-In this study I analyzed the effects of COVID-19 related social media use on well-being.
-The data come from a panel study with 24 waves and are representative of the Austrian population.
-In a random effects model I separated between person relations from within-person effects and controlled for a large number of both stable and varying covariates, thereby aiming to assess causality.
-The results showed that within-person effects were trivial.
-People who used social media more than usual to learn about COVID-19 didn't show meaningful changes in their well-being.
-
-The results imply that COVID-19 related social media use doesn't seem to be particularly relevant for well-being.
-Other factors among the third variables that were measured revealed larger effects or relations, suggesting that well-being is determined by alternative aspects such as health, satisfaction with democracy, locus of control, or exercising.
-According to this study, popular fears that ``doomscrolling'' or overusing social media during times of crises don't seem to be justified.
-
-On the one hand, the results are not aligned with several recent studies analyzing similar or closely related research questions.
-This includes a study by Bendau et al. (2021), which showed negative relations between social media and well-being.
-However, note that Bendau et al. (2021) analyzed cross-sectional data on a between-person level while not controlling for third variables, which is not optimal for investigating causal effects.
-On the other hand, the results are well-aligned with recent studies and meta-analyses analyzing the effects of social media use from a more general perspective or from a somewhat different angle.
-These studies have found that the effects of various types of social media use on several well-being indicators are small at best, often too small to matter (Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), which echoes the results obtained here.
-
-If anything, two preliminary and subtle trends can be observed.
-First, of all the three COVID-19 related social media activities, people who read about the pandemic more than others showed slightly decreased levels of positive affect, and people who actively posted about the pandemic more than others showed slightly increased levels of negative affect.
-On the other hand, however, people who posted more about COVID-19 also showed slightly higher levels of positive affects, so taken together the results are ambivalent.
-Second, in terms of media channels, using Twitter more than usual was related to slightly decreased levels of life satisfaction.
-Twitter is considered to have more negative affordances and tonality as compared to other networks such as Instagram (Waterloo, Baumgartner, Peter, \& Valkenburg, 2018), which might help explain the results.
-Instagram, on the other hand, was related to slightly increased levels of life satisfaction.
-To speculate, the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.
-That said, all these effects were still very small and arguably too small to matter.
-But future research might elaborate on these specific relations to probe their stability and relevance.
-
-Finally, another interesting observation is that life satisfaction was remarkably stable.
-Hence, even in times of a pandemic, it seems that such broad assessment of life vary only mildly.
-This supports the hypothesis that life satisfaction seems to be determined largely by stable factors such as one's genes (Brown \& Rohrer, 2019).
+Based on a panel study with 32 waves largely representative of the Austrian population, this study analyzed the effects of COVID-19 related social media use on well-being.
+Between person correlation analyses showed that more active users of COVID-19 related content on social media also reported decreased well-being.
+The within-person relations, which are more informative regarding causal effects, showed a different pattern, however.
+If people consumed more COVID-19 content on social media than usual, this did not meaningfully reduce their well-being.
+Several statistically significant effects were found, but these were very small.
+For example, people who liked and shared more COVID-19 related posts than usual reported slightly higher levels of negative affect.
+Posting more content about COVID-19 than usual slightly increased negative affect.
+People who read more COVID-19 related posts than usual reported slightly decreased positive affect.
+Using Twitter for COVID-19 related content slightly increased negative affect.
+Again, although all of these within-person effects were statistically significant, they were very small, smaller than the predefined smallest effect size of interest.
+According to the preregistered procedure, they should hence be considered irrelevant.
+Other factors, for which we would expect to find meaningful effects, such as health or sports, indeed showed substantial and meaningful impacts on well-being.
+In conclusion, COVID-19 related activity on social media was not a particularly strong influence on peoples' well-being.
+The results do not support the popular fears that ``doomscrolling'' or overusing social media during times of crises constitutes a relevant risk for well-being.
+
+These specific observations notwithstanding, several general trends can be observed.
+First, overall the results do suggest that effects of COVID-19 related social media use on well-being tend to take place in the negative as opposed to the positive spectrum.
+Although very small, four statistically significant negative results of COVID-19 related social media use on well-being were found.
+Not a single positive effect emerged.
+Also note that in the analyses several control variables were included, ruling out plausible alternative explanations for the negative results.
+For example, it was controlled for whether or not participants contracted a COVID-19 infection during a specific wave.
+Hence, we can rule out the alternative explanation that having an infection was the root cause of increased communication and reduced well-being.
+
+Second, significant outcomes emerged for positive or negative affect, but not for life satisfaction.
+Life satisfaction was stable and not affected by any type or channel of social media communication.
+The more fluctuating positive and negative affect, however, were affected.
+Liking, sharing, and posting COVID-19 related content, and spending time on Twitter to browse COVID-19 related content, all negatively influenced affect.
+This is aligned with prior findings which showed that social media can trigger negative affect but does not reduce life satisfaction (Huang, 2017).
+Conversations about COVID-19 on social media are often extreme, negative, or aggressive (Fan et al., 2020).
+More deeply engaging with this type of content could negatively affect active authors.
+The hypothesis that tonality could explain the negative effects is especially supported by the observation that spending more time on Twitter than usual increased negative affect.
+Communication on Twitter is often found to be more negative and impolite compared to other SNSs (Halpern \& Gibbs, 2013).
+Consuming more negative and misleading information could hence explain the (slightly) increased levels of negative affect.
+
+Third, the results show that it makes sense to analyze different communication types and communication channels.
+Reading reduced positive affect, while liking, sharing and commenting increased negative affect.
+Whereas it was often stated that passive use is bad and active use good (Verduyn et al., 2015), this pattern thus did not emerge here.
+Instead, all three types were negative.
+The results hence support the findings from Valkenburg et al. (2022), who also could not confirm the claim that active use is good and that passive use is bad.
+Focusing on communication channels, Twitter seems to be more negative, as has often been observed (Halpern \& Gibbs, 2013).
+But, again, all of these effects are very small.
+Future research might elaborate on these specific relations to probe their stability and relevance.
+
+Taken together, the results are hence aligned with the underlying theoretical models and prior empirical results.
+The findings support the Different Susceptibility to Media Effects Theory, which states that effects are generally small and depend on the type of communication.
+The results are well-aligned with mood management theory (Zillmann, 1988) or the uses and gratifications approach (Katz et al., 1973), whose premises preclude particularly negative effects of routine and widespread media consumption.
+If the effects of social media were indeed profoundly negative on average, then people likely would not spend so much time on social media engaging with COVID-19 content.
+Likewise, recent empirical studies and meta-analyses reported rather small negative effects, too.
+Several studies found that the effects of various types of social media use on well-being are small, often too small to matter (Bendau et al., 2021; Ferguson et al., 2021; Meier \& Reinecke, 2020; Orben, 2020), echoing the results obtained here.
+
+In light of the overall very small effects, engaging in COVID 19-related social media use should not be a major concern for one's well-being.
+That said, the results still imply that it does make sense to critically reflect upon COVID-19 related social media use.
+On average, it is likely more beneficial to post less actively about COVID-19 on social media and to spend less time on Twitter consuming COVID-19 related content.
+The results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.
 
@@
 
-The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders.
-Together, this allows for an improved perspective on assessing causality.
-That said, causality necessitates temporal order, and the cause needs to precede the effect.
-Regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
-In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
-However, even then we don't know for certain if we actually measured the right interval.
-Effects depend on the intensity of use or the length of the interval, and to borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
-Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
-In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.
-
-Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
+Focusing on within-person effects and controlling for several potential confounders, this study provides an improved perspective on assessing causality.
+However, three major challenges remain.
+First, it could be that the timing of the interval was wrong; it could be that effects unfold in shorter or longer intervals.
+Next, it could be that there was not sufficient variance in media use and well-being.
+Without sufficient variability, statistical models cannot detect existing effects.
+Third, it is necessary to control for \emph{all} relevant confounding third variables.
+Although this study included are large list of confounders, it could still be that crucial variables were missed (Zillmann, 1988).
+To borrow the words from Rohrer and Murayama (2021), there is no such thing as ``the'' effect of social media use on well-being.
+The results reported here are hence contingent on the design I used.
+To document how effects unfold, future research needs to employ different study designs probing different intervals.
+More thought needs to be invested in what factors to control for, and I hope this study provides a first step into this direction.
+
+Although I had already reduced the predefined SESOIs to be less conservative, one could argue they were still too large.
 Media use is only one aspect of several factors that simultaneously affect well-being.
-Is it really realistic to expect that extremely changing only \emph{one} of these aspects should manifest in a detectable change in well-being?
+Is it realistic to expect that changing only \emph{one} of these aspects should already manifest in a detectable change in well-being?
 Or would it make more sense to expect that thoroughly committing to say \emph{two} activities (e.g.~regularly exercising \emph{and} establishing a reading habit) should then cause a detectable improvement in well-being?
-Practically, this would imply a SESOI half as large as I have defined here, namely \emph{b} = \textbar.15\textbar{} for well-being and \emph{b} = \textbar.075\textbar{} for affect.
-In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effect would still be completely in the null region, and no effect would be outside of the null region.
-However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
-Again, with this study I hope to provide some first input and guidelines.
+Practically, this would imply a SESOI half the size defined here, namely \emph{b} = \textbar.15\textbar{} for life satisfaction and \emph{b} = \textbar.075\textbar{} for affect.
+In the case of this study, however, even halving the SESOI would not make a difference.
+All but one effect would still be completely in the null region, and no effect would fall completely outside of the null region.
+I encourage future research to elaborate on what effect sizes are considered meaningful and what not.
 
 Both media use and well-being were measured using self-reports.
-Measuring well-being with self-reports is adequate, because it by definition requires introspection.
-However, it would be preferable to measure social media use objectively, because people cannot reliably estimate their use (Scharkow, 2016).
-That said, objective measures often cannot capture the content or the motivation of the use, and only very complicated tools recording the actual content (such as the Screenome project) might produce such data.
-Unfortunately, such procedures introduce other problems, especially related to privacy.
-Hence, for this type of research question it still seems necessary to use self-reported measures.
-
-Because the data were collected in a single country, the generalizability of the results is limited.
-The results apply primarily to the more Western sphere, and might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
-That said, because this is a comparatively large study representative of an entire country, and because several waves were collected across a large time span, the results should be at least as generalizable as other typical empirical studies collected in the social sciences.
+Because assessing well-being necessarily requires introspection, using self-reports for affect and life satisfaction is adequate.
+However, for social media use objective measures are clearly preferable, as people often cannot reliably estimate their use (Scharkow, 2016).
+At the same time, most objective measures cannot capture the content or the motivation of use.
+Hence, for the type of research question analyzed here, it still seems necessary to use self-reported measures.
+In many cases they can still be informative (Verbeij et al., 2021).
+
+Being collected in a single country, the generalizability of the results is limited.
+The results apply primarily to the more Western sphere.
+They might not hold true in other cultures, especially cultures with a different media landscape or alternative social media channels.
 
@@
 
-In this study, COVID-19 related social media use didn't meaningfully affect several indicators of well-being, including life satisfaction, positive affect, and negative affect.
-However, factors other than social media use were meaningfully related to well-being, such as physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
-If it's our aim to improve well-being, it might hence be more fruitful not to focus so much on social media but to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
+In this study, COVID-19 related social media use did not meaningfully affect well-being.
+Very small negative effects were found for writing COVID-19 related posts, sharing COVID-19 related content, and spending more time than usual on Twitter.
+Factors other than social media use, however, were meaningfully related to well-being, including physical health, exercise, satisfaction with democracy, or believing that one is in control of one's life.
+Hence, when trying to improve well-being during a pandemic, instead of focusing on social media it seems more fruitful to address other, more pertinent societal problems related to health care, regular exercise, or a functioning democratic system.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-baguleyStandardizedSimpleEffect2009}{}%
-Baguley, T. (2009). Standardized or simple effect size: {What} should be reported? \emph{British Journal of Psychology}, \emph{100}(3), 603--617. \url{https://doi.org/10.1348/000712608X377117}
-
-\leavevmode\hypertarget{ref-bellFixedRandomEffects2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bellFixedRandomEffects2019}{}}%
 Bell, A., Fairbrother, M., \& Jones, K.
@@
 
-\leavevmode\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bendauAssociationsCOVID19Related2021}{}}%
 Bendau, A., Petzold, M.
-B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., \ldots{} Plag, J.
+B., Pyrkosch, L., Mascarell Maricic, L., Betzler, F., Rogoll, J., Große, J., Ströhle, A., \& Plag, J.
 (2021).
-Associations between {COVID}-19 related media consumption and symptoms of anxiety, depression and {COVID}-19 related fear in the general population in {Germany}.
+Associations between {COVID-19} related media consumption and symptoms of anxiety, depression and {COVID-19} related fear in the general population in {Germany}.
 \emph{European Archives of Psychiatry and Clinical Neuroscience}, \emph{271}(2), 283--291.
@@
 
-\leavevmode\hypertarget{ref-beyensSocialMediaUse2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-beyensSocialMediaUse2021}{}}%
 Beyens, I., Pouwels, J.
-L., Driel, I.
-I.
-van, Keijsers, L., \& Valkenburg, P.
+L., van Driel, I.
+I., Keijsers, L., \& Valkenburg, P.
 M.
@@
 \emph{Communication Research}.
-\url{https://doi.org/10.31234/osf.io/ftygp}
-
-\leavevmode\hypertarget{ref-brownEasyHappinessPie2019}{}%
-Brown, N.
-J.
-L., \& Rohrer, J.
-M.
-(2019).
-Easy as (happiness) pie?
-{A} critical evaluation of a popular model of the determinants of well-being.
-\emph{Journal of Happiness Studies}.
-\url{https://doi.org/10.1007/s10902-019-00128-4}
-
-\leavevmode\hypertarget{ref-choiMediatedCommunicationMatters2021}{}%
+\url{https://doi.org/10.1177/00936502211038196}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bradleyStressMoodSmartphone2021}{}}%
+Bradley, A., \& Howard, A. (2021). \emph{Stress, mood, and smartphone use in {University} students: {A} 12-week longitudinal study}. \url{https://doi.org/10.31219/osf.io/frvpb}
+
+\leavevmode\vadjust pre{\hypertarget{ref-buchiDigitalWellbeingTheory2021}{}}%
+Büchi, M. (2021). Digital well-being theory and research. \emph{New Media \& Society}, 146144482110568. \url{https://doi.org/10.1177/14614448211056851}
+
+\leavevmode\vadjust pre{\hypertarget{ref-choiMediatedCommunicationMatters2021}{}}%
 Choi, M., \& Choung, H.
 (2021).
-Mediated communication matters during the {COVID}-19 pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
+Mediated communication matters during the {COVID-19} pandemic: {The} use of interpersonal and masspersonal media and psychological well-being.
 \emph{Journal of Social and Personal Relationships}, \emph{38}(8), 2397--2418.
@@
 
-\leavevmode\hypertarget{ref-cohenPowerPrimer1992}{}%
-Cohen, J. (1992). A power primer. \emph{Psychological Bulletin}, \emph{112}(1), 155--159. \url{https://doi.org/10.1037/0033-2909.112.1.155}
-
-\leavevmode\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dienerAdvancesOpenQuestions2018}{}}%
 Diener, E., Lucas, R.
@@
 
-\leavevmode\hypertarget{ref-dienesUsingBayesGet2014}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dienesUsingBayesGet2014}{}}%
 Dienes, Z. (2014). Using {Bayes} to get the most out of non-significant results. \emph{Frontiers in Psychology}, \emph{5}. \url{https://doi.org/10.3389/fpsyg.2014.00781}
 
-\leavevmode\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dienlinImpactDigitalTechnology2020}{}}%
 Dienlin, T., \& Johannes, N.
@@
 
-\leavevmode\hypertarget{ref-edenMediaCopingCOVID192020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dienlinDisplacementReinforcementReciprocity2017}{}}%
+Dienlin, T., Masur, P.
+K., \& Trepte, S.
+(2017).
+Displacement or reinforcement?
+{The} reciprocity of {FtF}, {IM}, and {SNS} communication and their effects on loneliness and life satisfaction.
+\emph{Journal of Computer-Mediated Communication}, \emph{22}(2), 71--87.
+\url{https://doi.org/10.1111/jcc4.12183}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dormannOptimalTimeLags2015}{}}%
+Dormann, C., \& Griffin, M. A. (2015). Optimal time lags in panel studies. \emph{Psychological Methods}, \emph{20}(4), 489--505. \url{https://doi.org/10.1037/met0000041}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dornemannHowGoodBad2021}{}}%
+Dörnemann, A., Boenisch, N., Schommer, L., Winkelhorst, L., \& Wingen, T.
+(2021).
+\emph{How do good and bad news impact mood during the {Covid-19} pandemic?
+{The} role of similarity}.
+\url{https://doi.org/10.31219/osf.io/sy2kd}
+
+\leavevmode\vadjust pre{\hypertarget{ref-edenMediaCopingCOVID192020}{}}%
 Eden, A.
@@
 (2020).
-Media for coping during {COVID}-19 social distancing: {Stress}, anxiety, and psychological well-being.
+Media for coping during {COVID-19} social distancing: {Stress}, anxiety, and psychological well-being.
 \emph{Frontiers in Psychology}, \emph{11}, 577639.
@@
 
-\leavevmode\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-europeansocialsurveyESS9Edition20182021}{}}%
+European Social Survey. (2021). \emph{{ESS9} edition 3.1 - 2018 {Documentation Report}}. \url{https://www.europeansocialsurvey.org/docs/round9/survey/ESS9_data_documentation_report_e03_1.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fanStigmatizationSocialMedia2020}{}}%
+Fan, L., Yu, H., \& Yin, Z.
+(2020).
+Stigmatization in social media: {Documenting} and analyzing hate speech for {\textsc{COVID}} ‐19 on {Twitter}.
+\emph{Proceedings of the Association for Information Science and Technology}, \emph{57}(1).
+\url{https://doi.org/10.1002/pra2.313}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fergusonThisMetaanalysisScreen2021}{}}%
 Ferguson, C.
@@
 K., Branley-Bell, D., Markey, P., Ivory, J.
-D., Klisanin, D., \ldots{} Wilson, J.
+D., Klisanin, D., Elson, M., Smyth, M., Hogg, J.
+L., McDonnell, D., Nichols, D., Siddiqui, S., Gregerson, M., \& Wilson, J.
 (2021).
@@
 
-\leavevmode\hypertarget{ref-funderEvaluatingEffectSize2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-funderEvaluatingEffectSize2019}{}}%
 Funder, D.
@@
 
-\leavevmode\hypertarget{ref-greenspoonIntegrationSubjectiveWellbeing2001}{}%
-Greenspoon, P.
-J., \& Saklofske, D.
-H.
-(2001).
-Toward an integration of subjective well-being and psychopathology.
-\emph{Social Indicators Research}, \emph{54}(1), 81--108.
-\url{https://doi.org/10.1023/A:1007219227883}
-
-\leavevmode\hypertarget{ref-hamakerWhyResearchersShould2014}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-guazziniSecondWaveAnalysis2022}{}}%
+Guazzini, A., Pesce, A., Marotta, L., \& Duradoni, M.
+(2022).
+Through the second wave: {Analysis} of the psychological and perceptive changes in the {Italian} population during the {COVID-19} pandemic.
+\emph{International Journal of Environmental Research and Public Health}, \emph{19}(3), 1635.
+\url{https://doi.org/10.3390/ijerph19031635}
+
+\leavevmode\vadjust pre{\hypertarget{ref-halpernSocialMediaCatalyst2013}{}}%
+Halpern, D., \& Gibbs, J.
+(2013).
+Social media as a catalyst for online deliberation?
+{Exploring} the affordances of {Facebook} and {YouTube} for political expression.
+\emph{Computers in Human Behavior}, \emph{29}(3), 1159--1168.
+\url{https://doi.org/10.1016/j.chb.2012.10.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hamakerWhyResearchersShould2014}{}}%
 Hamaker, E.
@@
 Csikszentmihalyi (Eds.), \emph{Handbook of research methods for studying daily life} (Paperback ed.).
-New York, NY: Guilford.
-
-\leavevmode\hypertarget{ref-huangTimeSpentSocial2017}{}%
+{Guilford}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-huangTimeSpentSocial2017}{}}%
 Huang, C.
@@
 
-\leavevmode\hypertarget{ref-kerestesAdolescentsOnlineSocial2020}{}%
-Keresteš, G., \& Štulhofer, A.
-(2020).
-Adolescents' online social network use and life satisfaction: {A} latent growth curve modeling approach.
-\emph{Computers in Human Behavior}, \emph{104}, 106187.
-\url{https://doi.org/10.1016/j.chb.2019.106187}
-
-\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2020}{}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
-(2020).
-\emph{Austrian {Corona} {Panel} {Project} ({SUF} edition)}.
-AUSSDA.
-\url{https://doi.org/10.11587/28KQNS}
-
-\leavevmode\hypertarget{ref-kittelAustrianCoronaPanel2021}{}%
-Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., \ldots{} Schlogl, L.
+\leavevmode\vadjust pre{\hypertarget{ref-katzUsesGratificationsResearch1973}{}}%
+Katz, E., Blumler, J. G., \& Gurevitch, M. (1973). Uses and {Gratifications Research}. \emph{Public Opinion Quarterly}, \emph{37}(4), 509. \url{https://doi.org/10.1086/268109}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2021}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
 (2021).
-The {Austrian} {Corona} {Panel} {Project}: Monitoring individual and societal dynamics amidst the {COVID}-19 crisis.
+The {Austrian Corona Panel Project}: Monitoring individual and societal dynamics amidst the {COVID-19} crisis.
 \emph{European Political Science}, \emph{20}(2), 318--344.
@@
 
-\leavevmode\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}%
-Klein, J. (2021). \emph{The darkly soothing compulsion of 'doomscrolling'}. Retrieved from \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
-
-\leavevmode\hypertarget{ref-klinePrinciplesPracticeStructural2016}{}%
-Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling} (4th ed.). New York, NY: The Guilford Press.
-
-\leavevmode\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-kittelAustrianCoronaPanel2020}{}}%
+Kittel, B., Kritzinger, S., Boomgaarden, H., Prainsack, B., Eberl, J.-M., Kalleitner, F., Lebernegg, N.
+S., Partheymüller, J., Plescia, C., Schiestl, D.
+W., \& Schlogl, L.
+(2020).
+\emph{Austrian {Corona Panel Project} ({SUF} edition)} {[}Data set{]}.
+{AUSSDA}.
+\url{https://doi.org/10.11587/28KQNS}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kleinDarklySoothingCompulsion2021}{}}%
+Klein, J. (2021, March 3). \emph{The darkly soothing compulsion of 'doomscrolling'}. \url{https://www.bbc.com/worklife/article/20210226-the-darkly-soothing-compulsion-of-doomscrolling}
+
+\leavevmode\vadjust pre{\hypertarget{ref-latikkaLonelinessPsychologicalDistress2022}{}}%
+Latikka, R., Koivula, A., Oksa, R., Savela, N., \& Oksanen, A.
+(2022).
+Loneliness and psychological distress before and during the {COVID-19} pandemic: {Relationships} with social media identity bubbles.
+\emph{Social Science \& Medicine}, \emph{293}, 114674.
+\url{https://doi.org/10.1016/j.socscimed.2021.114674}
+
+\leavevmode\vadjust pre{\hypertarget{ref-liuRelationOfficialWhatsAppdistributed2020}{}}%
 Liu, J.
@@
 (2020).
-The relation between official {WhatsApp}-distributed {COVID}-19 news exposure and psychological symptoms: Cross-sectional survey study.
+The relation between official {WhatsApp-distributed COVID-19} news exposure and psychological symptoms: Cross-sectional survey study.
 \emph{Journal of Medical Internet Research}, \emph{22}(9), e22142.
@@
 
-\leavevmode\hypertarget{ref-mcelreathYesterdayClass2021}{}%
-McElreath, R. (2021). Yesterday in class, ... {[}Tweet{]}. Retrieved from \url{https://twitter.com/rlmcelreath/status/1354786005996482563}
-
-\leavevmode\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lucasAdaptationSetpointModel2007}{}}%
+Lucas, R.
+E.
+(2007).
+Adaptation and the set-point model of subjective well-being.
+\emph{Current Directions in Psychological Science}, \emph{16}(2), 75--79.
+\url{https://doi.org/10.1111/j.1467-8721.2007.00479.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lucasWhyCrossLaggedPanel2022}{}}%
+Lucas, R. E. (2022). \emph{Why the {Cross-Lagged Panel Model} is {Almost Never} the {Right Choice}} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/pkec7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lykkenHappinessWhatStudies1999}{}}%
+Lykken, D. T. (1999). \emph{Happiness: {What} studies on twins show us about nature, nurture, and the happiness set-point}. {Golden Books}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marcianoDynamicsAdolescentsSmartphone2022}{}}%
+Marciano, L., Driver, C.
+C., Schulz, P.
+J., \& Camerini, A.-L.
+(2022).
+Dynamics of adolescents' smartphone use and well-being are positive but ephemeral.
+\emph{Scientific Reports}, \emph{12}(1), 1316.
+\url{https://doi.org/10.1038/s41598-022-05291-y}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meierDoesPassiveSocial2022}{}}%
+Meier, A., \& Krause, H.-V. (2022). \emph{Does passive social media use harm well-being? {An} adversarial review} {[}Preprint{]}. {PsyArXiv}. \url{https://doi.org/10.31234/osf.io/nvbwh}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meierComputermediatedCommunicationSocial2020a}{}}%
 Meier, A., \& Reinecke, L.
@@
 
-\leavevmode\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-normanInterpretationChangesHealthrelated2003}{}}%
 Norman, G., Sloan, J., \& Wyrwich, K.
@@
 \emph{Medical Care}, \emph{41}(5), 582--592.
-Retrieved from \url{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}
-
-\leavevmode\hypertarget{ref-orbenTeenagersScreensSocial2020}{}%
+\href{Retrieved\%20from\%20http://www.jstor.org/stable/3768017}{Retrieved from http://www.jstor.org/stable/3768017}
+
+\leavevmode\vadjust pre{\hypertarget{ref-orbenTeenagersScreensSocial2020}{}}%
 Orben, A.
@@
 
-\leavevmode\hypertarget{ref-orbenSocialMediaEnduring2019}{}%
-Orben, A., Dienlin, T., \& Przybylski, A.
-K.
-(2019).
-Social media's enduring effect on adolescent life satisfaction.
-\emph{Proceedings of the National Academy of Sciences of the United States of America}, \emph{116}(21), 10226--10228.
-\url{https://doi.org/10.1073/pnas.1902058116}
-
-\leavevmode\hypertarget{ref-przybylskiDoesTakingShort2021a}{}%
-Przybylski, A.
-K., Nguyen, T.
-T., Law, W., \& Weinstein, N.
-(2021).
-Does taking a short break from social media have a positive effect on well-being?
-{Evidence} from three preregistered field experiments.
-\emph{Journal of Technology in Behavioral Science}, \emph{6}(3), 507--514.
-\url{https://doi.org/10.1007/s41347-020-00189-w}
-
-\leavevmode\hypertarget{ref-przybylskiLargescaleTestGoldilocks2017}{}%
-Przybylski, A. K., \& Weinstein, N. (2017). A large-scale test of the {Goldilocks} hypothesis. \emph{Psychological Science}, \emph{28}(2), 204--215. \url{https://doi.org/10.1177/0956797616678438}
-
-\leavevmode\hypertarget{ref-riehmAssociationsMediaExposure2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pelletierOneSizeDoesn2020}{}}%
+Pelletier, M.
+J., Krallman, A., Adams, F.
+G., \& Hancock, T.
+(2020).
+One size doesn't fit all: A uses and gratifications analysis of social media platforms.
+\emph{Journal of Research in Interactive Marketing}, \emph{14}(2), 269--284.
+\url{https://doi.org/10.1108/JRIM-10-2019-0159}
+
+\leavevmode\vadjust pre{\hypertarget{ref-riehmAssociationsMediaExposure2020}{}}%
 Riehm, K.
 E., Holingue, C., Kalb, L.
-G., Bennett, D., Kapteyn, A., Jiang, Q., \ldots{} Thrul, J.
+G., Bennett, D., Kapteyn, A., Jiang, Q., Veldhuis, C.
+B., Johnson, R.
+M., Fallin, M.
+D., Kreuter, F., Stuart, E.
+A., \& Thrul, J.
 (2020).
 Associations between media exposure and mental distress among {U}.{S}.
-Adults at the beginning of the {COVID}-19 pandemic.
+Adults at the beginning of the {COVID-19} pandemic.
 \emph{American Journal of Preventive Medicine}, \emph{59}(5), 630--638.
@@
 
-\leavevmode\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-rohrerThinkingClearlyCorrelations2018}{}}%
 Rohrer, J.
@@
 
-\leavevmode\hypertarget{ref-rohrerTheseAreNot2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-rohrerTheseAreNot2021}{}}%
 Rohrer, J.
@@
 (2021).
-\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis} {[}Preprint{]}.
-PsyArXiv.
+\emph{These are not the effects you are looking for: {Causality} and the within-/between-person distinction in longitudinal data analysis}.
+{PsyArXiv}.
 \url{https://doi.org/10.31234/osf.io/tg4vj}
 
-\leavevmode\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-sandstromDoomscrollingCOVIDNews2021}{}}%
 Sandstrom, G., Buchanan, K., Aknin, L., \& Lotun, S.
-(2021).
-Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place.
-Retrieved from \url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
-
-\leavevmode\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}%
+(2021, October 22).
+\emph{Doomscrolling {COVID} news takes an emotional toll -- here's how to make your social media a happier place}.
+{The Conversation}.
+\url{http://theconversation.com/doomscrolling-covid-news-takes-an-emotional-toll-heres-how-to-make-your-social-media-a-happier-place-170342}
+
+\leavevmode\vadjust pre{\hypertarget{ref-scharkowAccuracySelfreportedInternet2016}{}}%
 Scharkow, M.
@@
 
-\leavevmode\hypertarget{ref-schemerImpactInternetSocial2021}{}%
-Schemer, C., Masur, P.
-K., Geiß, S., Müller, P., \& Schäfer, S.
+\leavevmode\vadjust pre{\hypertarget{ref-sewallObjectivelyMeasuredDigital2021}{}}%
+Sewall, C.
+J.
+R., Goldstein, T.
+R., \& Rosen, D.
 (2021).
-The {Impact} of {Internet} and {Social} {Media} {Use} on {Well}-{Being}: {A} {Longitudinal} {Analysis} of {Adolescents} {Across} {Nine} {Years}.
-\emph{Journal of Computer-Mediated Communication}, \emph{26}(1), 1--21.
-\url{https://doi.org/10.1093/jcmc/zmaa014}
-
-\leavevmode\hypertarget{ref-schnauber-stockmannMobileDevicesTools2020}{}%
-Schnauber-Stockmann, A., \& Karnowski, V.
-(2020).
-Mobile devices as tools for media and communication research: {A} scoping review on collecting self-report data in repeated measurement designs.
-\emph{Communication Methods and Measures}, \emph{14}(3), 145--164.
-\url{https://doi.org/10.1080/19312458.2020.1784402}
-
-\leavevmode\hypertarget{ref-stainbackCOVID1924News2020}{}%
+Objectively measured digital technology use during the {COVID-19} pandemic: {Impact} on depression, anxiety, and suicidal ideation among young adults.
+\emph{Journal of Affective Disorders}, \emph{288}, 145--147.
+\url{https://doi.org/10.1016/j.jad.2021.04.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sheldonStabilityHappinessTheories2014}{}}%
+Sheldon, K. M., \& Lucas, R. E. (2014). \emph{Stability of happiness: Theories and evidence on whether happiness can change}. \url{http://site.ebrary.com/id/10891875}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stainbackCOVID1924News2020}{}}%
 Stainback, K., Hearne, B.
@@
 (2020).
-{COVID}-19 and the 24/7 {News} {Cycle}: {Does} {COVID}-19 {News} {Exposure} {Affect} {Mental} {Health}?
+{COVID-19} and the 24/7 {News Cycle}: {Does COVID-19 News Exposure Affect Mental Health}?
 \emph{Socius}, \emph{6}, 2378023120969339.
@@
 
-\leavevmode\hypertarget{ref-statistaAverageDailyTime2021}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-statistaAverageDailyTime2021}{}}%
 Statista.
+(2021, May 21).
+\emph{Average daily time spent on social networks by users in the {United States} from 2018 to 2022}.
+\url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-valkenburgDifferentialSusceptibilityMedia2013}{}}%
+Valkenburg, P. M., \& Peter, J. (2013). The differential susceptibility to media effects model. \emph{Journal of Communication}, \emph{63}(2), 221--243. \url{https://doi.org/10.1111/jcom.12024}
+
+\leavevmode\vadjust pre{\hypertarget{ref-valkenburgAssociationsActivePassive2022}{}}%
+Valkenburg, P.
+M., van Driel, I.
+I., \& Beyens, I.
+(2022).
+The associations of active and passive social media use with well-being: {A} critical scoping review.
+\emph{New Media \& Society}, \emph{24}(2), 530--549.
+\url{https://doi.org/10.1177/14614448211065425}
+
+\leavevmode\vadjust pre{\hypertarget{ref-verbeijSelfreportedMeasuresSocial2021}{}}%
+Verbeij, T., Pouwels, J.
+L., Beyens, I., \& Valkenburg, P.
+M.
 (2021).
-\emph{Average daily time spent on social networks by users in the {United} {States} from 2018 to 2022}.
-Retrieved from \url{https://www.statista.com/statistics/1018324/us-users-daily-social-media-minutes/}
-
-\leavevmode\hypertarget{ref-waterlooNormsOnlineExpressions2018}{}%
-Waterloo, S.
-F., Baumgartner, S.
-E., Peter, J., \& Valkenburg, P.
-M.
+\emph{Self-reported measures of social media use show high predictive validity} {[}Preprint{]}.
+{PsyArXiv}.
+\url{https://doi.org/10.31234/osf.io/c9bj7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-verduynPassiveFacebookUsage2015}{}}%
+Verduyn, P., Lee, D.
+S., Park, J., Shablack, H., Orvell, A., Bayer, J., Ybarra, O., Jonides, J., \& Kross, E.
+(2015).
+Passive {Facebook} usage undermines affective well-being: {Experimental} and longitudinal evidence.
+\emph{Journal of Experimental Psychology: General}, \emph{144}(2), 480--488.
+\url{https://doi.org/10.1037/xge0000057}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wagnerAUTNESOnlinePanel2018}{}}%
+Wagner, M., Aichholzer, J., Eberl, J.-M., Meyer, T.
+M., Berk, N., Büttner, N., Boomgaarden, H., Kritzinger, S., \& Müller, W.
+C.
 (2018).
-Norms of online expressions of emotion: {Comparing} {Facebook}, {Twitter}, {Instagram}, and {WhatsApp}.
-\emph{New Media \& Society}, \emph{20}(5), 1813--1831.
-\url{https://doi.org/10.1177/1461444817707349}
+\emph{{AUTNES Online Panel Study} 2017 ({SUF} edition)} {[}Data set{]}.
+{AUSSDA}.
+\url{https://doi.org/10.11587/I7QIYJ}
+
+\leavevmode\vadjust pre{\hypertarget{ref-worldhealthorganizationWellbeingMeasuresPrimary1998}{}}%
+World Health Organization. (1998). \emph{Wellbeing measures in primary health care/{The Depcare Project}}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-yuePassiveSocialMedia2022}{}}%
+Yue, Z., Zhang, R., \& Xiao, J.
+(2022).
+Passive social media use and psychological well-being during the {COVID-19} pandemic: {The} role of social comparison and emotion regulation.
+\emph{Computers in Human Behavior}, \emph{127}, 107050.
+\url{https://doi.org/10.1016/j.chb.2021.107050}
+
+\leavevmode\vadjust pre{\hypertarget{ref-zillmannMoodManagementCommunication1988}{}}%
+Zillmann, D. (1988). Mood {Management Through Communication Choices}. \emph{American Behavioral Scientist}, \emph{31}(3), 327340. \url{https://doi.org/10.1177/000276488031003005}
 
 \end{CSLReferences}
-
-\newpage
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://tdienlin.github.io/Austrian_Corona_Panel/index.html}).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
 
@@
 
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+I would like to thank BLINDED for providing valuable feedback on this manuscript.
+
 \end{document}

### 2023-01-30 preregistration_changes.Rmd
- major updates to preprint

@@
 
-This includes (a) text-based media news consumption, (b) video-based media news consumption, (c) residency is Vienna, (d) self-reported physical health, (e) living space (in squaremeter), (f) access to balcony, (g) access to garden, (h) employment status, (i) disposition to take risks, and (j) locus of control.
+This includes residency is Vienna, self-reported physical health, living space (in squaremeter), access to balcony, access to garden, and employment status.
 
-The following variables I originally planned to include as control variables.
-However, their inclusion led to the deletion of too many cases/observations.
-I hence decided against their inclusion.
-This includes (a) household size, (b) work hours per week, (c) home office, (d) household income.
+## Imputation
+
+After having received feedback from colleagues who are experts on missing data analyses, I changed the imputation strategy.
+
+- Before: predictive mean matching with n = 1 data-sets; now: multiple imputation (also predictive mean matching) with n = 5 data-sets.
+- Before: remove participants with > 50% missing data; now: include all participants
+
+## Estimation
+
+Positive affect and negative affect were modeled as mean scores and not as factor scores. The reason was that it wasn't possible to extract factor scores for the data set with multiple imputation.

### 2023-01-30 manuscript.Rmd
- Changed URLs

@@
 In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project) (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online [companion website](https://tdienlin.github.io/Austrian_Corona_Panel) (https://tdienlin.github.io/Austrian_Corona_Panel).
 
@@
 I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
-For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project).
-<!-- The additional analyses are reported on the [companion website](https://XMtRA.github.io/Austrian_Corona_Panel_Project/analyses_additional.html). -->
+For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see [companion website](https://tdienlin.github.io/Austrian_Corona_Panel).
 <!-- For example, all within-person standardized COVID-19 related types of social media use or channels were significantly smaller than $\beta$ = |.05|, again supporting that effects were negligible.
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://XMtRA.github.io/Austrian_Corona_Panel_Project).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (https://tdienlin.github.io/Austrian_Corona_Panel).
 

### 2023-01-30 manuscript.tex
- Changed URLs

@@
 In some cases, it was impossible to execute the analyses as I had originally planned, for example because some properties of the variables only became apparent when seeing the actual data.
-The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website} (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
+The most relevant deviations are reported below, and a complete list of all changes can be found in the online \href{https://tdienlin.github.io/Austrian_Corona_Panel}{companion website} (\url{https://tdienlin.github.io/Austrian_Corona_Panel}).
 
@@
 I reran the analyses (a) with additional not-preregistered covariates such as trust in media or government, (b) without covariates, (c) with single imputation, and (d) without imputation.
-For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see \href{https://XMtRA.github.io/Austrian_Corona_Panel_Project}{companion website}.
+For more information on the analyses, a complete documentation of the models and results, and all additional analyses, see \href{https://tdienlin.github.io/Austrian_Corona_Panel}{companion website}.
 
@@
 
-All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://XMtRA.github.io/Austrian_Corona_Panel_Project}).
+All the stimuli, presentation materials, analysis scripts, and a reproducible version of the manuscript can be found on the companion website (\url{https://tdienlin.github.io/Austrian_Corona_Panel}).
 

### 2023-01-30 manuscript.Rmd
- changed abstract

@@
 
-abstract          : "In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
-To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop 'doomscrolling'.
-In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects within between models, controlling for several stable and varying confounders.
-Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded."
+abstract          : "In times of crisis such as the COVID-19 pandemic, citizens need to stay informed about recent political events.
+To this end, people increasingly use social media.
+However, being particularly engaging, many find it hard to disconnect from social media, especially during times of crisis.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 32 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use reduced facets of well-being.
+However, the effects were very small.
+Other factors such as health, income, exercise, or internal locus of control showed larger and more meaningful effects."
 keywords          : "COVID-19, well-being, social media, news use, panel study."

### 2023-01-30 manuscript.tex
- changed abstract

@@
 \abstract{%
-In times of crisis such as the Corona pandemic citizens need to stay informed about recent events, the latest political decisions, or mandatory protection measures.
-To this end, many people use various types of media, and increasingly social media.
-However, because social media are particularly engaging, some find it hard to disconnect and cannot stop `doomscrolling'.
-In this preregistered study, I investigate whether using social media for COVID-19 related reasons affects psychological well-being.
-To answer this question I analyzed data from the Austrian Corona Panel Project, which consists of 24 waves with overall 3,018 participants.
-I ran three random effects within between models, controlling for several stable and varying confounders.
-Results showed that the effects of COVID-19 related social media use on well-being were very small, arguably too small to matter.
-The findings suggest that fears that social media use during times of crisis impairs well-being are likely to be unfounded.
+In times of crisis such as the COVID-19 pandemic, citizens need to stay informed about recent political events.
+To this end, people increasingly use social media.
+However, being particularly engaging, many find it hard to disconnect from social media, especially during times of crisis.
+Using data from the Austrian Corona Panel Project consisting of 3,485 participants from 32 waves, controlling for several stable and varying confounders, the results showed that COVID-19 related social media use reduced facets of well-being.
+However, the effects were very small.
+Other factors such as health, income, exercise, or internal locus of control showed larger and more meaningful effects.
 }
